### 2012-04-03 docs/mixture_models.tex
started repo, outline of mixtures note, readme file, etc

% Copyright 2012 David W. Hogg (NYU).
% All rights reserved.

\documentclass[12pt]{article}

\begin{document}

\section*{Replacing standard galaxy profiles with \\ mixtures of Gaussians}

\noindent
David W. Hogg \\
\textsl{New York University} \\
\textsl{Max-Planck-Institut f\"ur Astronomie} \\[1ex]
Dustin Lang \\
\textsl{Princeton University Observatory}

\begin{abstract}
Exponential, de~Vaucouleurs, and Sersic profiles are simple and
successful models for fitting two-dimensional images of galaxies.  One
numerical issue encountered in this kind of fitting is the pixel
rendering and convolution (or correlation) of the models with the
telescope point-spread function (PSF); these operations are slow, and
easy to get slightly wrong at small radii.  Here we exploit the
realization that these models can be approximated to arbitrary
accuracy with a mixture (linear sum) of two-dimensional Gaussians.
Mixtures of Gaussians are fast to render, fast to affine-transform,
and fast to convolve with mixture-of-Gaussian PSF models, all at
machine precision.  We present worked examples that can be directly
used in image fitting; we are using them ourselves.  We also advocate
modeling PSFs also as arbitrary mixtures of Gaussians.  Amusingly, in
the optically thin limit, a mixture-of-Gaussian two-dimensional model
directly implies its own three-dimensional de-projection.
\end{abstract}

...Some other authors have worked with mixtures of Gaussians...Very
related work...thinking more generally...Here we want to solve a very
specific set of numerical problems...

...Mixtures of Gaussians sometimes used for PSF fitting...Used in
XD...Other places?...

...Rendering of Sersic profiles (of which exp and dev are types) is
non-trivial at the center...for the same reason convolution is a
bitch...

...If we wanted to approximate these models, what objective function
would we use?  What tolerances are okay?...

...What do we get by running the relevant approximations for exp and
dev?...Can we make a continuous approximation for the Sersic
profiles?...

...How to use these in a real setting...An example from the Tractor...

\end{document}

### 2012-04-19 docs/mixture_models.tex
loaded in a whole lot o math

@@
 \documentclass[12pt]{article}
+\usepackage{amssymb,amsmath,mathrsfs}
+
+\newcommand{\tmatrix}[1]{\boldsymbol{#1}}
+\newcommand{\inverse}[1]{{#1}^{-1}}
+\newcommand{\transpose}[1]{{#1}^{\mathsf T}}
+\newcommand{\tvector}[1]{\boldsymbol{#1}}
+\newcommand{\pos}{\tvector{x}}
+\newcommand{\spos}{\tvector{\xi}}
+\newcommand{\mean}{\tvector{m}}
+\newcommand{\var}{\tmatrix{V}\!}
+\newcommand{\Gm}{\tmatrix{G}}
+\newcommand{\Hm}{\tmatrix{H}}
+\newcommand{\Rm}{\tmatrix{R}}
+\newcommand{\uv}{\tvector{u}}
+\newcommand{\zero}{\tmatrix{0}}
+\newcommand{\identity}{\tmatrix{I}}
+\newcommand{\normal}{N}
+\newcommand{\given}{\,|\,}
+\renewcommand{\star}{\mathrm{star}}
+\newcommand{\dev}{\mathrm{dev}}
 
@@
 modeling PSFs also as arbitrary mixtures of Gaussians.  Amusingly, in
-the optically thin limit, a mixture-of-Gaussian two-dimensional model
-directly implies its own three-dimensional de-projection.
+the optically thin limit, a circularly symmetric mixture-of-Gaussian
+two-dimensional model directly implies its own spherically symmetric
+three-dimensional de-projection.
 \end{abstract}
@@
 
+\begin{eqnarray}\displaystyle
+\normal(\pos\given\mean,\var) &\equiv& \frac{1}{2\pi}\,\det(\var)^{-1/2}\,\exp(-\frac{1}{2}\,\transpose{[\pos-\mean]}\cdot\inverse{\var}\cdot[\pos-\mean])
+\quad ,
+\end{eqnarray}
+
+\begin{eqnarray}\displaystyle
+Q_{\exp}(\spos) &=& \sum_{m=1}^{M_g} g_m\,\normal(\spos\given\zero,\Gm_m)
+\\
+1 &=& \sum_{m=1}^{M_g} g_m
+\\
+\Gm_m &=& \sigma^2_{\exp,m}\,\identity
+\quad ,
+\end{eqnarray}
+
+\begin{eqnarray}\displaystyle
+Q_{\dev}(\spos) &=& \sum_{m=1}^{M_h} h_m\,\normal(\spos\given\zero,\Hm_m)
+\\
+1 &=& \sum_{m=1}^{M_h} h_m
+\\
+\Hm_m &=& \sigma^2_{\dev,m}\,\identity
+\quad ,
+\end{eqnarray}
+
+\begin{eqnarray}\displaystyle
+\psi(\Delta\pos) &=& \sum_{k=1}^K p_k\,\normal(\Delta\pos\given\mean_k,\var_k)
+\\
+1 &=& \sum_{k=1}^K p_k
+\\
+I(\pos\given\star,S_s,\pos_s) &=& \sum_{k=1}^K S_s\,p_k\,\normal(\pos\given\pos_s+\mean_k,\var_k)
+\quad ,
+\end{eqnarray}
+
+\begin{eqnarray}\displaystyle
+I(\pos\given\exp,S_g,\pos_g,\var_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_g} S_g\,g_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gn}+\var_k)
+\\
+\var_{gn} &\equiv& \Rm_g\cdot\Gm_m\cdot\transpose{\Rm_g}
+\\
+\var_g &\equiv& \Rm_g\cdot\transpose{\Rm_g}
+\\
+\Rm_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
+\quad ,
+\end{eqnarray}
+where $a$ and $b$ are the major and minor axis lengths of the galaxy
+ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
+eigenvectors on the sky pointing in the major-axis and minor-axis
+directions respectively.  Since implicitly all vectors are
+two-dimensional column vectors, this makes $\Rm_g$ a $2\times 2$
+affine transformation matrix.
+
+\begin{eqnarray}\displaystyle
+I(\pos\given\dev,S_h,\pos_h,\var_h) &=& \sum_{k=1}^K \sum_{m=1}^{M_h} S_h\,h_m\,p_k\,\normal(\pos\given\pos_h+\mean_k,\var_{hn}+\var_k)
+\\
+\var_{hn} &\equiv& \Rm_h\cdot\Hm_m\cdot\transpose{\Rm_h}
+\\
+\var_h &\equiv& \Rm_h\cdot\transpose{\Rm_h}
+\quad ,
+\end{eqnarray}
+where...
+
 \end{document}

### 2012-04-19 docs/mixture_models.tex
more words

@@
 \usepackage{amssymb,amsmath,mathrsfs}
+
+\newcommand{\documentname}{\textsl{Note}}
+\newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\tractor}{\project{Tractor}}
 
@@
 \newcommand{\Hm}{\tmatrix{H}}
-\newcommand{\Rm}{\tmatrix{R}}
+\newcommand{\affine}{\tmatrix{R}}
 \newcommand{\uv}{\tvector{u}}
@@
 \begin{abstract}
-Exponential, de~Vaucouleurs, and Sersic profiles are simple and
+Exponential and de~Vaucouleurs profiles are simple and
 successful models for fitting two-dimensional images of galaxies.  One
@@
 
-...What do we get by running the relevant approximations for exp and
-dev?...Can we make a continuous approximation for the Sersic
-profiles?...
+...Limit scope of this document to just what we need for the
+\tractor...
 
-...How to use these in a real setting...An example from the Tractor...
-
+In what follows we will exclusively use two-dimensional Gaussian or
+Normal distributions, which look like
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
+where $\pos$ and $\mean$ are two-dimensional vectors (usually in the
+focal plane or on the sky or something like that), and $\var$ is a
+symmetric $2\times 2$ variance tensor or matrix, and implicitly the
+vectors are column vectors.  A \emph{mixture of Gaussians} is a linear
+superposition of Gaussians.  Any positive two-dimensional function
+with finite support and finite total integral---including as a special
+case any two-dimensional probability distribution function---can be
+represented as a mixture of Gaussians to arbitrary accuracy (DSTN: It
+would be nice to have a reference here).
 
+We wish to make an approximation to the two-dimensional circular
+exponential profile
+\begin{eqnarray}\displaystyle
+Q_{\exp}(\spos) &\approx& \frac{1}{Z_{\exp}}\,\exp(-\frac{|\spos|}{r_{\exp}})
+\\
+Z_{\exp} &=& XXXXX
+\\
+r_{\exp} &=& XXXXX
+\quad ,
+\end{eqnarray}
+where $\spos$ is a dimensionless focal-plane position, and $Z_{\exp}$
+and $r_{\exp}$ are parameters that ensure that the profile integrates
+to unit flux and has unit half-light radius.  The position $\spos$ is
+dimensionless because it parameterizes the unit-size and unit-flux
+dimensionless function.  We seek the best (where ``best'' will be
+defined below) $M_g$-Gaussian mixture approximation
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-
+where all of the means are exactly zero and all of the variances
+$\Gm_m$ in the mixture are scalar multiples of the identity matrix
+$\identity$ because we are requiring this dimensionless function to be
+precisely circular (so every component is itself circular and
+concentric).  Similarly for the de~Vaucouleurs profile
 \begin{eqnarray}\displaystyle
+Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\left[\frac{|\spos|}{r_{\dev}}\right]^{1/4})
+\\
+Z_{\dev} &=& XXXXX
+\\
+r_{\dev} &=& XXXXX
+\\
 Q_{\dev}(\spos) &=& \sum_{m=1}^{M_h} h_m\,\normal(\spos\given\zero,\Hm_m)
@@
 \Hm_m &=& \sigma^2_{\dev,m}\,\identity
-\quad ,
+\quad .
 \end{eqnarray}
 
+...Discuss the cutoffs used by PHOTO and adjust $Z$ and $r$ accordingly...
+
+...Define a sensible scalar objective...Give fitting procedure...give actual values for the parameters...
+
+...Show plots demonstrating quality of fit, one-d and two-d plots...
+
+The value of these mixture-of-Gaussian galaxy components comes when
+they are to be convolved with a PSF (usually in fact a pixel-convolved
+PSF) that is itself represented as a mixture of Gaussians.  In this
+scenario, the PSF $\psi(\Delta\pos)$---which is thought of as a
+function of focal-plane displacement $\Delta\pos$ away from, say, a
+true stellar position---is represented as a $K$-Gaussian mixture
 \begin{eqnarray}\displaystyle
@@
 1 &=& \sum_{k=1}^K p_k
-\\
-I(\pos\given\star,S_s,\pos_s) &=& \sum_{k=1}^K S_s\,p_k\,\normal(\pos\given\pos_s+\mean_k,\var_k)
 \quad ,
 \end{eqnarray}
+where the means $\mean_k$ are \emph{not} required to vanish because
+the PSF can have arbitrarily non-trivial structure (think speckles and
+the like) and the variances $\var_k$ will not in general be
+proportional to the identity or even diagonal because the PSF will not
+in general be round.  An example that illustrates the use of this PSF
+is the following: It implies that a star of flux $S_s$ at focal-plane
+position $\pos_s$ will lead to an image (PSF-convolved intensity map)
+of the form
+\begin{eqnarray}\displaystyle
+I(\pos\given\star,S_s,\pos_s) &=& \sum_{k=1}^K S_s\,p_k\,\normal(\pos\given\pos_s+\mean_k,\var_k)
+\quad .
+\end{eqnarray}
+That is, when the PSF is represented as a mixture of Gaussians, any
+image of a star---or indeed any image of any set of stars---is also
+represented as a mixture of Gaussians.
 
+Applying this PSF to an exponential or de~Vaucouleurs galaxy is
+slightly more complicated, because the galaxy has not just a flux
+$S_g$ and a central position $\pos_g$; it also has a shape.  Because
+we are only considering these simple galaxies, we are only permitting
+ellipsoidal shapes, which can be represented by a semi-major axis $a$,
+a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
+eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or by a symmetric
+variance tensor $\var_g$ (which is a general representation of an
+ellipse).  The galaxy is distorted by this variance tensor
+\emph{prior} to PSF convolution, so the focal-plane image
+(PSF-convolved intensity field) for a general exponential galaxy is
+given by
 \begin{eqnarray}\displaystyle
@@
 \\
-\var_{gn} &\equiv& \Rm_g\cdot\Gm_m\cdot\transpose{\Rm_g}
+\var_{gn} &\equiv& \affine_g\cdot\Gm_m\cdot\transpose{\affine_g}
 \\
-\var_g &\equiv& \Rm_g\cdot\transpose{\Rm_g}
+\var_g &\equiv& \affine_g\cdot\transpose{\affine_g}
 \\
-\Rm_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
+\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
 \quad ,
@@
 directions respectively.  Since implicitly all vectors are
-two-dimensional column vectors, this makes $\Rm_g$ a $2\times 2$
-affine transformation matrix.
-
+two-dimensional column vectors, this makes $\affine_g$ a $2\times 2$
+affine transformation matrix and $\var_g$ a symmetric $2\times 2$
+variance tensor.  Similarly, the general de~Vaucouleurs galaxy is
+given by
 \begin{eqnarray}\displaystyle
@@
 \\
-\var_{hn} &\equiv& \Rm_h\cdot\Hm_m\cdot\transpose{\Rm_h}
+\var_{hn} &\equiv& \affine_h\cdot\Hm_m\cdot\transpose{\affine_h}
 \\
-\var_h &\equiv& \Rm_h\cdot\transpose{\Rm_h}
+\var_h &\equiv& \affine_h\cdot\transpose{\affine_h}
 \quad ,
 \end{eqnarray}
-where...
+where the matrix mathematics is the same as that in the exponential
+case.  Note the important and key result of this \documentname, to
+wit, that a mixture-of-Gaussian galaxy model (with $M$ components)
+convolved with a mixture-of-Gaussian PSF model (with $K$ components)
+yields a mixture-of-Gaussian image (with $[M\,K]$ components).
+
+...pegagogical figure demonstrating these operations and their results...
+
+...what would you want to do to generalize to Sersic profiles?...
+
+...what about mixtures of exponentials and de~Vaucouleurs:  No problem!...
 

### 2012-04-19 docs/mixture_models.tex
abstract and text tweaks

@@
-% Copyright 2012 David W. Hogg (NYU).
+% This file is part of the Tractor project.
+% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton).
 % All rights reserved.
@@
 \begin{abstract}
-Exponential and de~Vaucouleurs profiles are simple and
-successful models for fitting two-dimensional images of galaxies.  One
-numerical issue encountered in this kind of fitting is the pixel
-rendering and convolution (or correlation) of the models with the
-telescope point-spread function (PSF); these operations are slow, and
-easy to get slightly wrong at small radii.  Here we exploit the
-realization that these models can be approximated to arbitrary
-accuracy with a mixture (linear sum) of two-dimensional Gaussians.
-Mixtures of Gaussians are fast to render, fast to affine-transform,
-and fast to convolve with mixture-of-Gaussian PSF models, all at
-machine precision.  We present worked examples that can be directly
-used in image fitting; we are using them ourselves.  We also advocate
-modeling PSFs also as arbitrary mixtures of Gaussians.  Amusingly, in
-the optically thin limit, a circularly symmetric mixture-of-Gaussian
-two-dimensional model directly implies its own spherically symmetric
-three-dimensional de-projection.
+Exponential and de~Vaucouleurs profiles are simple and successful
+models for fitting two-dimensional images of galaxies.  One numerical
+issue encountered in this kind of fitting is the pixel rendering and
+convolution (or correlation) of the models with the telescope
+point-spread function (PSF); these operations are slow, and easy to
+get slightly wrong at small radii.  Here we exploit the realization
+that these models can be approximated to arbitrary accuracy with a
+mixture (linear superposition) of two-dimensional Gaussians.  Mixtures
+of Gaussians are fast to render and fast to affine-transform.  Most
+importantly, if you have a mixture-of-Gaussian PSF model (which we
+advocate), the PSF-convolved, affine-transformed galaxy models are
+themselves mixtures of Gaussians and therefore very fast to compute
+precisely.  We present worked examples that can be directly used in
+image fitting; we are using them ourselves.  We discuss extensions to
+Sersic models and applications of these ideas to three-dimensional
+deprojection.
 \end{abstract}
@@
 
-In what follows we will exclusively use two-dimensional Gaussian or
+...We need to put somewhere Lupton-like comments about sampling the
+pixel-convolved-PSF-convolved scene...
+
+In what follows we will make heavy use of two-dimensional Gaussian or
 Normal distributions, which look like
@@
 
-We wish to make an approximation to the two-dimensional circular
-exponential profile
+We wish to make an approximation $Q_{\exp}(\cdot)$ to the
+two-dimensional circular exponential profile
 \begin{eqnarray}\displaystyle
@@
 \\
-Z_{\exp} &=& XXXXX
+Z_{\exp} &=& X.XXXX
 \\
-r_{\exp} &=& XXXXX
+r_{\exp} &=& X.XXXX
 \quad ,
@@
 \\
-Z_{\dev} &=& XXXXX
+Z_{\dev} &=& X.XXXX
 \\
-r_{\dev} &=& XXXXX
+r_{\dev} &=& X.XXXX
 \\

### 2012-05-08 docs/mixture_models.tex
to-do list
structure like a paper, intro text, reference list started

@@
 
-\documentclass[12pt]{article}
+% to-do
+% -----
+% - zeroth draft
+% - make useful figures
+% - send to Jim Bosch and Kevin Bundy for comments
+
+\documentclass[12pt,pdftex,preprint]{aastex}
 \usepackage{amssymb,amsmath,mathrsfs}
 
+\newcommand{\foreign}[1]{\textit{#1}}
+\newcommand{\etal}{et\,al.}
 \newcommand{\documentname}{\textsl{Note}}
@@
 \newcommand{\tractor}{\project{Tractor}}
+\newcommand{\sdss}{\project{SDSS}}
 
@@
 
-\section*{Replacing standard galaxy profiles with \\ mixtures of Gaussians}
-
-\noindent
-David W. Hogg \\
-\textsl{New York University} \\
-\textsl{Max-Planck-Institut f\"ur Astronomie} \\[1ex]
-Dustin Lang \\
-\textsl{Princeton University Observatory}
+\title{Replacing standard galaxy profiles with \\ mixtures of Gaussians}
+\author{David W. Hogg\altaffilmark{1,2,3} \&
+        Dustin Lang\altaffilmark{4}}
+\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
+\altaffiltext{2}{New York University}
+\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
+\altaffiltext{4}{Princeton University Observatory}
 
@@
 
-...Some other authors have worked with mixtures of Gaussians...Very
-related work...thinking more generally...Here we want to solve a very
-specific set of numerical problems...
+Gaussians are remarkable distribution functions.  They have the
+incredible properties that---in any number of dimensions---the
+convolution (or correlation) of one multivariate Gaussian with another
+is itself a multivariate Gaussian, and any product of multivariate
+Gaussians is itself a multivariate Gaussian but with a different
+normalization.  Furthermore, the means and variance tensors of the
+Gaussians output by these operations are related simply to the means
+and variance tensors of the inputs.  Add to these wonders the fact
+that Gaussians form a complete basis for representing (smooth)
+probability distribution functions and it becomes remarkable that we
+don't do everything we do in terms of Gaussians.
 
-...Mixtures of Gaussians sometimes used for PSF fitting...Used in
-XD...Other places?...
+To elaborate, a mixture of multivariate Gaussians---a linear
+superposition---can be used to represent any reasonable distribution
+in any number of dimensions to any reasonable precision.  Convolution
+(or correlation) by any other distribution that has also been
+represented by a mixture of multivariate Gaussians creates a new
+mixture of Gaussians with simply adjusted amplitudes, means, and
+variance tensors.  The ubiquity of convolution operations in astronomy
+suggests the widespread adoption of mixture-of-Gaussian modeling.  We
+have pioneered this in the area of distribution modeling in large
+numbers of dimensions \citep{xd, xdqso, xdqsoz}, where convolution
+occurs because the true or noise-free distribution is convolved with
+the noise before being observed.  Here we are going to capitalize on
+the convolution properties of mixtures of Gaussians in modeling galaxy
+morphologies in imaging data, where the true or
+high-angular-resolution intensity field is convolved with the
+point-spread function (PSF) before being observed.  This is
+particularly useful in data sets in which the PSF itself has also been
+modeled as a mixture of Gaussians, which is not uncommon.  For
+example, the \project{Sloan Digital Sky Survey} (\sdss; [HOGG cite
+  York]) imaging pipelines [HOGG cite Lupton] makes such models for
+every imaging field.
 
-...Rendering of Sersic profiles (of which exp and dev are types) is
-non-trivial at the center...for the same reason convolution is a
-bitch...
+We are not the first in this space....modeling galaxies with mixtures
+of Gaussians has been done very successfully before [HOGG cite].
+However, the idea behind that work was to use the mixture of Gaussians
+to provide a very free form for the isophotes of large, nearby, highly
+resolved galaxies.  Here our goals are very limited: We want to
+improve the performance of standard galaxy image model fitting by
+expressing the standard galaxy models---the exponential and
+de~Vaucouleurs profiles---as rigid mixtures of Gaussians.
 
-...If we wanted to approximate these models, what objective function
-would we use?  What tolerances are okay?...
+The performance advantages we will obtain from using mixtures of
+Gaussians are not simply that convolution itself is trivial.  The
+standard unconvolved galaxy models, especially the de~Vaucouleurs and
+S\'ersic models, are very ill-behaved near the galaxy center.
+Rendering these profiles precisely near the center can be very
+challenging numerically.  In addition, most of the rendering time is
+spent at very small radii, where PSF convolution is going to erase all
+structure anyway.  That is, a mixture-of-Gaussians description of the
+de~Vaucouleurs profile saves time both in rendering \emph{and} in PSF
+convolution; it produces profiles that are approximate but very high
+performance when all real uses of the profiles are PSF-convolved, as
+they usually are.
 
@@
 
-...Define a sensible scalar objective...Give fitting procedure...give actual values for the parameters...
+...If we wanted to approximate these models, what objective function
+would we use?  What tolerances are okay?...Define a sensible scalar
+objective...Give fitting procedure...give actual values for the
+parameters...
 
@@
 
+\acknowledgements
+It is a pleasure to thank...
+
+\begin{thebibliography}{70}
+\bibitem[Bovy \etal(2011a)]{xd}
+Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
+\bibitem[Bovy \etal(2011b)]{xdqso}
+Bovy,~J., \etal, 2011, \apj, 729, 141
+\bibitem[Bovy \etal(2012)]{xdqsoz}
+Bovy,~J. \etal, 2012, \aj, 749, 41
+\end{thebibliography}
+
 \end{document}

### 2012-05-08 docs/mixture_models.tex
galfit
references

@@
 % -----
-% - zeroth draft
+% - write zeroth draft
 % - make useful figures
@@
 \newcommand{\foreign}[1]{\textit{#1}}
-\newcommand{\etal}{et\,al.}
+\newcommand{\etal}{\foreign{et\,al.}}
 \newcommand{\documentname}{\textsl{Note}}
@@
 
-We are not the first in this space....modeling galaxies with mixtures
-of Gaussians has been done very successfully before [HOGG cite].
-However, the idea behind that work was to use the mixture of Gaussians
-to provide a very free form for the isophotes of large, nearby, highly
-resolved galaxies.  Here our goals are very limited: We want to
-improve the performance of standard galaxy image model fitting by
-expressing the standard galaxy models---the exponential and
-de~Vaucouleurs profiles---as rigid mixtures of Gaussians.
+We are not the first in this space deconvolution and modeling of
+galaxy images with mixtures of Gaussians has been done very
+successfully before \citep[for example,][]{bendinelli, emsellem,
+  bendinelli2, cappellari}.  However, the idea behind those projects
+was to use the mixture of Gaussians to provide a very free form for
+the isophotes or morphologies of resolved galaxies or other complex
+scenes.  Here our goals are very limited: We want to improve the
+performance of standard galaxy image model fitting by expressing the
+standard galaxy models---the exponential and de~Vaucouleurs
+profiles---as rigid mixtures of Gaussians.
 
@@
 \tractor...
+
+...somehow have to cite GALFIT \citep{galfit} in this section...
 
@@
 \begin{thebibliography}{70}
+\bibitem[Bendinelli(1991)]{bendinelli}
+Bendinelli,~O., 1991, \apj, 366, 599
+\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
+Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
 \bibitem[Bovy \etal(2011a)]{xd}
@@
 Bovy,~J. \etal, 2012, \aj, 749, 41
+\bibitem[Cappellari(2002)]{cappellari}
+Cappellari,~M., 2002, \mnras, 333, 400
+\bibitem[Emsellem \etal(1994)]{emsellem}
+Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
+\bibitem[Peng \etal(2002)]{galfit}
+Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
 \end{thebibliography}

### 2012-05-09 __init__.py
dropped everything but the optimize mixture code

@@ -1,30 +1,0 @@
-from engine import *
-from ducks import *
-from basics import *
-import sdss
-import cfht
-from nasasloan import *
-
-from ttime import Time
-__all__ = [
-	# modules
-	'sdss', 'fitpsf', 'emfit', 'sdss_galaxy',
-	# ducks
-	'Params', 'Sky', 'Source', 'Position', 'Brightness', 'PhotoCal',
-	'PSF', 
-	# utils
-	'BaseParams', 'ScalarParam', 'ParamList', 'MultiParams',
-	'NamedParams',
-	# basics
-	'ConstantSky', 'PointSource',
-	'Flux', 'Mag', 'Mags',
-	'PixPos', 'RaDecPos',
-	'NullPhotoCal', 'WCS', 'NullWCS',
-	'FitsWcs', 'GaussianMixturePSF',
-	'NCircularGaussianPSF', 
-	# engine
-	'Patch', 'Image', 'Images',
-	'Catalog', 'Tractor',
-	# ttime
-	'Time',
-	]

### 2012-05-09 basics.py

@@ -1,584 +1,0 @@
-"""
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`basics.py`
-===========
-
-Generally useful generic implementations of things the Tractor needs:
-Magnitudes, (RA,Dec) positions, FITS WCS, and so on.
-
-"""
-from math import ceil, floor, pi, sqrt, exp
-
-from engine import *
-#from ducks import *
-from utils import *
-import mixture_profiles as mp
-
-
-class Mags(ParamList):
-	'''
-	An implementation of `Brightness` that stores magnitudes in
-	multiple bands.
-	'''
-	def __init__(self, **kwargs):
-		'''
-		Mags(r=14.3, g=15.6, order=['r','g'])
-
-		The `order` parameter is optional; it determines the ordering
-		of the bands in the parameter vector (eg, `getParams()`).
-		'''
-		keys = kwargs.pop('order', None)
-		if keys is None:
-			keys = kwargs.keys()
-			keys.sort()
-		assert(len(kwargs) == len(keys))
-		assert(set(kwargs.keys()) == set(keys))
-		vals = []
-		for k in keys:
-			vals.append(kwargs[k])
-		super(Mags,self).__init__(*vals)
-		self.order = keys
-		self.addNamedParams(**dict((k,i) for i,k in enumerate(keys)))
-
-	def getMag(self, bandname):
-		''' Bandname: string
-		Returns: mag in the given band.
-		'''
-		return getattr(self, bandname)
-
-	def __setstate__(self, state):
-		'''For pickling.'''
-		self.__dict__ = state
-		self.addNamedParams(**dict((k,i) for i,k in enumerate(self.order)))
-
-
-class Mag(ScalarParam):
-	'''
-	An implementation of `Brightness` that stores a single magnitude.
-	'''
-	stepsize = 0.01
-	strformat = '%.3f'
-
-class Flux(ScalarParam):
-	'''
-	A `Brightness` implementation that stores raw counts.
-	'''
-	def __mul__(self, factor):
-		new = self.copy()
-		new.val *= factor
-		return new
-	__rmul__ = __mul__
[…]

### 2012-05-09 cache.py

@@ -1,130 +1,0 @@
-try:
-	# python 2.7
-	from collections import OrderedDict
-except:
-	#from .ordereddict import OrderedDict
-	from ordereddict import OrderedDict
-
-
-'''
-LRU cache.
-This code is based on: http://code.activestate.com/recipes/498245-lru-and-lfu-cache-decorators/
-By: Raymond Hettinger
-License: Python Software Foundation (PSF) license.
-'''
-class Cache(object):
-	class Entry(object):
-		pass
-	def __init__(self, maxsize=1000, sizeattr='size'):
-		self.clear()
-		self.maxsize = maxsize
-		self.sizeattr = sizeattr
-
-	def clear(self):
-		self.dict = OrderedDict()
-		self.hits = 0
-		self.misses = 0
-		
-	def __setitem__(self, key, val):
-		sz = 0
-		if hasattr(val, self.sizeattr):
-			try:
-				sz = int(getattr(val, self.sizeattr))
-			except:
-				pass
-		e = Cache.Entry()
-		e.val = val
-		e.size = sz
-		e.hits = 0
-		# purge LRU item
-		if len(self.dict) >= self.maxsize:
-			self.dict.popitem(0)
-		self.dict[key] = e
-
-	def __getitem__(self, key):
-		# pop
-		try:
-			e = self.dict.pop(key)
-		except KeyError:
-			self.misses += 1
-			raise
-		self.hits += 1
-		# reinsert (to record recent use)
-		self.dict[key] = e
-		if e is None:
-			return e
-		e.hits += 1
-		return e.val
-	def __len__(self):
-		return len(self.dict)
-	def put(self, k, v):
-		self[k] = v
-	def get(self, *args):
-		if len(args) == 1:
-			key = args[0]
-			return self.__getitem__(key)
-		assert(len(args) == 2)
-		key,default = args
-		try:
-			return self.__getitem__(key)
-		except:
-			return default
-	def about(self):
-		print 'Cache has', len(self), 'items:'
-		for k,v in self.dict.items():
-			if v is None:
-				continue
-			print '  size', v.size, 'hits', v.hits
-	def __str__(self):
-		s =  'Cache: %i items, total of %i hits, %i misses' % (len(self), self.hits, self.misses)
-		nnone = 0
-		hits = 0
-		size = 0
-		for k,v in self.dict.items():
-			if v is None:
-				nnone += 1
-				continue
-			if v.val is None:
-				nnone += 1
-				continue
-			hits += v.hits
-			size += v.size
[…]

### 2012-05-09 cfht.py

@@ -1,40 +1,0 @@
-from .utils import BaseParams
-
-class CfhtPhotoCal(BaseParams):
-	def __init__(self, hdr=None, bandname=None):
-		self.bandname = bandname
-		if hdr is not None:
-			self.exptime = hdr['EXPTIME']
-			self.phot_c = hdr['PHOT_C']
-			self.phot_k = hdr['PHOT_K']
-			self.airmass = hdr['AIRMASS']
-			print 'CFHT photometry:', self.exptime, self.phot_c, self.phot_k, self.airmass
-		# FIXME -- NO COLOR TERMS (phot_x)!
-		'''
-		COMMENT   Formula for Photometry, based on keywords given in this header:
-		COMMENT   m = -2.5*log(DN) + 2.5*log(EXPTIME)
-		COMMENT   M = m + PHOT_C + PHOT_K*(AIRMASS - 1) + PHOT_X*(PHOT_C1 - PHOT_C2)
-		'''
-	def hashkey(self):
-		return ('CfhtPhotoCal', self.exptime, self.phot_c, self.phot_k, self.airmass)
-
-	def getParams(self):
-		return [self.phot_c,]
-	def getStepSizes(self):
-		return [0.01]
-	def setParam(self, i, p):
-		assert(i == 0)
-		self.phot_c = p
-
-	def getParamNames(self):
-		return ['phot_c']
-
-	def brightnessToCounts(self, brightness):
-		M = brightness.getMag(self.bandname)
-		logc = (M - self.phot_c - self.phot_k * (self.airmass - 1.)) / -2.5
-		return self.exptime * 10.**logc
-
-	#def countsToBrightness(self, counts):
-	#	return Mag(-2.5 * np.log10(counts / self.exptime) +
-	#			   self.phot_c + self.phot_k * (self.airmass - 1.))
-		

### 2012-05-09 ducks.py

@@ -1,218 +1,0 @@
-"""
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`ducks.py`
-===========
-
-Duck-type definitions of types used by the Tractor.
-
-Most of this code is not actually used at all.  It's here for
-documentation purposes.
-
-"""
-
-class Params(object):
-	'''
-	A set of parameters that can be optimized by the Tractor.
-
-	This is a duck-type definition.
-	'''
-	def copy(self):
-		return None
-		
-	def hashkey(self):
-		'''
-		Returns a tuple containing the state of this `Params` object
-		for use as a cache key.
-
-		All elements must be hashable: see
-		http://docs.python.org/glossary.html#term-hashable
-		'''
-		return ()
-	def __hash__(self):
-		''' Params must be hashable. '''
-		return None
-	#def __eq__(self, other):
-
-	def getParamNames(self):
-		''' Returns a list of strings: the names of the parameters. '''
-		return []
-	def numberOfParams(self):
-		''' Returns the number of parameters (ie, number of scalar
-		values).'''
-		return len(self.getParams())
-	def getParams(self):
-		''' Returns a *copy* of the current parameter values as an
-		iterable (eg, list)'''
-		return []
-	def getStepSizes(self, *args, **kwargs):
-		''' Returns "reasonable" step sizes for the parameters.'''
-		return []
-	def setParams(self, p):
-		''' Sets the parameter values to the values in the given
-		iterable `p`.  The length of `p` will be equal to
-		`numberOfParams()`.'''
-		pass
-	def setParam(self, i, p):
-		'''
-		Sets parameter index 'i' to new value 'p'.
-
-		i: integer in the range [0, numberOfParams()).
-		p: float
-
-		Returns the old value.
-		'''
-		return None
-
-
-class Sky(Params):
-	'''
-	Duck-type definition for a sky model.
-	'''
-	def getParamDerivatives(self, img):
-		'''
-		Returns [ Patch, Patch, ... ], of length numberOfParams(),
-		containing the derivatives in the given `Image` for each
-		parameter.
-		'''
-		return []
-	def addTo(self, mod):
-		'''
-		Add the sky to the input synthetic image `mod`, a 2-D numpy
-		array.
-		'''
-		pass
-
-
-class Source(Params):
-	'''
-	This is the duck-type definition of a Source (star, galaxy, etc)
[…]

### 2012-05-09 engine.py

@@ -1,1458 +1,0 @@
-'''
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`engine.py`
-===========
-
-Core image modeling and fitting.
-'''
-
-from math import ceil, floor, pi, sqrt, exp
-import time
-import logging
-import random
-import os
-
-import numpy as np
-import pylab as plt
-
-from scipy.sparse import csr_matrix
-from scipy.sparse.linalg import lsqr
-
-from astrometry.util.miscutils import get_overlapping_region
-from astrometry.util.multiproc import *
-from .utils import MultiParams
-from .cache import *
-from .ttime import Time
-
-def logverb(*args):
-	msg = ' '.join([str(x) for x in args])
-	logging.debug(msg)
-def logmsg(*args):
-	msg = ' '.join([str(x) for x in args])
-	logging.info(msg)
-def isverbose():
-	# Ugly logging interface...
-	return (logging.getLogger().level <= logging.DEBUG)
-
-def set_fp_err():
-	'''Cause all floating-point errors to raise exceptions.
-	Returns the current error state so you can revert via:
-
-	    olderr = set_fp_err()
-	    # do stuff
-	    np.seterr(**olderr)
-	'''
-	return np.seterr(all='raise')
-
-class Image(MultiParams):
-	'''
-	An image plus its calibration information.  An `Image` has pixels,
-	inverse-variance map, WCS, PSF, photometric calibration
-	information, and sky level.  All these things are `Params`
-	instances, and `Image` is a `MultiParams` so that the Tractor can
-	optimize them.
-	'''
-	def __init__(self, **kwargs):
-		'''
-		Expected kwargs:
-		(data=None, invvar=None,
-		 psf=None, wcs=None, sky=None, photocal=None,
-		 name=None)
-		
-		'''
-		self.data = kwargs.pop('data', None)
-		self.invvar = kwargs.pop('invvar', None)
-		for i,x in enumerate(self.invvar):
-			for j,y in enumerate(x):
-				if y < 0:
-					self.invvar[i][j] = 0
-		self.inverr = np.sqrt(self.invvar)
-		self.origInvvar = self.invvar
-
[…]

### 2012-05-09 fitpsf.py

@@ -1,178 +1,0 @@
-#import matplotlib
-#matplotlib.use('Agg')
-import pylab as plt
-import numpy as np
-from math import pi, sqrt
-
-def makevector(w, mu, sig):
-	return np.append(w, np.append(mu.ravel(), sig.ravel()))
-
-def unpackvector(T, K):
-	return T[:K], T[K:K*3].reshape(K,2), T[K*3:].reshape(K,2,2)
-
-def em_update_theta(I, theta):
-	(K,N,II,xy) = I
-	(w, mu, sig) = unpackvector(theta, K)
-	w2,mu2,sig2 = em_update(K,N,II,xy,w,mu,sig)
-	return makevector(w2,mu2,sig2)
-
-def qn1fit(I, x0, y0, K=2, w=None, mu=None, sig=None):
-	(H,W) = I.shape
-	X,Y = np.meshgrid(np.arange(W)+x0, np.arange(H)+y0)
-	II = I.ravel()
-	II /= II.sum()
-	xy = np.array(zip(X.ravel(), Y.ravel())).astype(float)
-	N = len(II)
-
-	I = (K, N, II, xy)
-
-	def ghatfunc(theta):
-		tnew = em_update_theta(I, theta)
-		print 'ghat function:'
-		print '  theta = ', theta
-		print '  tnew  = ', tnew
-		return tnew - theta
-
-	# number of parameters
-	#    w, mu, sig
-	#P = (1 + 2 + 4) * K
-	w,mu,sig = em_init_params(K, w, mu, sig)
-	theta = makevector(w, mu, sig)
-	P = len(theta)
-
-	print 'w,mu,sig', w.shape, mu.shape, sig.shape
-	print 'theta', theta.shape
-
-	ghat = ghatfunc(theta)
-	#em_update_theta(I, theta) - theta
-	print 'ghat', ghat.shape
-	print ghat
-	A = -np.eye(P)
-	print 'A', A.shape
-
-	steps = 10
-	for step in range(steps):
-		deltheta = -np.dot(A, ghat)
-		print 'deltheta', deltheta.shape
-		print deltheta
-		delghat = ghatfunc(theta + deltheta) - ghat
-		print 'delghat', delghat.shape
-		print delghat
-
-		delA = (np.dot(np.dot((deltheta - np.dot(A, delghat)), deltheta.T), A)
[…]

### 2012-05-09 galaxy_profiles.py

@@ -1,73 +1,0 @@
-from numpy import exp, power, atleast_1d, isscalar
-
-### copied from photo: makeprof.c and phFitobj.h
-
-# r in half-light radii
-def profile_dev(r):
-	'''
-	>>> print round(profile_dev(1.), 3)
-	511.804
-
-	>>> print round(profile_dev(0.), 3)
-	61296.259
-
-	>>> print round(profile_dev(7.), 3)
-	4.191
-
-	>>> print round(profile_dev(8.), 3)
-	0.0
-	'''
-	DEVOUT = 8.0
-	DEVCUT = 7.*DEVOUT/8.
-
-	def dev_nc(r):
-		IE_DEV = 512.0
-		DEFAC = -7.66925
-		return IE_DEV * exp(DEFAC*(power(r*r + 0.0004, 0.125) - 1.0))
-
-	rr = atleast_1d(r)
-	p = dev_nc(rr)
-	p[rr > DEVCUT] *= (1 - ((rr[rr>DEVCUT] - DEVCUT)/(DEVOUT - DEVCUT))**2)**2
-	p[rr > DEVOUT] = 0.
-	if isscalar(r):
-		return p[0]
-	return p
-
-# r in half-light radii
-def profile_exp(r):
-	'''
-	>>> print round(profile_exp(1.), 3)
-	2048.0
-
-	>>> print round(profile_exp(0.), 3)
-	10970.542
-
-	>>> print round(profile_exp(3.), 3)
-	71.373
-
-	>>> print round(profile_exp(3.99), 3)
-	0.005
-
-	>>> print round(profile_exp(4.), 3)
-	0.0
-	'''
-
-	def exp_nc(r):
-		# Surface brightness at r_e
-		IE_EXP = 2048.0
-		EXPFAC = -1.67835
-		return IE_EXP * exp(EXPFAC*(r - 1.))
-
-	EXPOUT = 4.
-	EXPCUT = 3.*EXPOUT/4.
-	rr = atleast_1d(r)
-	p = exp_nc(rr)
-	p[rr > EXPCUT] *= (1 - ((rr[rr>EXPCUT] - EXPCUT)/(EXPOUT - EXPCUT))**2)**2
-	p[rr > EXPOUT] = 0.
-	if isscalar(r):
-		return p[0]
-	return p
-
-if __name__ == '__main__':
-	import doctest
-	doctest.testmod()

### 2012-05-09 hst.py

@@ -1,271 +1,0 @@
-# Copyright 2011 Dustin Lang and David W. Hogg.  All rights reserved.
-
-if __name__ == '__main__':
-	import matplotlib
-	matplotlib.use('Agg')
-import pyfits
-import pylab as plt
-
-from tractor import *
-
-import numpy as np
-from math import pi
-
-class HSTTractor(Tractor):
-
-	def createNewSource(self, img, x, y, ht):
-		# "ht" is the peak height (difference between image and model)
-		# convert to total flux by normalizing by my patch's peak pixel value.
-		#patch = img.getPsf().getPointSourcePatch(x, y)
-		#print 'psf patch:', patch.shape
-		#print 'psf patch: max', patch.max(), 'sum', patch.sum()
-		#print 'new source peak height:', ht, '-> flux', ht/patch.max()
-		#ht /= patch.max()
-		return PointSource(PixPos([x,y]), Flux(ht))
-
-def measure_sky_variance(img):
-	x0,y0 = 5,7
-	d = img[y0:,x0:] - img[:-y0,:-x0]
-	mad = abs(d.ravel())
-	I = np.argsort(mad)
-	mad = mad[I[len(I)/2]]
-	print 'median abs diff:', mad
-	sigmasq = mad**2 * pi / 4.
-	print 'sigma', np.sqrt(sigmasq)
-	return sigmasq
-
-
-def main():
-
-	if True:
-		P = pyfits.open('jbf108bzq_flt.fits')
-		img = P[1].data
-		err = P[2].data
-		#dq  = P[3].data
-
-		P = pyfits.open('jbf108bzq_dq1.fits')
-		dq  = P[0].data
-
-		#cut = [slice(900,1200), slice(2300,2600)]
-		cut = [slice(1400,1700), slice(2300,2600)]
-		img = img[cut]
-		err = err[cut]
-		dq  = dq[cut]
-
-		skyvar = measure_sky_variance(img)
-		skymed = np.median(img.ravel())
-		skysig = np.sqrt(skyvar)
-		print 'Estimate sky value', skymed, 'and sigma', skysig
-		zrange = np.array([-3.,+10.]) * skysig + skymed
-		invvar = 1. / (err**2)
-		invvar[dq > 0] = 0
-
-	else:
-		# Dealing with cosmic rays is a PITA so use DRZ for now...
-		P = pyfits.open('jbf108020_drz.fits')
[…]

### 2012-05-09 integral_image.py

@@ -1,65 +1,0 @@
-# Copyright 2011 Dustin Lang.  All rights reserved.
-
-from numpy import *
-
-def integral_image(I):
-	return cumsum(cumsum(I, axis=0), axis=1)
-
-def intimg_rect(img, xx0, xx1, yy0, yy1):
-	'''
-	Computes the area of the rectangular region in an integral image.
-
-	(xx0,yy0) are one less than the first pixel included in the sum.
-	(xx1,yy1) are the last pixel included in the sum.
-
-	(Note that the inclusiveness is opposite from usual.)
-
-	These can be (1-d) arrays, but they must be the same shape.
-
-	The return value is:
-
-        sum_{i=xx0 + 1}^{xx1} sum_{j = yy0 + 1}^{yy1} I[j,i]
-
-    where I is the original image
-
-	>>> if True:
-	...     from numpy.random import random_sample, seed
-	...     seed(42)
-	...     I = random_sample((10,10))
-	...     # This is the explicit sum
-	...     R0 = sum(I[3:5, 4:8])
-	...     print round(R0, 5)
-	...     # This is the integral-image sum
-	...     intimg = integral_image(I)
-	...     R1 = intimg_rect(intimg, 4-1, 8-1, 3-1, 5-1)
-	...     print round(R1, 5)
-	3.9294
-	3.9294
-	
-	'''
-	x0 = atleast_1d(xx0).astype(int)
-	x1 = atleast_1d(xx1).astype(int)
-	y0 = atleast_1d(yy0).astype(int)
-	y1 = atleast_1d(yy1).astype(int)
-	assert(x0.shape == x1.shape)
-	assert(x0.shape == y0.shape)
-	assert(x0.shape == y1.shape)
-	assert(all(x1 >= 0))
-	assert(all(y1 >= 0))
-	assert(all(x1 > x0))
-	assert(all(y1 > y0))
-	A = img[y1,x1]
-	I = logical_and((x0 >= 0), (y0 >= 0))
-	A[I] += img[y0[I],x0[I]]
-	I = (x0 >= 0)
-	A[I] -= img[y1[I],x0[I]]
-	I = (y0 >= 0)
-	A[I] -= img[y0[I],x1[I]]
-	if isscalar(xx0):
-		return A[0]
-	return A
-
-
-if __name__ == '__main__':
-	import doctest
-	doctest.testmod()

### 2012-05-09 mixture_profiles.py

@@ -1,292 +1,0 @@
-# Copyright 2011 David W. Hogg and Dustin Lang.  All rights reserved.
-
-if __name__ == '__main__':
-	import matplotlib
-	matplotlib.use('Agg')
-	import pylab as plt
-	import matplotlib.cm as cm
-import numpy as np
-import scipy.spatial.distance as scp
-
-# magic arrays, generated by running optimize_mixture_profiles.py:
-exp_amp = np.array([3.31636565e-05, 1.06478564e-03, 1.33260624e-02, 1.06217866e-01,
-		    6.09924868e-01, 2.43600369e+00, 5.34553250e+00, 3.41379672e+00])
-exp_var = np.array([4.91509189e-05, 7.91283025e-04, 5.06909854e-03, 2.30018199e-02,
-		    8.50523831e-02, 2.73398885e-01, 7.93675135e-01, 2.17065603e+00])
-dev_amp = np.array([1.36305372e-02, 1.08889599e-01, 3.68235229e-01, 9.26647361e-01,
-		    2.00002437e+00, 3.77384257e+00, 6.01053703e+00, 7.22968202e+00])
-dev_var = np.array([1.34654444e-04, 5.04128747e-04, 1.88518562e-03, 7.20439754e-03,
-		    2.87959626e-02, 1.25118904e-01, 6.38235086e-01, 4.76437813e+00])
-
-def get_exp_mixture():
-	return MixtureOfGaussians(exp_amp, np.zeros((exp_amp.size, 2)), exp_var)
-
-def get_dev_mixture():
-	return MixtureOfGaussians(dev_amp, np.zeros((dev_amp.size, 2)), dev_var)
-
-class MixtureOfGaussians():
-
-	# symmetrize is an unnecessary step in principle, but in practice?
-	def __init__(self, amp, mean, var):
-		self.amp = np.array(amp).astype(float)
-		self.mean = np.atleast_2d(np.array(mean)).astype(float)
-		(self.K, self.D) = self.mean.shape
-		self.set_var(var)
-		self.symmetrize()
-		#self.test()
-
-	def __str__(self):
-		result = "MixtureOfGaussians instance"
-		result += " with %d components in %d dimensions:\n" % (self.K, self.D)
-		result += " amp	 = %s\n" % self.amp.__str__()
[…]

### 2012-05-09 mpcache.py

@@ -1,47 +1,0 @@
-from multiprocessing import Manager
-from multiprocessing.managers import BaseManager
-#from .cache import Cache
-from cache import Cache
-
-class CacheManager(BaseManager):
-	pass
-CacheManager.register('Cache', Cache, exposed=('get','put','printStats'))
-
-def createManager():
-	manager = CacheManager()
-	manager.start()
-	return manager
-
-def createCache(**kwargs):
-	man = createManager()
-	#kwargs = {}
-	#if maxsize is not None:
-	#		kwargs['maxsize'] = maxsize
-	cache = man.Cache(**kwargs)
-	return cache
-
-
-
-def testProcess(cache):
-	import time
-	import os
-	for i in range(10):
-		time.sleep(1)
-		print os.getpid(), 'get', i, cache.get(i, None)
-		print os.getpid(), 'put', i
-		cache.put(i, i**3)
-
-if __name__ == '__main__':
-	from multiprocessing import Process
-
-	cache = createCache()
-	pp = []
-	for i in range(5):
-		p = Process(target=testProcess, args=(cache,))
-		p.start()
-		pp.append(p)
-
-	for p in pp:
-		p.join()
-
-	

### 2012-05-09 saveImg.py

@@ -1,171 +1,0 @@
-import os
-import logging
-import numpy as np
-import pylab as plt
-import matplotlib
-matplotlib.use('Agg')
-import pyfits
-
-from astrometry.util.file import *
-
-from tractor import *
-from tractor import sdss as st
-
-
-def save(idstr, tractor, zr,debug=False,plotAll=False,imgi=0):
-	print "Index: ", imgi
-	mod = tractor.getModelImages()[imgi]
-	chi = tractor.getChiImages()[imgi]
-
-	synthfn = 'synth-%s.fits' % idstr
-	print 'Writing synthetic image to', synthfn
-	pyfits.writeto(synthfn, mod, clobber=True)
-
-	pfn = 'tractor-%s.pickle' % idstr
-	print 'Saving state to', pfn
-	pickle_to_file(tractor, pfn)
-
-	timg = tractor.getImage(imgi)
-	data = timg.getImage()
-	ima = dict(interpolation='nearest', origin='lower',
-			   vmin=zr[0], vmax=zr[1])
-	imchi = ima.copy()
-	imchi.update(vmin=-10, vmax=10)
-	
-	if debug:
-		sources = tractor.getCatalog()
-		wcs = timg.getWcs()
-		allobjx = []
-		allobjy = []
-		allobjc = []
-		pointx = []
-		pointy = []
-		xplotx = []
-		xploty = []
-
-		for obj in sources:
-			if (isinstance(obj,PointSource)):
-				xt,yt = wcs.positionToPixel(obj.getPosition(), obj)
-				pointx.append(xt)
-				pointy.append(yt)
-				continue
-			shapes = []
-			attrType = []
-			if (isinstance(obj,st.CompositeGalaxy)):
-				for attr in 'shapeExp', 'shapeDev':
-					shapes.append(getattr(obj, attr))
-					attrType.append(attr)
-			else:
-				shapes.append(getattr(obj,'shape'))
-				attrType.append(' ')
-			x0,y0 = wcs.positionToPixel(obj.getPosition(), obj)
-			
-			cd = timg.getWcs().cdAtPixel(x0,y0)
-			for i,shape in enumerate(shapes):
-				xplotx.append(x0)
-				xploty.append(y0)
-				T=np.linalg.inv(shape.getTensor(cd))
-
-				x,y = [],[]
[…]

### 2012-05-09 sdss.py

@@ -1,1133 +1,0 @@
-"""
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`sdss.py`
-===========
-
-SDSS-specific implementation of Tractor elements:
-
- - retrieves and reads SDSS images, converting them to Tractor language
- - knows about SDSS photometric and astrometric calibration conventions
-
-"""
-import os
-from math import pi, sqrt, ceil, floor
-from datetime import datetime
-
-import pyfits
-import pylab as plt
-import numpy as np
-import matplotlib
-
-from engine import *
-from basics import *
-from sdss_galaxy import *
-
-from astrometry.sdss import * #DR7, band_name, band_index
-from astrometry.util.pyfits_utils import *
-from astrometry.util.file import *
-from astrometry.util.ngc2000 import ngc2000
-from astrometry.util.plotutils import setRadecAxes, redgreen
-
-# might want to override this to set the step size to ~ a pixel
-#class SdssRaDecPos(RaDecPos):
-#	def getStepSizes(self, img):
-#		return [1e-4, 1e-4]
-
-def _check_sdss_files(sdss, run, camcol, field, bandname, filetypes,
-					  retrieve=True):
-	bandnum = band_index(bandname)
-	for filetype in filetypes:
-		#fn = sdss.getFilename(filetype, run, camcol, field, bandname)
-		fn = sdss.getPath(filetype, run, camcol, field, bandname)
-		print 'Looking for file', fn
-		if not os.path.exists(fn):
-			if retrieve:
-				print 'Retrieving', fn
-				res = sdss.retrieve(filetype, run, camcol, field, bandnum)
-				if res is False:
-					raise RuntimeError('No such file on SDSS DAS: %s, rcfb %i/%i/%i/%s' % (filetype, run, camcol, field, bandname))
-			else:
-				raise os.OSError('no such file: "%s"' % fn)
-
-def _getBrightness(counts, tsf, bands):
-	allcounts = counts
-	order = []
[…]

### 2012-05-09 sdss_galaxy.py

@@ -1,501 +1,0 @@
-"""
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`sdss_galaxy.py`
-===========
-
-SDSS exponential and deVaucouleurs galaxy model classes.
-
-These models are not specific to SDSS *images*, they just use a
-slightly modified definition of the exp and dev profiles from SDSS
-Photo.
-"""
-import numpy as np
-
-import mixture_profiles as mp
-from engine import *
-from utils import *
-
-from cache import *
-
-_galcache = Cache(maxsize=10000)
-def get_galaxy_cache():
-	return _galcache
-
-class GalaxyShape(ParamList):
-	@staticmethod
-	def getNamedParams():
-		# re: arcsec
-		# ab: axis ratio, dimensionless, in [0,1]
-		# phi: deg, "E of N", 0=direction of increasing Dec, 90=direction of increasing RA
-		return dict(re=0, ab=1, phi=2)
-	def __repr__(self):
-		return 're=%g, ab=%g, phi=%g' % (self.re, self.ab, self.phi)
-	def __str__(self):
-		return 're=%.2f, ab=%.2f, phi=%.1f' % (self.re, self.ab, self.phi)
-	def copy(self):
-		return GalaxyShape(*self.vals)
-	def getParamNames(self):
-		return ['re', 'ab', 'phi']
-
-	def getStepSizes(self):
-		abstep = 0.01
-		if self.ab >= (1 - abstep):
-			abstep = -abstep
-		return [ 1., abstep, 1. ]
-
-	def setre(self, re):
-		if re < (1./30.):
-			#print 'Clamping re from', re, 'to 1/30'
-			pass
-		self.re = max(1./30., re)
-	def setab(self, ab):
-		if ab > 1.:
-			#print 'Converting ab from', ab, 'to', 1./ab
-			self.setab(1./ab)
-			self.setphi(self.phi+90.)
-		elif ab < (1./30.):
-			#print 'Clamping ab from', ab, 'to 1/30'
-			self.ab = 1./30
-		else:
-			self.ab = ab
-	def setphi(self, phi):
-		# limit phi to [-180,180]
-		self.phi = np.fmod(phi, 360.)
-		if self.phi < -180.:
-			self.phi += 360.
-		if self.phi > 180.:
-			self.phi -= 360.
-
-	# Note about flipping the galaxy when ab>1:
-	#
[…]

### 2012-05-09 setup-emfit.py

@@ -1,23 +1,0 @@
-from distutils.core import setup, Extension
-from numpy.distutils.misc_util import get_numpy_include_dirs
-
-numpy_inc = get_numpy_include_dirs()
-
-c_swig_module = Extension('_emfit',
-						  sources = ['emfit_wrap.c' ],
-						  include_dirs = numpy_inc,
-						  extra_objects = [],
-						  undef_macros=['NDEBUG'],
-						  #extra_compile_args=['-O0','-g'],
-						  #extra_link_args=['-O0', '-g'],
-						  )
-
-setup(name = 'EM fit of Gaussian mixture',
-	  version = '1.0',
-	  description = '',
-	  author = 'Lang & Hogg',
-	  author_email = 'dstn@astro.princeton.edu',
-	  url = 'http://astrometry.net',
-	  py_modules = [],
-	  ext_modules = [c_swig_module])
-

### 2012-05-09 setup-mix.py

@@ -1,23 +1,0 @@
-from distutils.core import setup, Extension
-from numpy.distutils.misc_util import get_numpy_include_dirs
-
-numpy_inc = get_numpy_include_dirs()
-
-c_swig_module = Extension('_mix',
-						  sources = ['mix_wrap.c' ],
-						  include_dirs = numpy_inc,
-						  extra_objects = [],
-						  undef_macros=['NDEBUG'],
-						  #extra_compile_args=['-O0','-g'],
-						  #extra_link_args=['-O0', '-g'],
-						  )
-
-setup(name = 'Gaussian mixtures',
-	  version = '1.0',
-	  description = '',
-	  author = 'Lang & Hogg',
-	  author_email = 'dstn@astro.princeton.edu',
-	  url = 'http://astrometry.net',
-	  py_modules = [],
-	  ext_modules = [c_swig_module])
-

### 2012-05-09 symmetric_integral_image.py

@@ -1,179 +1,0 @@
-# Copyright 2011 Dustin Lang.  All rights reserved.
-
-from numpy import *
-from integral_image import intimg_rect
-
-import unittest
-
-# Compute the area of the rectangular region in a first-quadrant
-# integral image.
-# (x0,y0) is the pixel one less than the first pixel included in the sum
-# (x1,y1) is the last  pixel included in the sum
-def symm_intimg_rect(img, xx0, xx1, yy0, yy1, midline):
-	'''
-	Computes the area of a given rectangle, given an integral image for
-	an image that is assumed to be mirror-symmetric in both axes.
-
-	That is, if the image size is N x N, then I[i, j] == I[N-i-1, j]
-	and I[i, j] == I[i, N-j-1].
-
-	The image must be odd-sized and square.
-
-	"midline" is (image size - 1)/2; if the image is 2N-1, midline == N.
-
-	"img" must be an integral image of size at least midline+1 x midline+1.
-
-	(xx0,yy0) are one less than the first pixel included in the sum.
-	(xx1,yy1) are the last pixel included in the sum.
-
-	(Note that the inclusiveness is opposite from convention.)
-
-	These can be (1-d) arrays, but they must be the same shape.
-
-	The return value is:
-
-        sum_{i=xx0 + 1}^{xx1} sum_{j = yy0 + 1}^{yy1} I[j,i]
-
-
-	>>> if True:
-	...     from numpy import fliplr, flipud
-	...     from numpy.random import random_sample, seed
-	...     from integral_image import *
-	...     seed(42)
-	...     I = random_sample((11,11))
-	...     # symmetrize by copying the first quadrant
-	...     I[:,6:] = fliplr(I[:,0:5])
-	...     I[6:,:] = flipud(I[0:5,:])
-	...     intimg = integral_image(I)
-	...     symmimg = integral_image(I[:6,:6])
-	...     for (x0,x1,y0,y1) in [(-1,0,-1,0), (3,7,4,8), (-1,10,-1,10), (5,6,-1,0)]:
-	...         R0 = intimg_rect(intimg, x0, x1, y0, y1)
[…]

### 2012-05-09 ttime.py

@@ -1,54 +1,0 @@
-import os
-import resource
-def memusage():
-	# print heapy.heap()
-	#ru = resource.getrusage(resource.RUSAGE_BOTH)
-	ru = resource.getrusage(resource.RUSAGE_SELF)
-	pgsize = resource.getpagesize()
-	print 'Memory usage:'
-	#print 'page size', pgsize
-	print 'max rss:', (ru.ru_maxrss * pgsize / 1e6), 'MB'
-	#print 'shared memory size:', (ru.ru_ixrss / 1e6), 'MB'
-	#print 'unshared memory size:', (ru.ru_idrss / 1e6), 'MB'
-	#print 'unshared stack size:', (ru.ru_isrss / 1e6), 'MB'
-	#print 'shared memory size:', ru.ru_ixrss
-	#print 'unshared memory size:', ru.ru_idrss
-	#print 'unshared stack size:', ru.ru_isrss
-	procfn = '/proc/%d/status' % os.getpid()
-	try:
-		t = open(procfn).readlines()
-		#print 'proc file:', t
-		d = dict([(line.split()[0][:-1], line.split()[1:]) for line in t])
-		#print 'dict:', d
-		for key in ['VmPeak', 'VmSize', 'VmRSS', 'VmData', 'VmStk' ]: # VmLck, VmHWM, VmExe, VmLib, VmPTE
-			print key, ' '.join(d.get(key, []))
-	except:
-		pass
-
-class Time(object):
-	@staticmethod
-	def add_measurement(m):
-		Time.measurements.append(m)
-	measurements = []
-
-	def __init__(self):
-		import datetime
-		from time import clock
-		self.wall = datetime.datetime.now()
-		#self.cpu = time.clock()
-		self.cpu = clock()
-		self.meas = [m() for m in Time.measurements]
-
-	def __sub__(self, other):
-		dwall = (self.wall - other.wall)
-		# python2.7
-		if hasattr(dwall, 'total_seconds'):
-			dwall = dwall.total_seconds()
-		else:
-			dwall = (dwall.microseconds + (dwall.seconds + dwall.days * 24. * 3600.) * 1e6) / 1e6
-		dcpu = (self.cpu - other.cpu)
-
-		meas = [m.format_diff(om) for m,om in zip(self.meas, other.meas)]
-		return ', '.join(['%f wall' % dwall, '%f cpu' % dcpu] + meas)
-
-

### 2012-05-09 utils.py

@@ -1,580 +1,0 @@
-"""
-This file is part of the Tractor project.
-Copyright 2011, 2012 Dustin Lang and David W. Hogg.
-Licensed under the GPLv2; see the file COPYING for details.
-
-`utils.py`
-===========
-
-Utility classes: low-level useful implementations of ducks the Tractor
-needs: sets of parameters stored in lists and such.  This framework
-could be useful outside the Tractor context.
-
-"""
-#from ducks import *
-
-def getClassName(obj):
-	name = getattr(obj.__class__, 'classname', None)
-	if name is not None:
-		return name
-	return obj.__class__.__name__
-
-class PlotSequence(object):
-	def __init__(self, basefn, format='%02i'):
-		self.ploti = 0
-		self.basefn = basefn
-		self.format = format
-		self.suff = 0
-	def skip(self, n=1):
-		self.ploti += n
-	def skipto(self, n):
-		self.ploti = n
-	def savefig(self):
-		fn = '%s-%s.png' % (self.basefn, self.format % self.ploti)
-		plt.savefig(fn)
-		print 'saved', fn
-		self.ploti += 1
-
-
-class BaseParams(object):
-	'''
-	A basic implementation of the `Params` duck type.
-	'''
-	def __repr__(self):
-		return getClassName(self) + repr(self.getParams())
-	def __str__(self):
-		return getClassName(self) + ': ' + str(self.getParams())
-	def copy(self):
-		return self.__class__(self.getParams())
-	def hashkey(self):
-		return (getClassName(self),) + tuple(self.getParams())
-	def __hash__(self):
-		return hash(self.hashkey())
-	def __eq__(self, other):
-		return hash(self) == hash(other)
-	def getParamNames(self):
-		''' Returns a list containing the names of the parameters. '''
-		return []
-	def numberOfParams(self):
-		''' Returns the number of parameters (ie, number of scalar
-		values).'''
-		return len(self.getParams())
-	def getParams(self):
-		''' Returns a *copy* of the current parameter values as an
-		iterable (eg, list)'''
-		return []
-	def getStepSizes(self, *args, **kwargs):
-		''' Returns "reasonable" step sizes for the parameters.'''
-		return []
-	def setParams(self, p):
-		'''
-		Sets the parameter values to the values in the given iterable
[…]

### 2012-05-09 code/optimize_mixture_profiles.py
project string

@@ -1,4 +1,5 @@
-# Copyright 2011 David W. Hogg.  All rights reserved.
+# this file is part of the Tractor project.
+# Copyright 2011 David W. Hogg.
 
 import matplotlib
 matplotlib.use('Agg')

### 2012-05-09 code/optimize_mixture_profiles.py
refactor to remove synchronized casing

@@ -11,158 +11,141 @@ import scipy.optimize as op
 # note wacky normalization because this is for 2-d Gaussians
 # (but only ever called in 1-d).  Wacky!
 def not_normal(x, V):
-	exparg = -0.5 * x**2 / V
-	result = np.zeros_like(x)
-	I = ((exparg > -1000) * (exparg < 1000))
-	result[I] = 1. / (2. * np.pi * V) * np.exp(exparg[I])
-	return result
+    exparg = -0.5 * x**2 / V
+    result = np.zeros_like(x)
+    I = ((exparg > -1000) * (exparg < 1000))
+    result[I] = 1. / (2. * np.pi * V) * np.exp(exparg[I])
+    return result
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
 def hogg_exp(x):
-	return np.exp(-1.67835 * (x - 1.))
+    return np.exp(-1.67835 * (x - 1.))
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
 def hogg_dev(x):
-	return np.exp(-7.66925 * ((x*x + 0.0004)**0.125 - 1.))
+    return np.exp(-7.66925 * ((x*x + 0.0004)**0.125 - 1.))
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
 def hogg_lup(x):
-	inner = 7.
-	outer = 8.
-	lup = hogg_dev(x)
-	outside = (x >= outer)
-	lup[outside] *= 0.
-	middle = (x >= inner) * (x <= outer)
-	lup[middle] *= (outer - x[middle]) / (outer - inner)
-	return lup
+    inner = 7.
+    outer = 8.
+    lup = hogg_dev(x)
+    outside = (x >= outer)
+    lup[outside] *= 0.
+    middle = (x >= inner) * (x <= outer)
+    lup[middle] *= (outer - x[middle]) / (outer - inner)
+    return lup
+
+def hogg_model(x, model):
+    if model == 'exp':
+        return hogg_exp(x)
+    if model == 'dev':
+        return hogg_dev(x)
+    if model == 'lup':
+        return hogg_lup(x)
+    assert(1 == 0)
+    return None
 
 def mixture_of_not_normals(x, pars):
-	K = len(pars)/2
-	y = 0.
-	for k in range(K):
-		y += pars[k] * not_normal(x, pars[k+K])
-	return y
-
[…]

### 2012-05-09 code/optimize_mixture_profiles.py
nonfunctional tweaks for Hogg style

@@ -82,7 +82,8 @@ def plot_mixture(pars, prefix, model):
     for k in range(K):
         plt.plot(x2, pars[k] * not_normal(x2, pars[k+K]), 'k-', alpha=0.5)
     plt.axvline(MAX_RADIUS, color='k', alpha=0.25)
-    plt.title(r"%s / $K=%d$ / maximum radius = $%.1f$ / badness = $%.2f\times 10^{%d}$" % (model, len(pars)/2, MAX_RADIUS, badness, LOG10_SQUARED_DEVIATION))
+    plt.title(r"%s / $K=%d$ / maximum radius = $%.1f$ / badness = $%.2f\times 10^{%d}$"
+              % (model, len(pars)/2, MAX_RADIUS, badness, LOG10_SQUARED_DEVIATION))
     plt.xlim(-0.1*MAX_RADIUS, 2.*MAX_RADIUS)
     plt.ylim(-0.1*np.max(y1), 1.1*np.max(y1))
     plt.savefig(prefix+'_'+model+'.png')
@@ -109,12 +110,12 @@ def main(model):
     bestbadness = badness
     for K in range(2,20):
         print 'working on K = %d' % K
-        newvar = 2.0 * np.max(np.append(var,1.0))
+        newvar = 2.0 * np.max(np.append(var, 1.0))
         newamp = 1.0 * newvar
         amp = np.append(newamp, amp)
         var = np.append(newvar, var)
         pars = np.append(amp, var)
-        for i in range(2*K):
+        for i in range(2 * K):
             (badness, pars) = optimize_mixture(K, pars, model)
             if (badness < bestbadness) or (i == 0):
                 print '%d %d improved' % (K, i)
@@ -122,9 +123,9 @@ def main(model):
                 bestbadness = badness
             else:
                 print '%d %d not improved' % (K, i)
-                print i, len(var), K, np.mod(i,K)
-                var[0] = 2.0 * var[np.mod(i,K)]
-                amp[0] = 0.5 * amp[np.mod(i,K)]
+                print i, len(var), K, np.mod(i, K)
+                var[0] = 2.0 * var[np.mod(i, K)]
+                amp[0] = 0.5 * amp[np.mod(i, K)]
                 pars = np.append(amp, var)
                 if (bestbadness < 0.5 * lastKbadness) and (i > 4):
                     print '%d %d improved enough' % (K, i)
[…]

### 2012-05-10 code/optimize_mixture_profiles.py
no more global variables; multiprocessing; plotting tweaks

@@ -50,51 +50,51 @@ def mixture_of_not_normals(x, pars):
     K = len(pars)/2
     y = 0.
     for k in range(K):
-        y += pars[k] * not_normal(x, pars[k+K])
+        y += pars[k] * not_normal(x, pars[k + K])
     return y
 
 # note that you can do (x * ymix - x * ytrue)**2 or (ymix - ytrue)**2
 # each has disadvantages.
 # note magic number in the penalty for large variance
-def badness_of_fit(lnpars, model):
+def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation):
     pars = np.exp(lnpars)
-    x = np.arange(0.0005, MAX_RADIUS, 0.001)
+    x = np.arange(0.0005, max_radius, 0.001)
     badness = np.mean((mixture_of_not_normals(x, pars)
-               - hogg_model(x, model))**2) / 10.**LOG10_SQUARED_DEVIATION
+               - hogg_model(x, model))**2) / 10.**log10_squared_deviation
     K = len(pars) / 2
     var = pars[K:]
-    extrabadness = 0.0001 * np.sum(var) / MAX_RADIUS**2
+    extrabadness = 0.0001 * np.sum(var) / max_radius**2
     return badness + extrabadness
 
-def optimize_mixture(K, pars, model):
-    newlnpars = op.fmin_bfgs(badness_of_fit, np.log(pars), args=(model, ))
-    return (badness_of_fit(newlnpars, model), np.exp(newlnpars))
+def optimize_mixture(K, pars, model, max_radius, log10_squared_deviation):
+    newlnpars = op.fmin_bfgs(badness_of_fit, np.log(pars), args=(model, max_radius, log10_squared_deviation))
+    return (badness_of_fit(newlnpars, model, max_radius, log10_squared_deviation), np.exp(newlnpars))
 
-def plot_mixture(pars, prefix, model):
-    x2 = np.arange(0., 10.*MAX_RADIUS, 0.001)
+def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation):
+    x2 = np.arange(0., 10. * max_radius, 0.001)
     y1 = hogg_model(x2, model)
-    badness = badness_of_fit(np.log(pars), model)
+    badness = badness_of_fit(np.log(pars), model, max_radius, log10_squared_deviation)
     K = len(pars) / 2
     y2 = mixture_of_not_normals(x2, pars)
[…]

### 2012-05-10 code/optimize_mixture_profiles.py
go to 8, small tweaks, add Sersic placeholder

@@ -11,7 +11,7 @@ import scipy.optimize as op
 # note wacky normalization because this is for 2-d Gaussians
 # (but only ever called in 1-d).  Wacky!
 def not_normal(x, V):
-    exparg = -0.5 * x**2 / V
+    exparg = -0.5 * x * x / V
     result = np.zeros_like(x)
     I = ((exparg > -1000) * (exparg < 1000))
     result[I] = 1. / (2. * np.pi * V) * np.exp(exparg[I])
@@ -23,7 +23,7 @@ def hogg_exp(x):
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
 def hogg_dev(x):
-    return np.exp(-7.66925 * ((x*x + 0.0004)**0.125 - 1.))
+    return np.exp(-7.66925 * ((x * x + 0.0004)**0.125 - 1.))
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
 def hogg_lup(x):
@@ -35,6 +35,12 @@ def hogg_lup(x):
     middle = (x >= inner) * (x <= outer)
     lup[middle] *= (outer - x[middle]) / (outer - inner)
     return lup
+
+# NEEDS MAGIC NUMBERS
+def hogg_ser(x, n):
+    def amp:
+        return 1.
+    return np.exp(amp(n) * ((x * x + 0.0004)**(0.5 * n) - 1.))
 
 def hogg_model(x, model):
     if model == 'exp':
@@ -60,7 +66,7 @@ def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation):
     pars = np.exp(lnpars)
     x = np.arange(0.0005, max_radius, 0.001)
     badness = np.mean((mixture_of_not_normals(x, pars)
-               - hogg_model(x, model))**2) / 10.**log10_squared_deviation
+                       - hogg_model(x, model))**2) / 10.**log10_squared_deviation
     K = len(pars) / 2
     var = pars[K:]
     extrabadness = 0.0001 * np.sum(var) / max_radius**2
@@ -143,14 +149,13 @@ def main(model):
             txtfile = open(prefix + '_' + model + '.txt', "w")
             txtfile.write("pars = %s" % repr(pars))
             txtfile.close
-        if bestbadness < 1.0:
+        if bestbadness < 1.0 and K >= 8:
             break
 
 if __name__ == '__main__':
     if True: # use multiprocessing
         from multiprocessing import Pool
[…]

### 2012-05-10 docs/mixture_models.tex
added figures but not captions
ciotti, lupton

@@
 \newcommand{\dev}{\mathrm{dev}}
+
+\newlength{\figwidth}
+\setlength{\figwidth}{0.49\textwidth}
 
@@
 
-...Discuss the cutoffs used by PHOTO and adjust $Z$ and $r$ accordingly...
+...Cite \citep{ciotti} re half-light radius constants...
+
+...Lupton profile as different from de~Vaucouleurs...Sersic profiles...
+
+...Discuss the cutoffs and regularization used by PHOTO and adjust $Z$ and $r$ accordingly...
+
+...Note normalization issues and how easy it is to normalize the mixture-of-Gaussian model...
 
@@
 Cappellari,~M., 2002, \mnras, 333, 400
+\bibitem[Ciotti \& Bertin(1999)]{ciotti}
+Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
 \bibitem[Emsellem \etal(1994)]{emsellem}
@@
 
+\clearpage
+\begin{figure}
+\includegraphics[width=\figwidth]{K08_MR08_LSD04_dev.png}
+\includegraphics[width=\figwidth]{K08_MR08_LSD04_dev_log.png} \\
+\includegraphics[width=\figwidth]{K08_MR08_LSD04_lup.png}
+\includegraphics[width=\figwidth]{K08_MR08_LSD04_lup_log.png}
+\end{figure}
+
 \end{document}

### 2012-05-10 code/optimize_mixture_profiles.py
sersic profiles, ciotti numbers

@@ -17,15 +17,18 @@ def not_normal(x, V):
     result[I] = 1. / (2. * np.pi * V) * np.exp(exparg[I])
     return result
 
-# magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
+# magic number from Ciotti & Bertin, A&A, 352, 447 (1999)
+# normalized to return 1. at x=1.
 def hogg_exp(x):
-    return np.exp(-1.67835 * (x - 1.))
+    return np.exp(-1.67834699 * (x - 1.))
 
-# magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
+# magic number from Ciotti & Bertin, A&A, 352, 447 (1999)
+# normalized to return 1. at x=1.
 def hogg_dev(x):
-    return np.exp(-7.66925 * ((x * x + 0.0004)**0.125 - 1.))
+    return np.exp(-7.66924944 * ((x * x)**0.125 - 1.))
 
 # magic numbers from Lupton (makeprof.c and phFitobj.h) via dstn
+# normalized to return 1. at x=1.
 def hogg_lup(x):
     inner = 7.
     outer = 8.
@@ -36,11 +39,16 @@ def hogg_lup(x):
     lup[middle] *= (outer - x[middle]) / (outer - inner)
     return lup
 
-# NEEDS MAGIC NUMBERS
-def hogg_ser(x, n):
-    def amp:
-        return 1.
-    return np.exp(amp(n) * ((x * x + 0.0004)**(0.5 * n) - 1.))
+# magic numbers from Ciotti & Bertin, A&A, 352, 447 (1999)
+# normalized to return 1. at x=1.
+def hogg_ser2(x):
+    return np.exp(-3.67206075 * ((x * x)**0.25 - 1.))
+def hogg_ser3(x):
+    return np.exp(-5.67016119 * ((x * x)**(1./6.) - 1.))
+def hogg_ser4(x):
+    return np.exp(-7.66924944 * ((x * x)**0.125 - 1.))
+def hogg_ser5(x):
+    return np.exp(-9.66871461 * ((x * x)**0.1 - 1.))
 
 def hogg_model(x, model):
     if model == 'exp':
@@ -49,6 +57,14 @@ def hogg_model(x, model):
         return hogg_dev(x)
     if model == 'lup':
         return hogg_lup(x)
+    if model == 'ser2':
+        return hogg_ser2(x)
+    if model == 'ser3':
[…]

### 2012-05-11 code/optimize_mixture_profiles.py
restored lupton tweaks

@@ -32,7 +32,7 @@ def hogg_dev(x):
 def hogg_lup(x):
     inner = 7.
     outer = 8.
-    lup = hogg_dev(x)
+    lup = np.exp(-7.66925 * ((x * x + 0.0004)**0.125 - 1.))
     outside = (x >= outer)
     lup[outside] *= 0.
     middle = (x >= inner) * (x <= outer)
@@ -93,7 +93,7 @@ def optimize_mixture(K, pars, model, max_radius, log10_squared_deviation):
     return (badness_of_fit(newlnpars, model, max_radius, log10_squared_deviation), np.exp(newlnpars))
 
 def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation):
-    x2 = np.arange(0., 10. * max_radius, 0.001)
+    x2 = np.arange(0.0005, np.sqrt(5. * max_radius), 0.001)**2 # note non-linear spacing
     y1 = hogg_model(x2, model)
     badness = badness_of_fit(np.log(pars), model, max_radius, log10_squared_deviation)
     K = len(pars) / 2
@@ -165,7 +165,7 @@ def main(model):
             txtfile = open(prefix + '_' + model + '.txt', "w")
             txtfile.write("pars = %s" % repr(pars))
             txtfile.close
-        if bestbadness < 1.0 and K >= 8:
+        if bestbadness < 1.0 and K > 9:
             break
 
 if __name__ == '__main__':

### 2012-05-11 code/optimize_mixture_profiles.py
no more ser4; better log output

@@ -45,8 +45,6 @@ def hogg_ser2(x):
     return np.exp(-3.67206075 * ((x * x)**0.25 - 1.))
 def hogg_ser3(x):
     return np.exp(-5.67016119 * ((x * x)**(1./6.) - 1.))
-def hogg_ser4(x):
-    return np.exp(-7.66924944 * ((x * x)**0.125 - 1.))
 def hogg_ser5(x):
     return np.exp(-9.66871461 * ((x * x)**0.1 - 1.))
 
@@ -61,8 +59,6 @@ def hogg_model(x, model):
         return hogg_ser2(x)
     if model == 'ser3':
         return hogg_ser3(x)
-    if model == 'ser4':
-        return hogg_ser4(x)
     if model == 'ser5':
         return hogg_ser5(x)
     assert(1 == 0)
@@ -135,7 +131,7 @@ def main(model):
     lastKbadness = badness
     bestbadness = badness
     for K in range(2, 20):
-        print 'working on K = %d' % K
+        print 'working on %s at K = %d' % (model, K)
         newvar = 2.0 * np.max(np.append(var, 1.0))
         newamp = 1.0 * newvar
         amp = np.append(newamp, amp)
@@ -144,17 +140,16 @@ def main(model):
         for i in range(2 * K):
             (badness, pars) = optimize_mixture(K, pars, model, max_radius, log10_squared_deviation)
             if (badness < bestbadness) or (i == 0):
-                print '%d %d improved' % (K, i)
+                print '%s %d %d improved' % (model, K, i)
                 bestpars = pars
                 bestbadness = badness
             else:
-                print '%d %d not improved' % (K, i)
-                print i, len(var), K, np.mod(i, K)
+                print '%s %d %d not improved' % (model, K, i)
                 var[0] = 2.0 * var[np.mod(i, K)]
                 amp[0] = 0.5 * amp[np.mod(i, K)]
                 pars = np.append(amp, var)
                 if (bestbadness < 0.5 * lastKbadness) and (i > 4):
-                    print '%d %d improved enough' % (K, i)
+                    print '%s %d %d improved enough' % (model, K, i)
                     break
             lastKbadness = bestbadness
             pars = rearrange_pars(bestpars)
@@ -174,7 +169,5 @@ if __name__ == '__main__':
         pmap = Pool(8).map
     else: # don't use multiprocessing
         pmap = map
[…]

### 2012-05-11 code/optimize_mixture_profiles.py
do radius-weighted badness too

@@ -1,5 +1,10 @@
-# this file is part of the Tractor project.
-# Copyright 2011 David W. Hogg.
+'''
+this file is part of the Tractor project.
+Copyright 2011 David W. Hogg.
+
+### bugs:
+- should use levmar not bfgs!
+'''
 
 import matplotlib
 matplotlib.use('Agg')
@@ -74,24 +79,30 @@ def mixture_of_not_normals(x, pars):
 # note that you can do (x * ymix - x * ytrue)**2 or (ymix - ytrue)**2
 # each has disadvantages.
 # note magic number in the penalty for large variance
-def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation):
+def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation, radius_weighted):
     pars = np.exp(lnpars)
     x = np.arange(0.0005, max_radius, 0.001)
-    badness = np.mean((mixture_of_not_normals(x, pars)
-                       - hogg_model(x, model))**2) / 10.**log10_squared_deviation
+    residual = mixture_of_not_normals(x, pars) - hogg_model(x, model)
+    if radius_weighted:
+        numerator = np.sum(x * residual * residual)
+        denominator = np.sum(x)
+    else:
+        numerator = np.sum(residual * residual)
+        denominator = float(x.size)
+    badness = numerator / denominator / 10.**log10_squared_deviation
     K = len(pars) / 2
     var = pars[K:]
     extrabadness = 0.0001 * np.sum(var) / max_radius**2
     return badness + extrabadness
 
-def optimize_mixture(K, pars, model, max_radius, log10_squared_deviation):
-    newlnpars = op.fmin_bfgs(badness_of_fit, np.log(pars), args=(model, max_radius, log10_squared_deviation))
-    return (badness_of_fit(newlnpars, model, max_radius, log10_squared_deviation), np.exp(newlnpars))
+def optimize_mixture(K, pars, model, max_radius, log10_squared_deviation, radius_weighted):
+    newlnpars = op.fmin_bfgs(badness_of_fit, np.log(pars), args=(model, max_radius, log10_squared_deviation, radius_weighted))
+    return (badness_of_fit(newlnpars, model, max_radius, log10_squared_deviation, radius_weighted), np.exp(newlnpars))
 
-def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation):
+def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted):
[…]

### 2012-05-11 code/optimize_mixture_profiles.py
LSD tweaks

@@ -135,7 +135,7 @@ def rearrange_pars(pars):
 def main(input):
     model, radius_weighted = input
     max_radius = 8.0
-    log10_squared_deviation = -4
+    log10_squared_deviation = -6
     amp = np.array([1.0])
     var = np.array([1.0])
     pars = np.append(amp, var)
@@ -167,14 +167,14 @@ def main(input):
             pars = rearrange_pars(bestpars)
             amp = pars[0:K]
             var = pars[K:K+K]
-            prefix = 'K%02d_MR%02d_LSD%02d' % (K, int(round(max_radius) + 0.01), -1 * log10_squared_deviation)
+            prefix = 'K%02d_MR%02d' % (K, int(round(max_radius) + 0.01))
             if radius_weighted:
                 prefix += "_w"
             plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
             txtfile = open(prefix + '_' + model + '.txt', "w")
             txtfile.write("pars = %s" % repr(pars))
             txtfile.close
-        if bestbadness < 1.0 and K > 9:
+        if bestbadness < 100.0 and K > 9:
             break
 
 if __name__ == '__main__':

### 2012-05-11 code/optimize_mixture_profiles.py
more lsd tweaks

@@ -114,7 +114,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
     badname = "badness"
     if radius_weighted: badname = "weighted " + badname
     plt.title(r"%s / $K=%d$ / max radius = $%.1f$ / %s = $%.3f\times 10^{%d}$"
-              % (model, len(pars) / 2, max_radius, badname, badness, log10_squared_deviation))
+              % (model, len(pars) / 2, max_radius, badname, badness * 100., log10_squared_deviation - 2.))
     plt.xlim(-0.1 * max_radius, 1.2 * max_radius)
     plt.ylim(-0.1 * 8.0, 1.1 * 8.0)
     plt.savefig(prefix+'_'+model+'.png')
@@ -135,7 +135,7 @@ def rearrange_pars(pars):
 def main(input):
     model, radius_weighted = input
     max_radius = 8.0
-    log10_squared_deviation = -6
+    log10_squared_deviation = -4
     amp = np.array([1.0])
     var = np.array([1.0])
     pars = np.append(amp, var)
@@ -174,7 +174,7 @@ def main(input):
             txtfile = open(prefix + '_' + model + '.txt', "w")
             txtfile.write("pars = %s" % repr(pars))
             txtfile.close
-        if bestbadness < 100.0 and K > 9:
+        if bestbadness < 1.0 and K > 9:
             break
 
 if __name__ == '__main__':

### 2012-05-11 code/optimize_mixture_profiles.py
plot names

@@ -117,11 +117,11 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
               % (model, len(pars) / 2, max_radius, badname, badness * 100., log10_squared_deviation - 2.))
     plt.xlim(-0.1 * max_radius, 1.2 * max_radius)
     plt.ylim(-0.1 * 8.0, 1.1 * 8.0)
-    plt.savefig(prefix+'_'+model+'.png')
+    plt.savefig(model + '_' + prefix + '.png')
     plt.loglog()
     plt.xlim(0.001, 5. * max_radius)
     plt.ylim(3.e-5, 1.5 * np.max(y1))
-    plt.savefig(prefix+'_'+model+'_log.png')
+    plt.savefig(model + '_' + prefix + '_log.png')
 
 def rearrange_pars(pars):
     K = len(pars) / 2

### 2012-05-11 code/optimize_mixture_profiles.py
plot names

@@ -117,11 +117,11 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
               % (model, len(pars) / 2, max_radius, badname, badness * 100., log10_squared_deviation - 2.))
     plt.xlim(-0.1 * max_radius, 1.2 * max_radius)
     plt.ylim(-0.1 * 8.0, 1.1 * 8.0)
-    plt.savefig(model + '_' + prefix + '.png')
+    plt.savefig(prefix + '.png')
     plt.loglog()
     plt.xlim(0.001, 5. * max_radius)
     plt.ylim(3.e-5, 1.5 * np.max(y1))
-    plt.savefig(model + '_' + prefix + '_log.png')
+    plt.savefig(prefix + '_log.png')
 
 def rearrange_pars(pars):
     K = len(pars) / 2
@@ -167,11 +167,11 @@ def main(input):
             pars = rearrange_pars(bestpars)
             amp = pars[0:K]
             var = pars[K:K+K]
-            prefix = 'K%02d_MR%02d' % (K, int(round(max_radius) + 0.01))
+            prefix = '%s_K%02d_MR%02d' % (model, K, int(round(max_radius) + 0.01))
             if radius_weighted:
                 prefix += "_w"
             plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
-            txtfile = open(prefix + '_' + model + '.txt', "w")
+            txtfile = open(prefix + '.txt', "w")
             txtfile.write("pars = %s" % repr(pars))
             txtfile.close
         if bestbadness < 1.0 and K > 9:

### 2012-05-11 code/optimize_mixture_profiles.py
fixed indentation bug

@@ -143,7 +143,10 @@ def main(input):
     lastKbadness = badness
     bestbadness = badness
     for K in range(2, 20):
-        print 'working on %s at K = %d' % (model, K)
+        prefix = '%s_K%02d_MR%02d' % (model, K, int(round(max_radius) + 0.01))
+        if radius_weighted:
+            prefix += "_w"
+        print 'working on %s at K = %d (%s)' % (model, K, prefix)
         newvar = 2.0 * np.max(np.append(var, 1.0))
         newamp = 1.0 * newvar
         amp = np.append(newamp, amp)
@@ -163,17 +166,14 @@ def main(input):
                 if (bestbadness < 0.5 * lastKbadness) and (i > 4):
                     print '%s %d %d improved enough' % (model, K, i)
                     break
-            lastKbadness = bestbadness
-            pars = rearrange_pars(bestpars)
-            amp = pars[0:K]
-            var = pars[K:K+K]
-            prefix = '%s_K%02d_MR%02d' % (model, K, int(round(max_radius) + 0.01))
-            if radius_weighted:
-                prefix += "_w"
-            plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
-            txtfile = open(prefix + '.txt', "w")
-            txtfile.write("pars = %s" % repr(pars))
-            txtfile.close
+        lastKbadness = bestbadness
+        pars = rearrange_pars(bestpars)
+        amp = pars[0:K]
+        var = pars[K:K+K]
+        plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
+        txtfile = open(prefix + '.txt', "w")
+        txtfile.write("pars = %s" % repr(pars))
+        txtfile.close
         if bestbadness < 1.0 and K > 9:
             break
 
@@ -188,11 +188,5 @@ if __name__ == '__main__':
               ('ser2', True),
               ('ser3', True),
               ('dev', True),
-              ('ser5', True),
-              ('lup', False),
-              ('exp', False),
-              ('ser2', False),
-              ('ser3', False),
-              ('dev', False),
-              ('ser5', False)]
+              ('ser5', True), ]
     pmap(main, inputs)

### 2012-05-11 code/optimize_mixture_profiles.py
more Ks

@@ -174,7 +174,7 @@ def main(input):
         txtfile = open(prefix + '.txt', "w")
         txtfile.write("pars = %s" % repr(pars))
         txtfile.close
-        if bestbadness < 1.0 and K > 9:
+        if bestbadness < 1.0 and K > 15:
             break
 
 if __name__ == '__main__':

### 2012-05-11 code/optimize_mixture_profiles.py
latex fonts

@@ -1,17 +1,20 @@
 '''
-this file is part of the Tractor project.
+This file is part of the Tractor project.
 Copyright 2011 David W. Hogg.
 
 ### bugs:
-- should use levmar not bfgs!
+- Should use levmar not bfgs!
 '''
 
 import matplotlib
 matplotlib.use('Agg')
+from matplotlib import rc
+rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':12})
+rc('text', usetex=True)
 import pylab as plt
-import matplotlib.cm as cm
 import numpy as np
 import scipy.optimize as op
+from multiprocessing import Pool
 
 # note wacky normalization because this is for 2-d Gaussians
 # (but only ever called in 1-d).  Wacky!
@@ -179,7 +182,6 @@ def main(input):
 
 if __name__ == '__main__':
     if True: # use multiprocessing
-        from multiprocessing import Pool
         pmap = Pool(10).map
     else: # don't use multiprocessing
         pmap = map

### 2012-05-11 code/optimize_mixture_profiles.py
optimization adjustments

@@ -115,7 +115,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
         plt.plot(x2, pars[k] * not_normal(x2, pars[k + K]), 'k-', alpha=0.5)
     plt.axvline(max_radius, color='k', alpha=0.25)
     badname = "badness"
-    if radius_weighted: badname = "weighted " + badname
+    if not radius_weighted: badname = "unweighted " + badname
     plt.title(r"%s / $K=%d$ / max radius = $%.1f$ / %s = $%.3f\times 10^{%d}$"
               % (model, len(pars) / 2, max_radius, badname, badness * 100., log10_squared_deviation - 2.))
     plt.xlim(-0.1 * max_radius, 1.2 * max_radius)
@@ -159,16 +159,15 @@ def main(input):
             (badness, pars) = optimize_mixture(K, pars, model, max_radius, log10_squared_deviation, radius_weighted)
             if (badness < bestbadness) or (i == 0):
                 print '%s %d %d improved' % (model, K, i)
-                bestpars = pars
+                bestpars = rearrange_pars(pars)
                 bestbadness = badness
             else:
                 print '%s %d %d not improved' % (model, K, i)
+                amp = bestpars[0:K]
+                var = bestpars[K:K+K]
                 var[0] = 2.0 * var[np.mod(i, K)]
                 amp[0] = 0.5 * amp[np.mod(i, K)]
                 pars = np.append(amp, var)
-                if (bestbadness < 0.5 * lastKbadness) and (i > 4):
-                    print '%s %d %d improved enough' % (model, K, i)
-                    break
         lastKbadness = bestbadness
         pars = rearrange_pars(bestpars)
         amp = pars[0:K]

### 2012-05-11 code/optimize_mixture_profiles.py
write pickles too

@@ -15,6 +15,7 @@ import pylab as plt
 import numpy as np
 import scipy.optimize as op
 from multiprocessing import Pool
+import cPickle as pickle
 
 # note wacky normalization because this is for 2-d Gaussians
 # (but only ever called in 1-d).  Wacky!
@@ -174,8 +175,11 @@ def main(input):
         var = pars[K:K+K]
         plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
         txtfile = open(prefix + '.txt', "w")
-        txtfile.write("pars = %s" % repr(pars))
+        txtfile.write("%s\n" % repr(pars))
         txtfile.close
+        pckfile = open(prefix + '.pickle', "wb")
+        pickle.dump(pars, pckfile)
+        pckfile.close
         if bestbadness < 1.0 and K > 15:
             break
 

### 2012-05-11 code/optimize_mixture_profiles.py
naming tweaks

@@ -148,8 +148,8 @@ def main(input):
     bestbadness = badness
     for K in range(2, 20):
         prefix = '%s_K%02d_MR%02d' % (model, K, int(round(max_radius) + 0.01))
-        if radius_weighted:
-            prefix += "_w"
+        if not radius_weighted:
+            prefix += "_uw"
         print 'working on %s at K = %d (%s)' % (model, K, prefix)
         newvar = 2.0 * np.max(np.append(var, 1.0))
         newamp = 1.0 * newvar
@@ -177,9 +177,9 @@ def main(input):
         txtfile = open(prefix + '.txt', "w")
         txtfile.write("%s\n" % repr(pars))
         txtfile.close
-        pckfile = open(prefix + '.pickle', "wb")
-        pickle.dump(pars, pckfile)
-        pckfile.close
+        picklefile = open(prefix + '.pickle', "wb")
+        pickle.dump(pars, picklefile)
+        picklefile.close
         if bestbadness < 1.0 and K > 15:
             break
 

### 2012-05-11 code/optimize_mixture_profiles.py
huh

@@ -164,8 +164,8 @@ def main(input):
                 bestbadness = badness
             else:
                 print '%s %d %d not improved' % (model, K, i)
-                amp = bestpars[0:K]
-                var = bestpars[K:K+K]
+                amp = 1. * bestpars[0:K]
+                var = 1. * bestpars[K:K+K]
                 var[0] = 2.0 * var[np.mod(i, K)]
                 amp[0] = 0.5 * amp[np.mod(i, K)]
                 pars = np.append(amp, var)
@@ -188,10 +188,10 @@ if __name__ == '__main__':
         pmap = Pool(10).map
     else: # don't use multiprocessing
         pmap = map
-    inputs = [('lup', True),
+    inputs = [('dev', True),
+              ('lup', True),
               ('exp', True),
               ('ser2', True),
               ('ser3', True),
-              ('dev', True),
               ('ser5', True), ]
     pmap(main, inputs)

### 2012-05-12 code/optimize_mixture_profiles.py
dynamically adjust badness amplitude (lsd)

@@ -118,7 +118,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
     badname = "badness"
     if not radius_weighted: badname = "unweighted " + badname
     plt.title(r"%s / $K=%d$ / max radius = $%.1f$ / %s = $%.3f\times 10^{%d}$"
-              % (model, len(pars) / 2, max_radius, badname, badness * 100., log10_squared_deviation - 2.))
+              % (model, len(pars) / 2, max_radius, badname, badness, log10_squared_deviation))
     plt.xlim(-0.1 * max_radius, 1.2 * max_radius)
     plt.ylim(-0.1 * 8.0, 1.1 * 8.0)
     plt.savefig(prefix + '.png')
@@ -139,7 +139,7 @@ def rearrange_pars(pars):
 def main(input):
     model, radius_weighted = input
     max_radius = 8.0
-    log10_squared_deviation = -4
+    log10_squared_deviation = -2.
     amp = np.array([1.0])
     var = np.array([1.0])
     pars = np.append(amp, var)
@@ -162,6 +162,9 @@ def main(input):
                 print '%s %d %d improved' % (model, K, i)
                 bestpars = rearrange_pars(pars)
                 bestbadness = badness
+                while bestbadness < 1.:
+                    bestbadness *= 10.
+                    log10_squared_deviation -= 1.
             else:
                 print '%s %d %d not improved' % (model, K, i)
                 amp = 1. * bestpars[0:K]

### 2012-05-13 code/optimize_mixture_profiles.py
close is a function

@@ -179,10 +179,10 @@ def main(input):
         plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radius_weighted)
         txtfile = open(prefix + '.txt', "w")
         txtfile.write("%s\n" % repr(pars))
-        txtfile.close
+        txtfile.close()
         picklefile = open(prefix + '.pickle', "wb")
         pickle.dump(pars, picklefile)
-        picklefile.close
+        picklefile.close()
         if bestbadness < 1.0 and K > 15:
             break
 

### 2012-05-13 code/optimize_mixture_profiles.py
dynamically adjust badness amplitude (lsd)

@@ -187,13 +187,13 @@ def main(input):
             break
 
 if __name__ == '__main__':
-    if True: # use multiprocessing
+    if False: # use multiprocessing
         pmap = Pool(10).map
     else: # don't use multiprocessing
         pmap = map
-    inputs = [('dev', True),
+    inputs = [('exp', True),
               ('lup', True),
-              ('exp', True),
+              ('dev', True),
               ('ser2', True),
               ('ser3', True),
               ('ser5', True), ]

### 2012-05-13 code/optimize_mixture_profiles.py
plotting and lsd tweaks

@@ -4,6 +4,7 @@ Copyright 2011 David W. Hogg.
 
 ### bugs:
 - Should use levmar not bfgs!
+- Spews enormous numbers of warnings when log10_squared_deviation gets very negative.
 '''
 
 import matplotlib
@@ -123,8 +124,10 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, radiu
     plt.ylim(-0.1 * 8.0, 1.1 * 8.0)
     plt.savefig(prefix + '.png')
     plt.loglog()
-    plt.xlim(0.001, 5. * max_radius)
-    plt.ylim(3.e-5, 1.5 * np.max(y1))
+    xmin = 0.001
+    yint = np.interp([xmin, ], x2, y1)[0]
+    plt.xlim(xmin, 5. * max_radius)
+    plt.ylim(1.5e-6 * yint, 1.5 * yint)
     plt.savefig(prefix + '_log.png')
 
 def rearrange_pars(pars):
@@ -162,7 +165,7 @@ def main(input):
                 print '%s %d %d improved' % (model, K, i)
                 bestpars = rearrange_pars(pars)
                 bestbadness = badness
-                while bestbadness < 1.:
+                while bestbadness < 1. and log10_squared_deviation > -5.5:
                     bestbadness *= 10.
                     log10_squared_deviation -= 1.
             else:
@@ -187,14 +190,16 @@ def main(input):
             break
 
 if __name__ == '__main__':
-    if False: # use multiprocessing
-        pmap = Pool(10).map
+    if True: # use multiprocessing
+        pmap = Pool(6).map
     else: # don't use multiprocessing
         pmap = map
-    inputs = [('exp', True),
-              ('lup', True),
-              ('dev', True),
-              ('ser2', True),
-              ('ser3', True),
-              ('ser5', True), ]
+    inputs = [
+        ('exp', True),
+        ('dev', True),
+        ('lup', True),
+        ('ser2', True),
+        ('ser3', True),
+        ('ser5', True),
+        ]
     pmap(main, inputs)

### 2012-07-28 documents/herschel_dust.tex
start Herschel project; abstract for Herschel project

--- 2012-05-08 docs/mixture_models.tex
@@
 % -----
-% - write zeroth draft
-% - make useful figures
-% - send to Jim Bosch and Kevin Bundy for comments
+% - outline
+% - make figures
+% - write
+% - get comments
+% - submit
 
@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\tractor}{\project{Tractor}}
+\newcommand{\TheTractor}{\project{The~Tractor}}
 \newcommand{\sdss}{\project{SDSS}}
+\newcommand{\Herschel}{\project{Herschel}}
 
@@
 
+\newlength{\figwidth}
+\setlength{\figwidth}{0.49\textwidth}
+
 \begin{document}
 
-\title{Replacing standard galaxy profiles with \\ mixtures of Gaussians}
-\author{David W. Hogg\altaffilmark{1,2,3} \&
-        Dustin Lang\altaffilmark{4}}
-\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
+\title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
+\author{Lang, Hogg, etc}
+\altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
 \altaffiltext{2}{New York University}
@@
 \begin{abstract}
-Exponential and de~Vaucouleurs profiles are simple and successful
-models for fitting two-dimensional images of galaxies.  One numerical
-issue encountered in this kind of fitting is the pixel rendering and
-convolution (or correlation) of the models with the telescope
-point-spread function (PSF); these operations are slow, and easy to
-get slightly wrong at small radii.  Here we exploit the realization
-that these models can be approximated to arbitrary accuracy with a
-mixture (linear superposition) of two-dimensional Gaussians.  Mixtures
-of Gaussians are fast to render and fast to affine-transform.  Most
-importantly, if you have a mixture-of-Gaussian PSF model (which we
-advocate), the PSF-convolved, affine-transformed galaxy models are
-themselves mixtures of Gaussians and therefore very fast to compute
-precisely.  We present worked examples that can be directly used in
-image fitting; we are using them ourselves.  We discuss extensions to
-Sersic models and applications of these ideas to three-dimensional
-deprojection.
+The \Herschel\ Observatory takes images at wavelengths
+$100<\lambda<500\,\micron$ and thereby is excellent for measuring the
+density and temperature of interstellar dust.  Being
+diffraction-limited, however, the images have angular resolution (or
+beam size) that is a strong function of wavelength.  In this
+\documentname\ we show that it is possible, with very weak priors or
+regularization, to infer the properties of the dust---column,
+temperature, and emissivity parameter---at the angular resolution of
+the highest resolution images.  This is true even though much of the
+crucial information comes from the lowest-resolution images.  The
+method works by optimizing the posterior probability of a generative
+model for the data; the model involves convolution with the
+pixel-convolved beam for each image.  The method is demonstrated on a
+small patch of \Herschel\ imaging of the dust disk of M31.  The method
+can be used in similar circumstances for other \Herschel\ data or on
+data from other observatories.  All the code is available under an
+open-source license as part of \TheTractor\ image-modeling project.
 \end{abstract}
 
-Gaussians are remarkable distribution functions.  They have the
-incredible properties that---in any number of dimensions---the
-convolution (or correlation) of one multivariate Gaussian with another
-is itself a multivariate Gaussian, and any product of multivariate
-Gaussians is itself a multivariate Gaussian but with a different
-normalization.  Furthermore, the means and variance tensors of the
-Gaussians output by these operations are related simply to the means
-and variance tensors of the inputs.  Add to these wonders the fact
-that Gaussians form a complete basis for representing (smooth)
-probability distribution functions and it becomes remarkable that we
-don't do everything we do in terms of Gaussians.
-
-To elaborate, a mixture of multivariate Gaussians---a linear
-superposition---can be used to represent any reasonable distribution
-in any number of dimensions to any reasonable precision.  Convolution
-(or correlation) by any other distribution that has also been
-represented by a mixture of multivariate Gaussians creates a new
-mixture of Gaussians with simply adjusted amplitudes, means, and
-variance tensors.  The ubiquity of convolution operations in astronomy
-suggests the widespread adoption of mixture-of-Gaussian modeling.  We
-have pioneered this in the area of distribution modeling in large
-numbers of dimensions \citep{xd, xdqso, xdqsoz}, where convolution
-occurs because the true or noise-free distribution is convolved with
-the noise before being observed.  Here we are going to capitalize on
-the convolution properties of mixtures of Gaussians in modeling galaxy
-morphologies in imaging data, where the true or
-high-angular-resolution intensity field is convolved with the
-point-spread function (PSF) before being observed.  This is
-particularly useful in data sets in which the PSF itself has also been
-modeled as a mixture of Gaussians, which is not uncommon.  For
-example, the \project{Sloan Digital Sky Survey} (\sdss; [HOGG cite
-  York]) imaging pipelines [HOGG cite Lupton] makes such models for
-every imaging field.
-
-We are not the first in this space deconvolution and modeling of
-galaxy images with mixtures of Gaussians has been done very
-successfully before \citep[for example,][]{bendinelli, emsellem,
-  bendinelli2, cappellari}.  However, the idea behind those projects
-was to use the mixture of Gaussians to provide a very free form for
-the isophotes or morphologies of resolved galaxies or other complex
-scenes.  Here our goals are very limited: We want to improve the
-performance of standard galaxy image model fitting by expressing the
-standard galaxy models---the exponential and de~Vaucouleurs
-profiles---as rigid mixtures of Gaussians.
-
-The performance advantages we will obtain from using mixtures of
-Gaussians are not simply that convolution itself is trivial.  The
-standard unconvolved galaxy models, especially the de~Vaucouleurs and
-S\'ersic models, are very ill-behaved near the galaxy center.
-Rendering these profiles precisely near the center can be very
-challenging numerically.  In addition, most of the rendering time is
-spent at very small radii, where PSF convolution is going to erase all
-structure anyway.  That is, a mixture-of-Gaussians description of the
-de~Vaucouleurs profile saves time both in rendering \emph{and} in PSF
-convolution; it produces profiles that are approximate but very high
-performance when all real uses of the profiles are PSF-convolved, as
-they usually are.
-
-...Limit scope of this document to just what we need for the
-\tractor...
-
-...somehow have to cite GALFIT \citep{galfit} in this section...
-
-...We need to put somewhere Lupton-like comments about sampling the
-pixel-convolved-PSF-convolved scene...
-
-In what follows we will make heavy use of two-dimensional Gaussian or
-Normal distributions, which look like
-\begin{eqnarray}\displaystyle
-\normal(\pos\given\mean,\var) &\equiv& \frac{1}{2\pi}\,\det(\var)^{-1/2}\,\exp(-\frac{1}{2}\,\transpose{[\pos-\mean]}\cdot\inverse{\var}\cdot[\pos-\mean])
-\quad ,
-\end{eqnarray}
-where $\pos$ and $\mean$ are two-dimensional vectors (usually in the
-focal plane or on the sky or something like that), and $\var$ is a
-symmetric $2\times 2$ variance tensor or matrix, and implicitly the
-vectors are column vectors.  A \emph{mixture of Gaussians} is a linear
-superposition of Gaussians.  Any positive two-dimensional function
-with finite support and finite total integral---including as a special
-case any two-dimensional probability distribution function---can be
-represented as a mixture of Gaussians to arbitrary accuracy (DSTN: It
-would be nice to have a reference here).
-
-We wish to make an approximation $Q_{\exp}(\cdot)$ to the
-two-dimensional circular exponential profile
-\begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\approx& \frac{1}{Z_{\exp}}\,\exp(-\frac{|\spos|}{r_{\exp}})
-\\
-Z_{\exp} &=& X.XXXX
-\\
-r_{\exp} &=& X.XXXX
-\quad ,
-\end{eqnarray}
-where $\spos$ is a dimensionless focal-plane position, and $Z_{\exp}$
-and $r_{\exp}$ are parameters that ensure that the profile integrates
-to unit flux and has unit half-light radius.  The position $\spos$ is
-dimensionless because it parameterizes the unit-size and unit-flux
-dimensionless function.  We seek the best (where ``best'' will be
-defined below) $M_g$-Gaussian mixture approximation
-\begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &=& \sum_{m=1}^{M_g} g_m\,\normal(\spos\given\zero,\Gm_m)
-\\
-1 &=& \sum_{m=1}^{M_g} g_m
-\\
-\Gm_m &=& \sigma^2_{\exp,m}\,\identity
-\quad ,
-\end{eqnarray}
-where all of the means are exactly zero and all of the variances
-$\Gm_m$ in the mixture are scalar multiples of the identity matrix
-$\identity$ because we are requiring this dimensionless function to be
-precisely circular (so every component is itself circular and
-concentric).  Similarly for the de~Vaucouleurs profile
-\begin{eqnarray}\displaystyle
-Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\left[\frac{|\spos|}{r_{\dev}}\right]^{1/4})
-\\
-Z_{\dev} &=& X.XXXX
-\\
-r_{\dev} &=& X.XXXX
-\\
-Q_{\dev}(\spos) &=& \sum_{m=1}^{M_h} h_m\,\normal(\spos\given\zero,\Hm_m)
-\\
-1 &=& \sum_{m=1}^{M_h} h_m
-\\
-\Hm_m &=& \sigma^2_{\dev,m}\,\identity
-\quad .
-\end{eqnarray}
-
-...Discuss the cutoffs used by PHOTO and adjust $Z$ and $r$ accordingly...
-
-...If we wanted to approximate these models, what objective function
-would we use?  What tolerances are okay?...Define a sensible scalar
-objective...Give fitting procedure...give actual values for the
-parameters...
-
-...Show plots demonstrating quality of fit, one-d and two-d plots...
-
-The value of these mixture-of-Gaussian galaxy components comes when
-they are to be convolved with a PSF (usually in fact a pixel-convolved
-PSF) that is itself represented as a mixture of Gaussians.  In this
-scenario, the PSF $\psi(\Delta\pos)$---which is thought of as a
-function of focal-plane displacement $\Delta\pos$ away from, say, a
-true stellar position---is represented as a $K$-Gaussian mixture
-\begin{eqnarray}\displaystyle
-\psi(\Delta\pos) &=& \sum_{k=1}^K p_k\,\normal(\Delta\pos\given\mean_k,\var_k)
-\\
-1 &=& \sum_{k=1}^K p_k
-\quad ,
-\end{eqnarray}
-where the means $\mean_k$ are \emph{not} required to vanish because
-the PSF can have arbitrarily non-trivial structure (think speckles and
-the like) and the variances $\var_k$ will not in general be
-proportional to the identity or even diagonal because the PSF will not
-in general be round.  An example that illustrates the use of this PSF
-is the following: It implies that a star of flux $S_s$ at focal-plane
-position $\pos_s$ will lead to an image (PSF-convolved intensity map)
-of the form
-\begin{eqnarray}\displaystyle
-I(\pos\given\star,S_s,\pos_s) &=& \sum_{k=1}^K S_s\,p_k\,\normal(\pos\given\pos_s+\mean_k,\var_k)
-\quad .
-\end{eqnarray}
-That is, when the PSF is represented as a mixture of Gaussians, any
-image of a star---or indeed any image of any set of stars---is also
-represented as a mixture of Gaussians.
-
-Applying this PSF to an exponential or de~Vaucouleurs galaxy is
-slightly more complicated, because the galaxy has not just a flux
-$S_g$ and a central position $\pos_g$; it also has a shape.  Because
-we are only considering these simple galaxies, we are only permitting
-ellipsoidal shapes, which can be represented by a semi-major axis $a$,
-a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
-eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or by a symmetric
-variance tensor $\var_g$ (which is a general representation of an
-ellipse).  The galaxy is distorted by this variance tensor
-\emph{prior} to PSF convolution, so the focal-plane image
-(PSF-convolved intensity field) for a general exponential galaxy is
-given by
-\begin{eqnarray}\displaystyle
-I(\pos\given\exp,S_g,\pos_g,\var_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_g} S_g\,g_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gn}+\var_k)
-\\
-\var_{gn} &\equiv& \affine_g\cdot\Gm_m\cdot\transpose{\affine_g}
-\\
-\var_g &\equiv& \affine_g\cdot\transpose{\affine_g}
-\\
-\affine_g &=& \left[a\,\uv_1 , b\,\uv_2 \right]
-\quad ,
-\end{eqnarray}
-where $a$ and $b$ are the major and minor axis lengths of the galaxy
-ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
-eigenvectors on the sky pointing in the major-axis and minor-axis
-directions respectively.  Since implicitly all vectors are
-two-dimensional column vectors, this makes $\affine_g$ a $2\times 2$
-affine transformation matrix and $\var_g$ a symmetric $2\times 2$
-variance tensor.  Similarly, the general de~Vaucouleurs galaxy is
-given by
-\begin{eqnarray}\displaystyle
-I(\pos\given\dev,S_h,\pos_h,\var_h) &=& \sum_{k=1}^K \sum_{m=1}^{M_h} S_h\,h_m\,p_k\,\normal(\pos\given\pos_h+\mean_k,\var_{hn}+\var_k)
-\\
-\var_{hn} &\equiv& \affine_h\cdot\Hm_m\cdot\transpose{\affine_h}
-\\
-\var_h &\equiv& \affine_h\cdot\transpose{\affine_h}
-\quad ,
-\end{eqnarray}
-where the matrix mathematics is the same as that in the exponential
-case.  Note the important and key result of this \documentname, to
-wit, that a mixture-of-Gaussian galaxy model (with $M$ components)
-convolved with a mixture-of-Gaussian PSF model (with $K$ components)
-yields a mixture-of-Gaussian image (with $[M\,K]$ components).
-
-...pegagogical figure demonstrating these operations and their results...
-
-...what would you want to do to generalize to Sersic profiles?...
-
-...what about mixtures of exponentials and de~Vaucouleurs:  No problem!...
+This paper is a stub.  You can edit it and help \TheTractor, so long
+as you are either Hogg or Lang!
 

### 2012-08-03 data/visualize_jd_maps.py
start visualization of JD dust map

@@ -1,0 +1,18 @@
+# This file is part of The Tractor.
+
+if __name__ == "__main__":
+    import matplotlib
+    matplotlib.use('Agg')
+    from matplotlib import rc
+    rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':12})
+    rc('text', usetex=True)
+import pylab as plt
+import pyfits
+
+if __name__ == "__main__":
+    jdimage = pyfits.open("jd/brick15.medianebv.fits")[0].data
+    print jdimage.shape
+    plt.gray()
+    plt.clf()
+    plt.imshow(jdimage)
+    plt.savefig("jd_brick15.png")

### 2012-08-03 data/visualize_jd_maps.py
try to read pickle; failing

@@ -8,11 +8,20 @@ if __name__ == "__main__":
     rc('text', usetex=True)
 import pylab as plt
 import pyfits
+import cPickle as pickle
 
 if __name__ == "__main__":
+    fd = open("dstn/herschel-100-g100b.pickle")
+    tractor = pickle.load(fd)
+    fd.close()
+    print tractor
+
+if False:
     jdimage = pyfits.open("jd/brick15.medianebv.fits")[0].data
     print jdimage.shape
+    print jdimage.min()
+    print jdimage.max()
     plt.gray()
     plt.clf()
-    plt.imshow(jdimage)
+    plt.imshow(jdimage, interpolation="nearest", vmin=0., vmax=jdimage.max())
     plt.savefig("jd_brick15.png")

### 2012-08-04 documents/herschel_dust.tex
text and equations about the model

@@
 \newcommand{\TheTractor}{\project{The~Tractor}}
-\newcommand{\sdss}{\project{SDSS}}
 \newcommand{\Herschel}{\project{Herschel}}
+\newcommand{\PACS}{\project{PACS}}
+\newcommand{\SPIRE}{\project{SPIRE}}
+\newcommand{\PHAT}{\project{PHAT}}
 
@@
 \title{High angular resolution dust density and temperature maps from \Herschel\ imaging data}
-\author{Lang, Hogg, etc}
+\author{
+  David~W.~Hogg\altaffilmark{1,2,3},
+  Dustin~Lang\altaffilmark{4,5},
+  Brent~Groves\altaffilmark{3},
+  Hans-Walter~Rix\altaffilmark{3},
+  Karin~Sandstrom\altaffilmark{3}}
 \altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
-\altaffiltext{2}{New York University}
+\altaffiltext{2}{Center for Cosmology and Particle Physics, New York University}
 \altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
 \altaffiltext{4}{Princeton University Observatory}
+\altaffiltext{5}{Carnegie Mellon University}
 
@@
 
-This paper is a stub.  You can edit it and help \TheTractor, so long
-as you are either Hogg or Lang!
+How do you obtain reliable multi-wavelength information about a source
+or a set of sources when your imaging data span a huge range of
+angular scales?
 
-\acknowledgements
-It is a pleasure to thank...
+\section{model}
+
+The model consists of a \emph{physical emission model}, a
+\emph{spatial model}, a set of \emph{priors} and a \emph{likelihood
+  function}.  The physical emission model we choose is a toy model for
+emitting dust grains with a range of temperatures and sizes.  This toy
+model is that a tiny unresolved blob of dust $\Omega_k$ at temperature
+$T_k$ will emit flux density $F_{\nu,k}$ (energy per unit time per
+unit area per unit frequency)
+\begin{eqnarray}\displaystyle
+F_{\nu,k} &=& \Omega_k\,B_{\nu}(T_k)\,\left[\frac{\lambda}{\lambda_0}\right]^{-\beta_k}
+\quad,
+\end{eqnarray}
+where $B_{\nu}(T)$ is the black-body intensity (energy per unit time
+per unit area per unit solid angle per unit frequency), and $\beta_k$
+is an emission parameter accounting for the radiative and
+multi-temperature properties of the dust.  Usually the emission
+parameter $\beta$ is around 2.  In this formulation, the dust
+amplitude $\Omega_k$ has units of solid angle, and ought to be related
+(perhaps not trivially) to the literal solid angle subtended by the
+radiation-absorbing and emitting dust particles.  This model is
+obviously a toy or effective model; much more sophisticated models
+exist (CITATIONS).
+
+The spatial model we choose is a regular grid (on a small
+tangent-plane projection of the the celestial sphere) of
+delta-function pixels.  This is effectively a non-parametric model (in
+the sense that it has an enormous number of parameters).  To meet the
+goals of this project, this grid is made as fine (in an angular sense)
+as the pixel grid of the highest-resolution \Herschel\ \PACS\ images.
+At each two-dimensional pixel location $\theta_k$ on the sky, there is
+a dust solid angle $\Omega_k$, a temperature $T_k$, and an emission
+parameter $\beta_k$.  For any real image taken at some effective
+wavelength $\lambda$ there is an effective point-spread function
+$\psi(\Delta\theta\given\lambda)$ with which the grid of emitting
+pixels is convolved.  Thus the model intensity for an image is
+\begin{eqnarray}\displaystyle
+I_{\nu}(\theta,\lambda) &=& \sum_{k=1}^{K}
+F_{\nu,k}\,\psi(\theta-\theta_k\given\lambda) \quad ,
+\end{eqnarray}
+where the convolution is trivial because the unconvolved intensity
+model is a grid of delta functions.
+
+The likelihood function---the probability of the data given the model parameters---is just the Gaussian
+likelihood function
+\begin{eqnarray}\displaystyle
+-2\,\ln L &\approx& \chi^2
+\\
+\chi^2 &\equiv& \sum_{n=1}^N \frac{\left[d_n - I_{\nu}(\theta_n,\lambda_n)\right]^2}{\sigma_n^2}
+\quad,
+\end{eqnarray}
+where the ``$\approx$'' symbol is used as a reminder that this
+function is not properly normalized, each pixel $n$ has a data value
+$d_n$ that is assumed to be a properly calibrated intensity, and each
+pixel is at a known sky location $\theta_n$, taken in a bandpass at
+effective wavelengths $\lambda_n$, and with an independent Gaussian
+noise variance $\sigma_n^2$.
+
+In non-parametric models like this---even when we only want point
+estimates or optima of an objective function---we cannot avoid using
+prior information to regularize the fit.  The priors we choose are
+``smoothness'' priors inspired by Gaussian processes.  They penalize
+the sum of squares of differences between neighboring pixels.
+\begin{eqnarray}\displaystyle
+-2\,\ln p(\Omega_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_\Omega\,\left[\Omega_k - \Omega_{k'}\right]^2
+\\
+-2\,\ln p(T_k) &\approx& \sum_{\mathrm{neighbors}~k'} \epsilon_T\,\left[T_k - T_{k'}\right]^2
+\\
+-2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta - 2\right]^2
+                    + \sum_{\mathrm{neighbors}~k'} \epsilon_\beta\,\left[\beta_k - \beta_{k'}\right]^2
+\\
+\alpha &\equiv& \left\{\epsilon_\Omega, \epsilon_T, \epsilon_2, \epsilon_\beta\right\}
+\quad ,
+\end{eqnarray}
+where again the ``$\approx$'' symbols are used as a reminder that
+these priors are not properly normalized as written, the sum over
+``neighbors'' $k'$ is a sum, for each pixel, over its four nearest
+neighbors, the prior ``strength'' or information is set by the list
+$\alpha$ of hyper-parameters $\{\epsilon_\Omega, \epsilon_T,
+\epsilon_2, \epsilon_\beta\}$, and the emission parameter grid
+$\beta_k$ is given the additional prior that we expect values around
+2.
+
+...The hyperparameters are set by cross-validation...
+
+\section{experiments and results}
+
+...(in detail, sky)...
+
+...comparison to \PHAT\ extinction maps...
+
+\section{discussion}
+
+...Successes.
+
+...Limitations of the emission model.
+
+...Limitations of the spatial model, especially the PSFs.
+
+...Limitations of the priors.
+
+\acknowledgements It is a pleasure to thank Thomas Henning (MPIA) and
+Ben Weiner (Arizona).  This project made use of the NASA
+\project{Astrophysics Data System}, the Python \project{numpy},
+\project{scipy}, and \project{matplotlib} open-source software
+projects, and the \project{Astrometry.net} codebase.  All the code and
+documentation used in this project is available as part of
+\TheTractor\ at \url{http://TheTractor.org/}.
 

### 2012-08-04 data/visualize_jd_maps.py
kg image too

@@ -6,22 +6,21 @@ if __name__ == "__main__":
     from matplotlib import rc
     rc('font',**{'family':'serif','serif':'Computer Modern Roman','size':12})
     rc('text', usetex=True)
+import numpy as np
 import pylab as plt
 import pyfits
 import cPickle as pickle
 
 if __name__ == "__main__":
-    fd = open("dstn/herschel-100-g100b.pickle")
-    tractor = pickle.load(fd)
-    fd.close()
-    print tractor
+    kgimage = pyfits.open("kg/av_map_b12.fits")[0].data
+    vmax = kgimage[np.where(np.isfinite(kgimage))].max()
+    plt.gray()
+    plt.clf()
+    plt.imshow(kgimage, interpolation="nearest", vmin=0., vmax=vmax)
+    plt.savefig("kg_b12.png")
 
-if False:
     jdimage = pyfits.open("jd/brick15.medianebv.fits")[0].data
-    print jdimage.shape
-    print jdimage.min()
-    print jdimage.max()
     plt.gray()
     plt.clf()
     plt.imshow(jdimage, interpolation="nearest", vmin=0., vmax=jdimage.max())
-    plt.savefig("jd_brick15.png")
+    plt.savefig("jd_b15.png")

### 2012-08-05 documents/herschel_dust.tex
tiny typo
cross validation explained

@@
 \author{
+  Dustin~Lang\altaffilmark{4,5},
   David~W.~Hogg\altaffilmark{1,2,3},
-  Dustin~Lang\altaffilmark{4,5},
   Brent~Groves\altaffilmark{3},
@@
   Karin~Sandstrom\altaffilmark{3}}
-\altaffiltext{1}{To whom correspondence should be addressed; \texttt{foo@bar.edu}}
+\altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
 \altaffiltext{2}{Center for Cosmology and Particle Physics, New York University}
@@
 estimates or optima of an objective function---we cannot avoid using
-prior information to regularize the fit.  The priors we choose are
-``smoothness'' priors inspired by Gaussian processes.  They penalize
-the sum of squares of differences between neighboring pixels.
+prior information to regularize the fit.  There are no extremely good
+priors to put on images, and certainly not images of interstellar
+dust.  For this reason we choose simply ``smoothness'' priors inspired
+by Gaussian processes.  They penalize the sum of squares of
+differences between neighboring pixels.
 \begin{eqnarray}\displaystyle
@@
 \\
--2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta - 2\right]^2
+-2\,\ln p(\beta_k) &\approx& \epsilon_2\,\left[\beta_k - 2\right]^2
                     + \sum_{\mathrm{neighbors}~k'} \epsilon_\beta\,\left[\beta_k - \beta_{k'}\right]^2
@@
 
-...The hyperparameters are set by cross-validation...
+The posterior probability is proportional to the likelihood times the
+prior; its logarithm is sum sum of the log likelihood and the log
+prior.  Dust map inference---the determination of the solid-angle
+parameters $\Omega_k$, the temperature parameters $T_k$, and the
+emissivity parameters $\beta_k$ proceeds by optimization of the log
+posterior probability.  The parameters so derived---maximum \foreign{a
+  posteriori} or MAP---are taken to be the ``best-fit'' parameters in
+what follows.  The MAP dust-map parameters depend on the data, the
+spatial grid $\theta_k$, the data $d_n$, and the data noise variances
+$\sigma_n$.
+
+The hyperparameters $\alpha$---parameters of the priors---set the
+strictness or information content of the priors.  We want to give
+these hyperparameters values that have been dervied objectively from
+the data, so that the collection of \emph{all} data informs the prior
+probability for the parameters $\Omega_k, T_k, \beta_k$ affecting any
+small part of the data.  The best methods for such hyperparameter
+inference is hierarchical Bayesian inference, which is too
+computationally costly for the present effort.  The approximation to
+hierarchical inference adopted here is that of leave-some-out
+cross-validation: We drop randomly chosen small subsets of the data,
+perform the inference of the dust-map parameters $\Omega_k, T_k,
+\beta_k$ using the remaining (the non-left-out) data, predict the
+left-out data, and base the hyperparameter settings on the quality of
+the predictions.
+
+In detail, this works by choosing a random integer $1\leq s\leq S$
+(where $S$ is the number of leave-some-out trials, $S=8$ in what
+follows) for every data pixel $n$.  For each value of $s$---say
+3---MAP values for the dust-map parameters are obtained given all the
+data \emph{except} that marked with integer $s=3$.  The data marked
+with integer $s=3$ are predicted by this model, and a prediction
+$\chi^2$ value is computed for the left-out $s=3$ data.  This is
+repeated for all values of $s$ and all 8 prediction $\chi^2$ values
+are summed.  This total prediction leave-some-out $\chi^2$ is a
+function of the hyperparameters; the hyperparameters $\alpha$ at which
+it is minimized are chosen as the best hyperparameters.
+
+One extremely valuable side-effect of the leave-one-out trials is that
+they also provide uncertainty estimates on the dust-map parameters.
+The leave-some-out trials are effectively also a set of jackknife
+trials.  The variance (mean squared deviation) in the map parameters
+across the $S$ leave-some-out trials is multiplied by (HOGG: FORMULA
+HERE) to give a conservative empirical (bootstrap-like) estimate of
+the variance (mean squared error) in the map parameters.
 

### 2012-08-05 documents/herschel_dust.tex
small tweaks

@@
 
-In detail, this works by choosing a random integer $1\leq s\leq S$
-(where $S$ is the number of leave-some-out trials, $S=8$ in what
-follows) for every data pixel $n$.  For each value of $s$---say
+In detail, this works by choosing a random integer $s$ with $1\leq
+s\leq S$ (where $S$ is the number of leave-some-out trials, $S=8$ in
+what follows) for every data pixel $n$.  For each value of $s$---say
 3---MAP values for the dust-map parameters are obtained given all the
@@
 with integer $s=3$ are predicted by this model, and a prediction
-$\chi^2$ value is computed for the left-out $s=3$ data.  This is
-repeated for all values of $s$ and all 8 prediction $\chi^2$ values
-are summed.  This total prediction leave-some-out $\chi^2$ is a
+$\chi_s^2$ value is computed for the left-out $s=3$ data.  This is
+repeated for all values of $s$ and all $S$ prediction $\chi_s^2$
+values are summed.  This total prediction leave-some-out $\chi^2$ is a
 function of the hyperparameters; the hyperparameters $\alpha$ at which
@@
 
-...(in detail, sky)...
+...details of $\theta_k$, small patch, and so-on...
+
+...(in detail, sky level)...
+
+...what happened in the cross-validation...
+
+...results and jackknifes...
 

### 2012-08-05 documents/herschel_dust.tex
bibliography replacement

@@
 small part of the data.  The best methods for such hyperparameter
-inference is hierarchical Bayesian inference, which is too
-computationally costly for the present effort.  The approximation to
-hierarchical inference adopted here is that of leave-some-out
-cross-validation: We drop randomly chosen small subsets of the data,
-perform the inference of the dust-map parameters $\Omega_k, T_k,
-\beta_k$ using the remaining (the non-left-out) data, predict the
-left-out data, and base the hyperparameter settings on the quality of
-the predictions.
+inference is hierarchical Bayesian inference (for example,
+\citealt{eccentricity}, MORE REFERENCES), which is too computationally
+costly for the present effort.  The approximation to hierarchical
+inference adopted here is that of leave-some-out cross-validation: We
+drop randomly chosen small subsets of the data, perform the inference
+of the dust-map parameters $\Omega_k, T_k, \beta_k$ using the
+remaining (the non-left-out) data, predict the left-out data, and base
+the hyperparameter settings on the quality of the predictions.
 
@@
 \begin{thebibliography}{70}
-\bibitem[Bendinelli(1991)]{bendinelli}
-Bendinelli,~O., 1991, \apj, 366, 599
-\bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
-Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
-\bibitem[Bovy \etal(2011a)]{xd}
-Bovy,~J., Hogg,~D.~W., \& Roweis,~S., 2011, Ann. Appl. Stat., 5, 1657
-\bibitem[Bovy \etal(2011b)]{xdqso}
-Bovy,~J., \etal, 2011, \apj, 729, 141
-\bibitem[Bovy \etal(2012)]{xdqsoz}
-Bovy,~J. \etal, 2012, \aj, 749, 41
-\bibitem[Cappellari(2002)]{cappellari}
-Cappellari,~M., 2002, \mnras, 333, 400
-\bibitem[Ciotti \& Bertin(1999)]{ciotti}
-Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
-\bibitem[Emsellem \etal(1994)]{emsellem}
-Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
-\bibitem[Peng \etal(2002)]{galfit}
-Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
+\bibitem[Hogg \etal(2010)]{eccentricity}
+Hogg,~D.~W., Myers,~A.~D., \& Bovy,~J., 2010, \apj, 725, 2166
 \end{thebibliography}

### 2012-09-19 documents/mixture_models.tex
notes to self
references

--- 2012-05-10 docs/mixture_models.tex
@@
 % -----
-% - write zeroth draft
+% - Brewer suggestion about KLD and ack Brewer
+%   - need to discuss this and show results from it
+% - decide which models will make it in
 % - make useful figures
+% - make figures showing use of the models in a real situation
 % - send to Jim Bosch and Kevin Bundy for comments
@@
 have pioneered this in the area of distribution modeling in large
-numbers of dimensions \citep{xd, xdqso, xdqsoz}, where convolution
+numbers of dimensions (\citealt{xd, xdqso, xdqsoz}), where convolution
 occurs because the true or noise-free distribution is convolved with
@@
 particularly useful in data sets in which the PSF itself has also been
-modeled as a mixture of Gaussians, which is not uncommon.  For
-example, the \project{Sloan Digital Sky Survey} (\sdss; [HOGG cite
-  York]) imaging pipelines [HOGG cite Lupton] makes such models for
-every imaging field.
+modeled as a mixture of Gaussians, which is not uncommon (for example,
+the \project{Sloan Digital Sky Survey} imaging pipelines
+make approximate mixture-of-Gaussian PSF models for
+every imaging field; \citealt{lupton}).
 
@@
 galaxy images with mixtures of Gaussians has been done very
-successfully before \citep[for example,][]{bendinelli, emsellem,
-  bendinelli2, cappellari}.  However, the idea behind those projects
+successfully before (for example, \citealt{bendinelli, emsellem,
+  bendinelli2, cappellari}).  However, the idea behind those projects
 was to use the mixture of Gaussians to provide a very free form for
@@
 
-...somehow have to cite GALFIT \citep{galfit} in this section...
+...somehow have to cite GALFIT (\citealt{galfit}) in this section...
 
@@
 case any two-dimensional probability distribution function---can be
-represented as a mixture of Gaussians to arbitrary accuracy (DSTN: It
-would be nice to have a reference here).
+represented as a mixture of Gaussians to arbitrary accuracy; that is
+\begin{eqnarray}
+foo
+\end{eqnarray}
 
@@
 
-...Cite \citep{ciotti} re half-light radius constants...
+...Cite (\citealt{ciotti}) re half-light radius constants...
 
@@
 Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
+\bibitem[Lupton \etal(2001)]{lupton}
+Lupton,~R., Gunn,~J.~E., Ivezic,~Z., Knapp,~G.~R., Kent,~S.~M., \& Yasuda,~N., 2001, ASPC, 238, 269
 \bibitem[Peng \etal(2002)]{galfit}

### 2012-09-19 documents/mixture_models.tex
mixture of gaussians expl; tractor

@@
 modeled as a mixture of Gaussians, which is not uncommon (for example,
-the \project{Sloan Digital Sky Survey} imaging pipelines
+the \project{Sloan Digital Sky Survey} (\sdss) imaging pipelines
 make approximate mixture-of-Gaussian PSF models for
@@
 
-...Limit scope of this document to just what we need for the
-\tractor...
+We need good performance in two-dimensional image fitting because,
+with \thetractor (Lang \etal, forthcoming), we are building a
+comprehensive model of all the imaging data we can; this will have
+models of many millions of galaxies in imaging that contains on the
+order of $10^{13}$ pixels.  We are basing our models on the
+\sdss\ models (\citealt{lupton})...Limit scope of this document to
+just what we need for the \tractor...
 
@@
 \begin{eqnarray}
-foo
+p(\pos) &\approx& \sum_{k=1}^K a_k\,\normal(\pos\given\mean_k,\var_k)
+\\
+1 &=& \sum_{k=1}^K a_k
+\quad ,
 \end{eqnarray}
+where $p(\pos)$ is any probability distribution of a two-dimensional
+quantity $\pos$, the $\approx$ symbol implies approximation, $K$ is
+the number of Gaussians used in the mixture, and the $K$ Gaussians
+have amplitudes $a_k$, means $\mean_k$, and variance tensors $\var_k$.
+The sum-to-one condition ensures that the probability distribution
+approximation is properly normalized.
 

### 2012-09-19 documents/mixture_models.tex
intro sort-of done

@@
 \newcommand{\project}[1]{\textsl{#1}}
-\newcommand{\tractor}{\project{Tractor}}
+\newcommand{\thetractor}{\project{The~Tractor}}
 \newcommand{\sdss}{\project{SDSS}}
@@
 
-We need good performance in two-dimensional image fitting because,
-with \thetractor (Lang \etal, forthcoming), we are building a
-comprehensive model of all the imaging data we can; this will have
-models of many millions of galaxies in imaging that contains on the
-order of $10^{13}$ pixels.  We are basing our models on the
-\sdss\ models (\citealt{lupton})...Limit scope of this document to
-just what we need for the \tractor...
+Along those lines, it is important in any image-modeling situation to
+think of the PSF as the \emph{pixel-convolved} point-spread function.
+Under this choice, synthesis of a pixelized image involves only
+convolution with the PSF and sampling at the pixel centers.  That is,
+image synthesis (returned to at the end of this \documentname)
+consists of PSF convolution (an arithmetic operation on the two
+mixtures of Gaussians) followed by sampling (evaluation of the
+resulting mixture at the pixel centers).
 
-...somehow have to cite GALFIT (\citealt{galfit}) in this section...
+We need good performance in two-dimensional image synthesis (for
+fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
+building a comprehensive model of all the imaging data we can; this
+will have models of many millions of galaxies in imaging that contains
+on the order of $10^{13}$ pixels.  We are basing our models on the
+\sdss\ Catalog galaxy models, which only include exponential,
+de~Vaucouleurs, and composite (mixture of the two) radial profiles.
+In detail, in fact, the \sdss\ Catalog models are small modifications
+of these (\citealt{lupton}); details below.  So our goal here is to
+provide replacements for these models in order to improve the
+performance of galaxy image modeling and analysis software.  Some of
+the models we present have been used previously in the study of XXX
+(\cite{bundy}); they could also be used to speed other image-fitting
+systems, like the very successful \project{GALFIT} (\citealt{galfit}).
 
-...We need to put somewhere Lupton-like comments about sampling the
-pixel-convolved-PSF-convolved scene...
-
-In what follows we will make heavy use of two-dimensional Gaussian or
-Normal distributions, which look like
+Because we are thinking about two-dimensional imaging, we use here
+two-dimensional Gaussian or Normal distributions, which look like
 \begin{eqnarray}\displaystyle
@@
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\approx& \frac{1}{Z_{\exp}}\,\exp(-\frac{|\spos|}{r_{\exp}})
+Q_{\exp}(\spos) &\approx& \frac{1}{Z_{\exp}}\,\exp(-\frac{|\spos|}{\xi_{\exp}})
 \\
@@
 \\
-r_{\exp} &=& X.XXXX
+\xi_{\exp} &=& X.XXXX
 \quad ,
@@
 where $\spos$ is a dimensionless focal-plane position, and $Z_{\exp}$
-and $r_{\exp}$ are parameters that ensure that the profile integrates
+and $\xi_{\exp}$ are parameters that ensure that the profile integrates
 to unit flux and has unit half-light radius.  The position $\spos$ is
@@
 \begin{eqnarray}\displaystyle
-Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\left[\frac{|\spos|}{r_{\dev}}\right]^{1/4})
+Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\left[\frac{|\spos|}{\xi_{\dev}}\right]^{1/4})
 \\
@@
 \\
-r_{\dev} &=& X.XXXX
+\xi_{\dev} &=& X.XXXX
 \\
@@
 \end{eqnarray}
-
 ...Cite (\citealt{ciotti}) re half-light radius constants...
 
-...Lupton profile as different from de~Vaucouleurs...Sersic profiles...
+The Lupton profiles are...
 
-...Discuss the cutoffs and regularization used by PHOTO and adjust $Z$ and $r$ accordingly...
+...Lupton profiles as different from de~Vaucouleurs...Sersic profiles...Discuss the cutoffs and regularization used by PHOTO and adjust $Z$ and $r$ accordingly...
 

### 2012-09-19 code/optimize_mixture_profiles.py
nomenclature

@@ -49,7 +49,7 @@ def hogg_dev(x):
     """
     return np.exp(-7.66924944 * ((x * x)**0.125 - 1.))
 
-def hogg_lup(x):
+def hogg_luv(x):
     """
     One-dimensional Lupton approximation to DeVaucouleurs profile.
 
@@ -58,13 +58,13 @@ def hogg_lup(x):
     """
     inner = 7.
     outer = 8.
-    lup = np.exp(-7.66925 * ((x * x + 0.0004)**0.125 - 1.))
-    lup[x > outer] *= 0.
+    luv = np.exp(-7.66925 * ((x * x + 0.0004)**0.125 - 1.))
+    luv[x > outer] *= 0.
     middle = (x >= inner) * (x <= outer)
-    lup[middle] *= (1. - ((x[middle] - inner) / (outer - inner)) ** 2) ** 2
-    return lup
+    luv[middle] *= (1. - ((x[middle] - inner) / (outer - inner)) ** 2) ** 2
+    return luv
 
-def hogg_lue(x):
+def hogg_lux(x):
     """
     One-dimensional Lupton approximation to exponential profile.
 
@@ -73,11 +73,11 @@ def hogg_lue(x):
     """
     inner = 3.
     outer = 4.
-    lue = np.exp(-1.67835 * (x - 1.))
-    lue[x > outer] *= 0.
+    lux = np.exp(-1.67835 * (x - 1.))
+    lux[x > outer] *= 0.
     middle = (x >= inner) * (x <= outer)
-    lue[middle] *= (1. - ((x[middle] - inner) / (outer - inner)) ** 2) ** 2
-    return lue
+    lux[middle] *= (1. - ((x[middle] - inner) / (outer - inner)) ** 2) ** 2
+    return lux
 
 def hogg_ser2(x):
     """
@@ -114,10 +114,10 @@ def hogg_model(x, model):
         return hogg_exp(x)
     if model == 'dev':
         return hogg_dev(x)
-    if model == 'lup':
-        return hogg_lup(x)
-    if model == 'lue':
-        return hogg_lue(x)
+    if model == 'luv':
+        return hogg_luv(x)
+    if model == 'lux':
+        return hogg_lux(x)
     if model == 'ser2':
         return hogg_ser2(x)
     if model == 'ser3':
@@ -313,8 +313,8 @@ if __name__ == '__main__':
     inputs = [
         ('exp', 8.),
         ('dev', 8.),
[…]

### 2012-09-19 documents/mixture_models.tex
normalization and half-light radii

@@
 % - make figures showing use of the models in a real situation
+% - make sure we are consistent about normalization! (do the profiles sum to unity or are they unity at \xi = 1?)
 % - send to Jim Bosch and Kevin Bundy for comments
@@
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\approx& \frac{1}{Z_{\exp}}\,\exp(-\frac{|\spos|}{\xi_{\exp}})
+Q_{\exp}(\spos) &\approx& \exp(-\alpha_{\exp}\,|\spos|)
 \\
-Z_{\exp} &=& X.XXXX
-\\
-\xi_{\exp} &=& X.XXXX
+\alpha_{\exp} &=& -1.67834699
 \quad ,
 \end{eqnarray}
-where $\spos$ is a dimensionless focal-plane position, and $Z_{\exp}$
-and $\xi_{\exp}$ are parameters that ensure that the profile integrates
-to unit flux and has unit half-light radius.  The position $\spos$ is
-dimensionless because it parameterizes the unit-size and unit-flux
-dimensionless function.  We seek the best (where ``best'' will be
-defined below) $M_g$-Gaussian mixture approximation
+where $\spos$ is a dimensionless focal-plane position, and
+$\alpha_{\exp}$ ensures that the profile has unit half-light radius.
+The position $\spos$ is dimensionless because it parameterizes the
+unit-size and unit-flux dimensionless function.  We seek the best
+(where ``best'' will be defined below) $M_g$-Gaussian mixture
+approximation
 \begin{eqnarray}\displaystyle
 Q_{\exp}(\spos) &=& \sum_{m=1}^{M_g} g_m\,\normal(\spos\given\zero,\Gm_m)
-\\
-1 &=& \sum_{m=1}^{M_g} g_m
 \\
@@
 \begin{eqnarray}\displaystyle
-Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\left[\frac{|\spos|}{\xi_{\dev}}\right]^{1/4})
+Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\alpha_{\dev}\,|\spos|^{1/4})
 \\
-Z_{\dev} &=& X.XXXX
-\\
-\xi_{\dev} &=& X.XXXX
+\alpha_{\dev} &=& -7.66924944
 \\
 Q_{\dev}(\spos) &=& \sum_{m=1}^{M_h} h_m\,\normal(\spos\given\zero,\Hm_m)
-\\
-1 &=& \sum_{m=1}^{M_h} h_m
 \\

### 2012-09-19 code/optimize_mixture_profiles.py
which models to run

@@ -307,7 +307,7 @@ def main(input):
 
 if __name__ == '__main__':
     if True: # use multiprocessing
-        pmap = Pool(6).map
+        pmap = Pool(8).map
     else: # don't use multiprocessing
         pmap = map
     inputs = [
@@ -315,8 +315,8 @@ if __name__ == '__main__':
         ('dev', 8.),
         ('lup', 8.),
         ('lue', 4.),
-        ('ser2', 8.),
-        ('ser3', 8.),
-        ('ser5', 8.),
+        # ('ser2', 8.),
+        # ('ser3', 8.),
+        # ('ser5', 8.),
         ]
     pmap(main, inputs)

### 2012-09-19 documents/mixture_models.tex
citation; index typo
lupton profiles; bundy citation

@@
 % - make sure we are consistent about normalization! (do the profiles sum to unity or are they unity at \xi = 1?)
+% - does the Lupton paper give the profiles, or should we cite the code directly?
 % - send to Jim Bosch and Kevin Bundy for comments
@@
 \newcommand{\var}{\tmatrix{V}\!}
-\newcommand{\Gm}{\tmatrix{G}}
-\newcommand{\Hm}{\tmatrix{H}}
 \newcommand{\affine}{\tmatrix{R}}
@@
 \newcommand{\dev}{\mathrm{dev}}
+\newcommand{\lux}{\mathrm{lux}}
+\newcommand{\luv}{\mathrm{luv}}
 
@@
 performance of galaxy image modeling and analysis software.  Some of
-the models we present have been used previously in the study of XXX
-(\cite{bundy}); they could also be used to speed other image-fitting
-systems, like the very successful \project{GALFIT} (\citealt{galfit}).
+the models we present have been used previously in tools for precise
+photometry (Bundy \etal, forthcoming); they could also be used to
+speed other image-fitting systems, like the very successful
+\project{GALFIT} (\citealt{galfit}).
 
@@
 
-We wish to make an approximation $Q_{\exp}(\cdot)$ to the
-two-dimensional circular exponential profile
+We wish to make an approximation to the two-dimensional circular
+exponential profile $Q_{\exp}(\cdot)$
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\approx& \exp(-\alpha_{\exp}\,|\spos|)
+Q_{\exp}(\spos) &\equiv& \exp(-\alpha_{\exp}\,[|\spos| - 1])
 \\
-\alpha_{\exp} &=& -1.67834699
+\alpha_{\exp} &\equiv& -1.67834699
 \quad ,
@@
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &=& \sum_{m=1}^{M_g} g_m\,\normal(\spos\given\zero,\Gm_m)
+Q_{\exp}(\spos) &\approx& \sum_{m=1}^{M_{\exp}} a_{\exp,m}\,\normal(\spos\given\zero,\var_{\exp,m})
 \\
-\Gm_m &=& \sigma^2_{\exp,m}\,\identity
+\var_{\exp,m} &\equiv& \sigma^2_{\exp,m}\,\identity
 \quad ,
@@
 where all of the means are exactly zero and all of the variances
-$\Gm_m$ in the mixture are scalar multiples of the identity matrix
+$\var_m$ in the mixture are scalar multiples of the identity matrix
 $\identity$ because we are requiring this dimensionless function to be
@@
 \begin{eqnarray}\displaystyle
-Q_{\dev}(\spos) &\approx& \frac{1}{Z_{\dev}}\,\exp(-\alpha_{\dev}\,|\spos|^{1/4})
+Q_{\dev}(\spos) &\equiv& \exp(-\alpha_{\dev}\,[|\spos|^{1/4} - 1])
 \\
-\alpha_{\dev} &=& -7.66924944
+\alpha_{\dev} &\equiv& -7.66924944
 \\
-Q_{\dev}(\spos) &=& \sum_{m=1}^{M_h} h_m\,\normal(\spos\given\zero,\Hm_m)
+Q_{\dev}(\spos) &\approx& \sum_{m=1}^{M_{\dev}} a_{\dev,m}\,\normal(\spos\given\zero,\var_{\dev,m})
 \\
-\Hm_m &=& \sigma^2_{\dev,m}\,\identity
+\var_{\dev,m} &\equiv& \sigma^2_{\dev,m}\,\identity
 \quad .
 \end{eqnarray}
-...Cite (\citealt{ciotti}) re half-light radius constants...
+(The half-light radius parameters $\alpha_{\exp}$ and $\alpha_{\dev}$
+are taken from \citealt{ciotti}.)  The challenge we meet below is to
+determine the parameters
+\begin{eqnarray}
+\{a_{\exp,m},\sigma_{\exp,m}\}_{m=1}^{M_{\exp}} \mbox{~and~} \{a_{\dev,m},\sigma_{\dev,m}\}_{m=1}^{M_{\dev}}
+\end{eqnarray}
+to best approximate the
+traditional galaxy profile functions, under some sensible definiton of
+the word ``best'', as a function of the model complexity parameters
+(numbers of components) $(M_{\exp}, M_{\dev})$.
 
-The Lupton profiles are...
-
-...Lupton profiles as different from de~Vaucouleurs...Sersic profiles...Discuss the cutoffs and regularization used by PHOTO and adjust $Z$ and $r$ accordingly...
+The \sdss\ pipelines make use of modified profiles, which have been
+truncated smoothly at large radius and (in the case of the
+de~Vaucouleurs profile) ``softened'' at the center (\citealt{lupton}).
+The \sdss\ form of the exponential profile is
+\begin{eqnarray}\displaystyle
+Q_{\lux}(\spos) &\equiv& \left\{\begin{array}{ll}
+  \exp(-\alpha_{\lux}\,[|\spos| - 1]) & \mbox{for~}|\spos| < 3 \\
+  \exp(-\alpha_{\lux}\,[|\spos| - 1])
+  \,\left[1 - [|\spos| - 3]^2\right]^2 & \mbox{for~}3 < |\spos| < 4 \\
+  0                                   & \mbox{for~}4 < |\spos|
+\end{array}\right.
+\\
+\alpha_{\lux} &\equiv& -1.67835
+\quad ,
+\end{eqnarray}
+and the \sdss\ form of the de~Vaucouleurs profile is
+\begin{eqnarray}\displaystyle
+Q_{\luv}(\spos) &\equiv& \left\{\begin{array}{ll}
+  \exp(-\alpha_{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right]) & \mbox{for~}|\spos| < 7 \\
+  \exp(-\alpha_{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right])
+  \,\left[1 - [|\spos| - 7]^2\right]^2 & \mbox{for~}7 < |\spos| < 8 \\
+  0                                   & \mbox{for~}8 < |\spos|
+\end{array}\right.
+\\
+\alpha_{\luv} &\equiv& -7.66925
+\quad .
+\end{eqnarray}
+(The half-light radius parameters $\alpha_{\exp}$ and
+$\alpha_{\dev}$---and the softening and cutoff radius parameters---are
+taken from the \sdss\ codebase.)  The \sdss\ radial profiles are much
+better behaved than the standard forms; we recommend them strongly.
+Once again, the challenge is to determine the best parameters
+\begin{eqnarray}
+\{a_{\lux,m},\sigma_{\lux,m}\}_{m=1}^{M_{\lux}} \mbox{~and~} \{a_{\luv,m},\sigma_{\luv,m}\}_{m=1}^{M_{\luv}}
+\quad .
+\end{eqnarray}
 
@@
 \begin{eqnarray}\displaystyle
-I(\pos\given\exp,S_g,\pos_g,\var_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_g} S_g\,g_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gn}+\var_k)
+I(\pos\given\exp,S_g,\pos_g,\var_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_g} S_g\,g_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
 \\
-\var_{gn} &\equiv& \affine_g\cdot\Gm_m\cdot\transpose{\affine_g}
+\var_{gm} &\equiv& \affine_g\cdot\var_m\cdot\transpose{\affine_g}
 \\
@@
 \\
-\var_{hn} &\equiv& \affine_h\cdot\Hm_m\cdot\transpose{\affine_h}
+\var_{hn} &\equiv& \affine_h\cdot\var_m\cdot\transpose{\affine_h}
 \\

### 2012-09-19 documents/mixture_models.tex
small tweaks

@@
 fitting) because, with \thetractor\ (Lang \etal, forthcoming), we are
-building a comprehensive model of all the imaging data we can; this
+building a comprehensive model of all the imaging data we have; this
 will have models of many millions of galaxies in imaging that contains
@@
 The position $\spos$ is dimensionless because it parameterizes the
-unit-size and unit-flux dimensionless function.  We seek the best
-(where ``best'' will be defined below) $M_g$-Gaussian mixture
-approximation
+unit-size dimensionless function.  We seek the best (where ``best''
+will be defined below) $M_{\exp}$-Gaussian mixture approximation
 \begin{eqnarray}\displaystyle
@@
 where all of the means are exactly zero and all of the variances
-$\var_m$ in the mixture are scalar multiples of the identity matrix
-$\identity$ because we are requiring this dimensionless function to be
-precisely circular (so every component is itself circular and
-concentric).  Similarly for the de~Vaucouleurs profile
+$\var_{\exp,m}$ in the mixture are scalar multiples of the identity
+matrix $\identity$ because we are requiring this dimensionless
+function to be precisely circular (so every component is itself
+circular and concentric).  Similarly for the de~Vaucouleurs profile
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-to best approximate the
-traditional galaxy profile functions, under some sensible definiton of
-the word ``best'', as a function of the model complexity parameters
-(numbers of components) $(M_{\exp}, M_{\dev})$.
+to best approximate the traditional galaxy profile functions, under
+some sensible definiton of the word ``best'', as a function of the
+model complexity parameters (numbers of components) $(M_{\exp},
+M_{\dev})$.
+
+...In addition to these there are general Sersi\'c profiles, of which
+the exponential and de~Vaucouleurs profiles are special cases...
 
@@
 $\alpha_{\dev}$---and the softening and cutoff radius parameters---are
-taken from the \sdss\ codebase.)  The \sdss\ radial profiles are much
-better behaved than the standard forms; we recommend them strongly.
-Once again, the challenge is to determine the best parameters
-\begin{eqnarray}
-\{a_{\lux,m},\sigma_{\lux,m}\}_{m=1}^{M_{\lux}} \mbox{~and~} \{a_{\luv,m},\sigma_{\luv,m}\}_{m=1}^{M_{\luv}}
-\quad .
-\end{eqnarray}
+taken from the \sdss\ codebase.)
 

### 2012-09-20 documents/mixture_models.tex
clarification, simplification

@@
 in general be round.  An example that illustrates the use of this PSF
-is the following: It implies that a star of flux $S_s$ at focal-plane
-position $\pos_s$ will lead to an image (PSF-convolved intensity map)
-of the form
+is the following: A star of flux $S_s$ at focal-plane position
+$\pos_s$ will lead to an image (PSF-convolved intensity map) of the
+form
 \begin{eqnarray}\displaystyle
@@
 \begin{eqnarray}\displaystyle
-I(\pos\given\exp,S_g,\pos_g,\var_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_g} S_g\,g_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
+I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_{\exp}} S_g\,a_{\exp,m}\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
 \\
-\var_{gm} &\equiv& \affine_g\cdot\var_m\cdot\transpose{\affine_g}
-\\
-\var_g &\equiv& \affine_g\cdot\transpose{\affine_g}
+\var_{gm} &\equiv& \affine_g\cdot\var_{\exp,m}\cdot\transpose{\affine_g}
 \\
@@
 eigenvectors on the sky pointing in the major-axis and minor-axis
-directions respectively.  Since implicitly all vectors are
-two-dimensional column vectors, this makes $\affine_g$ a $2\times 2$
-affine transformation matrix and $\var_g$ a symmetric $2\times 2$
-variance tensor.  Similarly, the general de~Vaucouleurs galaxy is
-given by
-\begin{eqnarray}\displaystyle
-I(\pos\given\dev,S_h,\pos_h,\var_h) &=& \sum_{k=1}^K \sum_{m=1}^{M_h} S_h\,h_m\,p_k\,\normal(\pos\given\pos_h+\mean_k,\var_{hn}+\var_k)
-\\
-\var_{hn} &\equiv& \affine_h\cdot\var_m\cdot\transpose{\affine_h}
-\\
-\var_h &\equiv& \affine_h\cdot\transpose{\affine_h}
-\quad ,
-\end{eqnarray}
-where the matrix mathematics is the same as that in the exponential
-case.  Note the important and key result of this \documentname, to
-wit, that a mixture-of-Gaussian galaxy model (with $M$ components)
-convolved with a mixture-of-Gaussian PSF model (with $K$ components)
-yields a mixture-of-Gaussian image (with $[M\,K]$ components).
+directions respectively.  Implicitly all vectors are two-dimensional
+column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
+matrix that contains the ``shape'' (position angle, major-axis, and
+ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
+$\lux$ cases are all similar.  Note the important and key result of
+this \documentname, to wit, that a mixture-of-Gaussian galaxy model
+(with $M$ components) convolved with a mixture-of-Gaussian PSF model
+(with $K$ components) yields a mixture-of-Gaussian image (with
+$[M\,K]$ components).
 

### 2012-09-20 documents/mixture_models.tex
sign errors

@@
 \newcommand{\dev}{\mathrm{dev}}
+\newcommand{\ser}{\mathrm{ser}}
 \newcommand{\lux}{\mathrm{lux}}
@@
 image fitting; we are using them ourselves.  We discuss extensions to
-Sersic models and applications of these ideas to three-dimensional
+S\'ersic models and applications of these ideas to three-dimensional
 deprojection.
@@
 \\
-\alpha_{\exp} &\equiv& -1.67834699
+\alpha_{\exp} &\equiv& 1.67834699
 \quad ,
@@
 \\
-\alpha_{\dev} &\equiv& -7.66924944
+\alpha_{\dev} &\equiv& 7.66924944
 \\
@@
 \begin{eqnarray}
-\{a_{\exp,m},\sigma_{\exp,m}\}_{m=1}^{M_{\exp}} \mbox{~and~} \{a_{\dev,m},\sigma_{\dev,m}\}_{m=1}^{M_{\dev}}
+\{a_{\exp,m},\sigma^2_{\exp,m}\}_{m=1}^{M_{\exp}}~,~\{a_{\dev,m},\sigma^2_{\dev,m}\}_{m=1}^{M_{\dev}}
 \end{eqnarray}
@@
 
-...In addition to these there are general Sersi\'c profiles, of which
-the exponential and de~Vaucouleurs profiles are special cases...
+In addition to these, there are general S\'ersic profiles, of which
+the exponential and de~Vaucouleurs profiles are special cases.  The
+general S\'ersic profile has one parameter (the ``index'') $n$:
+\begin{eqnarray}\displaystyle
+Q_{\ser}(\spos\given n) &\equiv& \exp(-\alpha_{\ser}(n)\,[|\spos|^{1/n} - 1])
+\\
+\{\alpha_{\ser}(2), \alpha_{\ser}(3), \alpha_{\ser}(5)\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
+\quad ,
+\end{eqnarray}
+where we have given the constant $\alpha_{\ser}(n)$ for just a few
+values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles give
+the $n=1$ and $n=4$ values).
 
@@
 \\
-\alpha_{\lux} &\equiv& -1.67835
+\alpha_{\lux} &\equiv& 1.67835
 \quad ,
@@
 \\
-\alpha_{\luv} &\equiv& -7.66925
+\alpha_{\luv} &\equiv& 7.66925
 \quad .
@@
 
-...what would you want to do to generalize to Sersic profiles?...
+...what would you want to do to generalize to S\'ersic profiles?...
 

### 2012-09-20 documents/mixture_models.tex
objective function and optimization

@@
 
-...If we wanted to approximate these models, what objective function
-would we use?  What tolerances are okay?...Define a sensible scalar
-objective...Give fitting procedure...give actual values for the
-parameters...
+We seek the best mixture-of-Gaussian approximations.  This
+necessitates definition of the word ``best''.  If we think of the
+profiles as being two-diensional probability distribution functions
+(for, say, the arrivals of photons), then one natural choice is the
+K-L divergence or similar cross-entropy or information-theoretic
+measure.  However, in typical astronomical imaging, the galaxy is
+superimposed on a substantial, flat sky level, and the noise in the
+data is close to Gaussian.  This suggests more chi-squared-like
+objectives.  We adopt the latter, in part because it is most
+appropriate for our specific proposed application (modeling \sdss-like
+astronomical imaging), but experiments we have performed suggest that
+information-theoretic objectives also lead to good results.
+
+In detail, the chi-squared objective we minimize---the
+\emph{badness}---is a squared residual between function and
+approximation.  It is designed to be equivalent to a chi-squared
+statistic in a homoskedastic two-dimensional image of the profile
+taken with extremely high angular resolution (pixels of size 0.001 the
+half-light radius) and vanishing point-spread function.  The badness
+is computed in one dimension but weighted in radius (weight increasing
+linearly with radius) to make it equivalent to the two-dimensional
+chi-squared.  Optimization is performed by the \project{scipy}
+implementation of the BFGS algorithm, with many initializations to
+explore multiple local minima.  Further details are available in the
+code, which is publicly
+available.\footnote{\url{https://github.com/davidwhogg/TheTractor/}}
 
 ...Show plots demonstrating quality of fit, one-d and two-d plots...
+
+...give tables of results, including Bundy versions and recommended versions...
+
+...show dependence of amplitudes and variances on sersic n for some M...
 
@@
 a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
-eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or by a symmetric
-variance tensor $\var_g$ (which is a general representation of an
-ellipse).  The galaxy is distorted by this variance tensor
-\emph{prior} to PSF convolution, so the focal-plane image
-(PSF-convolved intensity field) for a general exponential galaxy is
-given by
+eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
+an affine transformation $\affine_g$ that takes a circle to the
+relevant ellipse (and is a therefore a general representation of an
+ellipse; it is also the matrix square root of the symmetric variance
+tensor describing the ellipse).  The galaxy is distorted by this
+affine transformation \emph{prior} to PSF convolution, so the
+focal-plane image (PSF-convolved intensity field) for a general (say)
+$\exp$ galaxy is given by
 \begin{eqnarray}\displaystyle
@@
 
-...what would you want to do to generalize to S\'ersic profiles?...
-
 ...what about mixtures of exponentials and de~Vaucouleurs:  No problem!...
 
-\acknowledgements
-It is a pleasure to thank...
+\acknowledgements It is a pleasure to thank Brendon Brewer (Auckland)
+and Kevin Bundy (IPMU) for valuable comments.  This work was supported
+in part by NASA (grant NNX12AI50G) and the NSF (grant IIS-1124794).
 

### 2012-09-20 documents/mixture_models.tex
wording tweaks

@@
 % -----
-% - Brewer suggestion about KLD and ack Brewer
-%   - need to discuss this and show results from it
 % - decide which models will make it in
@@
 
-We seek the best mixture-of-Gaussian approximations.  This
+We seek the \emph{best} mixture-of-Gaussian approximations.  This
 necessitates definition of the word ``best''.  If we think of the
-profiles as being two-diensional probability distribution functions
+profiles as being two-dimensional probability distribution functions
 (for, say, the arrivals of photons), then one natural choice is the
@@
 data is close to Gaussian.  This suggests more chi-squared-like
-objectives.  We adopt the latter, in part because it is most
+objectives.  We adopt the latter, in part because they are most
 appropriate for our specific proposed application (modeling \sdss-like
@@
 In detail, the chi-squared objective we minimize---the
-\emph{badness}---is a squared residual between function and
-approximation.  It is designed to be equivalent to a chi-squared
-statistic in a homoskedastic two-dimensional image of the profile
-taken with extremely high angular resolution (pixels of size 0.001 the
-half-light radius) and vanishing point-spread function.  The badness
-is computed in one dimension but weighted in radius (weight increasing
+\emph{badness}---is a squared residual between the exact profile
+function $Q(\spos)$ and its mixture-of-Gaussian approximation.  It is
+designed to be equivalent to a chi-squared statistic in a
+homoskedastic two-dimensional image of the profile taken with
+extremely high angular resolution (pixels of size 0.001 the half-light
+radius) and vanishing point-spread function.  Quantitatively the
+badness is defined to be the mean squared residual in the $Q(\spos)$
+functions, which are normalized to have unit intensity at the
+half-light radius, averaged over a two-dimensional circular region in
+the $\spos$ plane centered on the (circularly symmetric) profile and
+extending out to radius $\xi_{\max}$.  We use $\xi_{\max}=8$ for all
+profiles except the $\lux$ profile, for which we use $\xi_{\max}=4$.
+In practice, the badness is computed in a one-dimensional numerical
+integral but the integral is weighted in radius (weight increasing
 linearly with radius) to make it equivalent to the two-dimensional
-chi-squared.  Optimization is performed by the \project{scipy}
-implementation of the BFGS algorithm, with many initializations to
-explore multiple local minima.  Further details are available in the
-code, which is publicly
+chi-squared.
+
+Optimization (minimization) of the badness is performed by the
+\project{scipy} implementation of the BFGS algorithm, with many
+initializations to explore multiple local minima.  Further details are
+available in the code, which is publicly
 available.\footnote{\url{https://github.com/davidwhogg/TheTractor/}}
@@
 
-...show dependence of amplitudes and variances on sersic n for some M...
+...show dependence of amplitudes and variances on sersic n for some
+M...be sure to synchronize $\xi_{\max}$...
 

### 2012-09-20 code/optimize_mixture_profiles.py
re-plot without re-fitting

@@ -17,6 +17,7 @@ import numpy as np
 import scipy.optimize as op
 from multiprocessing import Pool
 import cPickle as pickle
+import os as os
 
 def not_normal(x, V):
     """
@@ -268,40 +269,53 @@ def main(input):
     bestbadness = badness
     for K in range(2, 20):
         prefix = '%s_K%02d_MR%02d' % (model, K, int(round(max_radius) + 0.01))
-        print 'working on %s at K = %d (%s)' % (model, K, prefix)
-        newvar = 2.0 * np.max(np.append(var, 1.0))
-        newamp = 1.0 * newvar
-        amp = np.append(newamp, amp)
-        var = np.append(newvar, var)
-        pars = np.append(amp, var)
-        for i in range(2 * K):
-            (badness, pars) = optimize_mixture(K, pars, model, max_radius, log10_squared_deviation, bad_fn)
-            if (badness < bestbadness) or (i == 0):
-                print '%s %d %d improved' % (model, K, i)
-                bestpars = rearrange_pars(pars)
-                bestbadness = badness
-                while bestbadness < 1. and log10_squared_deviation > -5.5:
-                    bestbadness *= 10.
-                    log10_squared_deviation = np.round(log10_squared_deviation - 1.)
-            else:
-                print '%s %d %d not improved' % (model, K, i)
-                amp = 1. * bestpars[0:K]
-                var = 1. * bestpars[K:K+K]
-                var[0] = 2.0 * var[np.mod(i, K)]
-                amp[0] = 0.5 * amp[np.mod(i, K)]
-                pars = np.append(amp, var)
-        lastKbadness = bestbadness
-        pars = rearrange_pars(bestpars)
+        picklefn = prefix + '.pickle'
+        if os.path.exists(picklefn):
+            print 'not working on %s at K = %d (%s); just reading %s' % (model, K, prefix, picklefn)
+            picklefile = open(picklefn, "r")
+            pars = pickle.load(picklefile)
+            picklefile.close()
+            bestbadness = bad_fn(np.log(pars), model, max_radius, log10_squared_deviation)
+            while bestbadness < 1. and log10_squared_deviation > -7.5:
+                bestbadness *= 10.
+                log10_squared_deviation = np.round(log10_squared_deviation - 1.)
+                print "K, bestbadness, log10_squared_deviation", K, bestbadness, log10_squared_deviation
+        else:
+            print 'working on %s at K = %d (%s)' % (model, K, prefix)
[…]

### 2012-09-20 code/optimize_mixture_profiles.py
stop at 12 no matter what

@@ -315,7 +315,7 @@ def main(input):
         txtfile = open(prefix + '.txt', "w")
         txtfile.write("%s\n" % repr(pars))
         txtfile.close()
-        if log10_squared_deviation < -5.5 and K > 11:
+        if K > 11:
             break
     return None
 

### 2012-10-18 documents/mixture_models.tex
put in full figure and table captions, adjust symbology, wording tweaks, normalization paragraph

@@
 % - decide which models will make it in
-% - make useful figures
+% - make pickles_to_latex_table.py function
 % - make figures showing use of the models in a real situation
-% - make sure we are consistent about normalization! (do the profiles sum to unity or are they unity at \xi = 1?)
 % - does the Lupton paper give the profiles, or should we cite the code directly?
@@
 \newcommand{\mean}{\tvector{m}}
-\newcommand{\var}{\tmatrix{V}\!}
+\newcommand{\var}{\tmatrix{V}}
 \newcommand{\affine}{\tmatrix{R}}
@@
 standard unconvolved galaxy models, especially the de~Vaucouleurs and
-S\'ersic models, are very ill-behaved near the galaxy center.
-Rendering these profiles precisely near the center can be very
-challenging numerically.  In addition, most of the rendering time is
-spent at very small radii, where PSF convolution is going to erase all
-structure anyway.  That is, a mixture-of-Gaussians description of the
-de~Vaucouleurs profile saves time both in rendering \emph{and} in PSF
-convolution; it produces profiles that are approximate but very high
-performance when all real uses of the profiles are PSF-convolved, as
-they usually are.
+S\'ersic models (\citealt{dev}, \citealt{ser}), are very ill-behaved
+near the galaxy center.  Rendering these profiles precisely near the
+center can be very challenging numerically.  In addition, most of the
+rendering time is spent at very small radii, where PSF convolution is
+going to erase all structure anyway.  That is, a mixture-of-Gaussians
+description of the de~Vaucouleurs profile saves time both in rendering
+\emph{and} in PSF convolution; it produces profiles that are
+approximate but very high performance when all real uses of the
+profiles are PSF-convolved, as they usually are.
 
@@
 We wish to make an approximation to the two-dimensional circular
-exponential profile $Q_{\exp}(\cdot)$
+exponential (exp) profile $Q^{\exp}(\cdot)$
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\equiv& \exp(-\alpha_{\exp}\,[|\spos| - 1])
+Q^{\exp}(\spos) &\equiv& \exp(-\alpha^{\exp}\,[|\spos| - 1])
 \\
-\alpha_{\exp} &\equiv& 1.67834699
+\alpha^{\exp} &\equiv& 1.67834699
 \quad ,
@@
 where $\spos$ is a dimensionless focal-plane position, and
-$\alpha_{\exp}$ ensures that the profile has unit half-light radius.
-The position $\spos$ is dimensionless because it parameterizes the
-unit-size dimensionless function.  We seek the best (where ``best''
-will be defined below) $M_{\exp}$-Gaussian mixture approximation
+$\alpha^{\exp}$ is a dimensionless inverse length set to ensure that
+the profile has unit half-light radius.  The position $\spos$ is
+dimensionless because it parameterizes the unit-size dimensionless
+function.  We seek the best (where ``best'' will be defined below)
+$M^{\exp}$-Gaussian mixture (where $M^{\exp}$ is an integer)
+approximation
 \begin{eqnarray}\displaystyle
-Q_{\exp}(\spos) &\approx& \sum_{m=1}^{M_{\exp}} a_{\exp,m}\,\normal(\spos\given\zero,\var_{\exp,m})
+Q^{\exp}(\spos) &\approx& \sum_{m=1}^{M^{\exp}} a^{\exp}_m\,\normal(\spos\given\zero,\var^{\exp}_m)
 \\
-\var_{\exp,m} &\equiv& \sigma^2_{\exp,m}\,\identity
+\var^{\exp}_m &\equiv& v^{\exp}_m\,\identity
 \quad ,
@@
 where all of the means are exactly zero and all of the variances
-$\var_{\exp,m}$ in the mixture are scalar multiples of the identity
-matrix $\identity$ because we are requiring this dimensionless
-function to be precisely circular (so every component is itself
-circular and concentric).  Similarly for the de~Vaucouleurs profile
+$\var^{\exp}_m$ in the mixture can be represented as a scalar
+$v^{\exp}_m$ multiplied by the identity matrix $\identity$
+because we are requiring this dimensionless function to be precisely
+circular (so every component is itself circular and concentric).
+Similarly for the de~Vaucouleurs (dev) profile
 \begin{eqnarray}\displaystyle
-Q_{\dev}(\spos) &\equiv& \exp(-\alpha_{\dev}\,[|\spos|^{1/4} - 1])
+Q^{\dev}(\spos) &\equiv& \exp(-\alpha^{\dev}\,[|\spos|^{1/4} - 1])
 \\
-\alpha_{\dev} &\equiv& 7.66924944
+\alpha^{\dev} &\equiv& 7.66924944
 \\
-Q_{\dev}(\spos) &\approx& \sum_{m=1}^{M_{\dev}} a_{\dev,m}\,\normal(\spos\given\zero,\var_{\dev,m})
+Q^{\dev}(\spos) &\approx& \sum_{m=1}^{M^{\dev}} a^{\dev}_m\,\normal(\spos\given\zero,\var^{\dev}_m)
 \\
-\var_{\dev,m} &\equiv& \sigma^2_{\dev,m}\,\identity
+\var^{\dev}_m &\equiv& v^{\dev}_m\,\identity
 \quad .
 \end{eqnarray}
-(The half-light radius parameters $\alpha_{\exp}$ and $\alpha_{\dev}$
-are taken from \citealt{ciotti}.)  The challenge we meet below is to
-determine the parameters
+The half-light inverse-radius parameters $\alpha^{\exp}$ and
+$\alpha^{\dev}$ are from \citet{ciotti}.  The challenge we
+meet below is to determine the parameters
 \begin{eqnarray}
-\{a_{\exp,m},\sigma^2_{\exp,m}\}_{m=1}^{M_{\exp}}~,~\{a_{\dev,m},\sigma^2_{\dev,m}\}_{m=1}^{M_{\dev}}
+\{a^{\exp}_m,v^{\exp}_m\}_{m=1}^{M^{\exp}}~,~\{a^{\dev}_m,v^{\dev}_m\}_{m=1}^{M^{\dev}}
 \end{eqnarray}
@@
 some sensible definiton of the word ``best'', as a function of the
-model complexity parameters (numbers of components) $(M_{\exp},
-M_{\dev})$.
+model complexity parameters (numbers of components) $(M^{\exp},
+M^{\dev})$.
 
-In addition to these, there are general S\'ersic profiles, of which
-the exponential and de~Vaucouleurs profiles are special cases.  The
-general S\'ersic profile has one parameter (the ``index'') $n$:
+In addition to these, there are general S\'ersic (ser) profiles, of
+which the exp and dev profiles are special cases.  The general ser
+profile has one parameter (the ``index'') $n$:
 \begin{eqnarray}\displaystyle
-Q_{\ser}(\spos\given n) &\equiv& \exp(-\alpha_{\ser}(n)\,[|\spos|^{1/n} - 1])
+Q^{\ser}(\spos\given n) &\equiv& \exp(-\alpha^{\ser}(n)\,[|\spos|^{1/n} - 1])
 \\
-\{\alpha_{\ser}(2), \alpha_{\ser}(3), \alpha_{\ser}(5)\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
+\{\alpha^{\ser}(2), \alpha^{\ser}(3), \alpha^{\ser}(5)\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
 \quad ,
 \end{eqnarray}
-where we have given the constant $\alpha_{\ser}(n)$ for just a few
+where we have given the constant $\alpha^{\ser}(n)$ for just a few
 values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles give
@@
 de~Vaucouleurs profile) ``softened'' at the center (\citealt{lupton}).
-The \sdss\ form of the exponential profile is
+The \sdss\ form of the exponential (lux) profile is
 \begin{eqnarray}\displaystyle
-Q_{\lux}(\spos) &\equiv& \left\{\begin{array}{ll}
-  \exp(-\alpha_{\lux}\,[|\spos| - 1]) & \mbox{for~}|\spos| < 3 \\
-  \exp(-\alpha_{\lux}\,[|\spos| - 1])
+Q^{\lux}(\spos) &\equiv& \left\{\begin{array}{ll}
+  \exp(-\alpha^{\lux}\,[|\spos| - 1]) & \mbox{for~}|\spos| < 3 \\
+  \exp(-\alpha^{\lux}\,[|\spos| - 1])
   \,\left[1 - [|\spos| - 3]^2\right]^2 & \mbox{for~}3 < |\spos| < 4 \\
@@
 \\
-\alpha_{\lux} &\equiv& 1.67835
+\alpha^{\lux} &\equiv& 1.67835
 \quad ,
 \end{eqnarray}
-and the \sdss\ form of the de~Vaucouleurs profile is
+and the \sdss\ form of the de~Vaucouleurs (luv) profile is
 \begin{eqnarray}\displaystyle
-Q_{\luv}(\spos) &\equiv& \left\{\begin{array}{ll}
-  \exp(-\alpha_{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right]) & \mbox{for~}|\spos| < 7 \\
-  \exp(-\alpha_{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right])
+Q^{\luv}(\spos) &\equiv& \left\{\begin{array}{ll}
+  \exp(-\alpha^{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right]) & \mbox{for~}|\spos| < 7 \\
+  \exp(-\alpha^{\luv}\,\left[[|\spos|^2 + 0.0004]^{1/8} - 1\right])
   \,\left[1 - [|\spos| - 7]^2\right]^2 & \mbox{for~}7 < |\spos| < 8 \\
@@
 \\
-\alpha_{\luv} &\equiv& 7.66925
+\alpha^{\luv} &\equiv& 7.66925
 \quad .
 \end{eqnarray}
-(The half-light radius parameters $\alpha_{\exp}$ and
-$\alpha_{\dev}$---and the softening and cutoff radius parameters---are
-taken from the \sdss\ codebase.)
+The half-light inverse-radius parameters $\alpha^{\exp}$ and
+$\alpha^{\dev}$---and the softening and cutoff radius parameters---are
+taken from the \sdss\ codebase.
 
-...Note normalization issues and how easy it is to normalize the mixture-of-Gaussian model...
+The profiles above are normalized to have unit intensity
+(approximately) at their half-light radii.  In many cases, the
+investigator wants profiles that are normalized to have unit total
+flux (intensity integrated over solid angle).  Although there is an
+analytic result for the dev profile, numerical integration of the
+concentrated profiles dev and luv to determine total fluxes can be
+challenging.  This is not true for the mixture-of-Gaussian
+approximations: Each Gaussian is normalized, so the sum of the
+amplitudes $\sum_m a^{\luv}_m$ (for the luv profile, say) gives the
+total flux for the mixture-of-Gaussians approximation to that profile.
 
@@
 
-...give tables of results, including Bundy versions and recommended versions...
+...give tables of results, including Bundy versions and recommended
+versions...Why are we giving and plotting root-variances not
+variances?...
 
@@
 
-Applying this PSF to an exponential or de~Vaucouleurs galaxy is
-slightly more complicated, because the galaxy has not just a flux
-$S_g$ and a central position $\pos_g$; it also has a shape.  Because
-we are only considering these simple galaxies, we are only permitting
-ellipsoidal shapes, which can be represented by a semi-major axis $a$,
-a semi-minor axis $b$, and a position angle $\phi$, or equivalently by
+Applying this PSF to an exp or dev galaxy is slightly more
+complicated, because the galaxy has not just a flux $S_g$ and a
+central position $\pos_g$; it also has a shape.  Because we are only
+considering these simple galaxies, we are only permitting ellipsoidal
+shapes, which can be represented by a semi-major axis $a$, a
+semi-minor axis $b$, and a position angle $\phi$, or equivalently by
 eigenvalues $a, b$ and eigenvectors $\uv_1, \uv_2$, or equivalently by
@@
 \begin{eqnarray}\displaystyle
-I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M_{\exp}} S_g\,a_{\exp,m}\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
+I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
 \\
-\var_{gm} &\equiv& \affine_g\cdot\var_{\exp,m}\cdot\transpose{\affine_g}
+\var_{gm} &\equiv& \affine_g\cdot\var^{\exp}_m\cdot\transpose{\affine_g}
 \\
@@
 
-...what about mixtures of exponentials and de~Vaucouleurs:  No problem!...
+...what about mixtures of exps and devs:  No problem!...
 
@@
 Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
+\bibitem[de~Vaucouleurs(19xx)]{dev}
+de~Vaucouleurs,~G., 19xx, whatever
 \bibitem[Emsellem \etal(1994)]{emsellem}
@@
 Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
+\bibitem[S\'ersic(19xx)]{ser}
+S\'ersic,~X., 19xx, whatever
 \end{thebibliography}
@@
 \clearpage
+\begin{table}
+\caption{The amplitudes and root-variances for the best
+  mixture-of-Gaussian approximations to the various profiles, for
+  different mixture sizes.  The badness and total (dimensionless) flux
+  are given for each approximation.}
+\end{table}
+
+\clearpage
 \begin{figure}
-\includegraphics[width=\figwidth]{K08_MR08_LSD04_dev.png}
-\includegraphics[width=\figwidth]{K08_MR08_LSD04_dev_log.png} \\
-\includegraphics[width=\figwidth]{K08_MR08_LSD04_lup.png}
-\includegraphics[width=\figwidth]{K08_MR08_LSD04_lup_log.png}
+\caption{\textsl{top-left:} The true exp profile (thin black line),
+  the best $M^{\exp}=6$ mixture-of-Gaussian approximation (thick grey
+  line), and the component Gaussians (multiplied by their
+  corresponding amplitudes) contributing to the approximation (thin
+  grey lines).  The plot title text gives $\xi_{\max}$ and the
+  badness. \textsl{top-right:} The same but shown logarithmically.
+  \textsl{bottom-left:} A representation of the residual or devation,
+  on which the badness is computed.  \textsl{bottom-right:} The same
+  but shown logarithmically.\label{fig:exp}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\caption{The dev profile and the best $M^{\dev}=8$ approximation.
+  The panels are equivalent to those in \figurename~\ref{fig:exp}.}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\caption{The luv and lux profiles and approximations.  The panels are
+  equivalent to those in the top-row of \figurename~\ref{fig:exp}.}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\caption{Three ser profiles and approximations.  The top-left,
+  top-right, and bottom-left panels are equivalent to those in the
+  top-right of \figurename~\ref{fig:exp}.  \textsl{bottom-right:} The
+  dependence on the amplitudes $a^{\ser}_m$ and root-variances
+  $\sqrt{v^{\ser}_m}$ on ser index $n$.}
+\end{figure}
+
+\clearpage
+\begin{figure}
+\caption{Comparisons of approximations.  \textsl{top-left:} The
+  dependence of amplitude $a^{\exp}_m$ and root-variance
+  $\sqrt{v^{\exp}_m}$ on $M^{\exp}$ for the exp profile.
+  \textsl{top-right:} The same but for the dev profile.
+  \textsl{bottom-left:} The same but for the lux profile.
+  \textsl{bottom-right:} The same but for the luv profile.}
 \end{figure}

### 2012-10-18 documents/mixture_models.tex
lupton profiles

@@
 % -----
-% - decide which models will make it in
 % - make pickles_to_latex_table.py function
@@
 \usepackage{amssymb,amsmath,mathrsfs}
+
+\newlength{\figurewidth}
+\setlength{\figurewidth}{0.49\textwidth}
 
@@
 
-\acknowledgements It is a pleasure to thank Brendon Brewer (Auckland)
-and Kevin Bundy (IPMU) for valuable comments.  This work was supported
-in part by NASA (grant NNX12AI50G) and the NSF (grant IIS-1124794).
+\acknowledgements It is a pleasure to thank
+      Brendon Brewer (Auckland)
+  and Kevin Bundy (IPMU)
+for valuable comments.  This work was supported in part by NASA (grant
+NNX12AI50G) and the NSF (grant IIS-1124794).
 
@@
 \begin{figure}
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.png}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_residual_log.png}
 \caption{\textsl{top-left:} The true exp profile (thin black line),
@@
 \begin{figure}
-\caption{The dev profile and the best $M^{\dev}=8$ approximation.
-  The panels are equivalent to those in \figurename~\ref{fig:exp}.}
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.png}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_residual_log.png}
+\caption{The dev profile and the best $M^{\dev}=10$ approximation.
+  The panels are equivalent to those in \figurename~\ref{fig:exp}.
+  The residuals look bad in the logarithmic plot, but they are only
+  bad very near the profile center (as expected), where the effect of
+  any realistic point-spread function will average them away.}
 \end{figure}
@@
 \begin{figure}
-\caption{The luv and lux profiles and approximations.  The panels are
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.png}%
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.png}\\
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.png}
+\caption{The lux and luv profiles and approximations.  The panels are
   equivalent to those in the top-row of \figurename~\ref{fig:exp}.}

### 2012-10-18 documents/mixtures/mixture_models.tex
new directory structure
references to dev and ser

--- 2012-10-18 documents/mixture_models.tex
@@
 % - send to Jim Bosch and Kevin Bundy for comments
+% - fix symbols on figures to match text or vice versa (K, M, v, sigma)
 
@@
 Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
-\bibitem[de~Vaucouleurs(19xx)]{dev}
-de~Vaucouleurs,~G., 19xx, whatever
+\bibitem[de~Vaucouleurs(1948)]{dev}
+de~Vaucouleurs,~G., 1948, Annales d'Astrophysique, 11, 247
 \bibitem[Emsellem \etal(1994)]{emsellem}
@@
 Peng,~C.~Y., Ho, L.~C., Impey,~C.~D., \& Rix,~H.-W., 2002, \aj, 124, 266
-\bibitem[S\'ersic(19xx)]{ser}
-S\'ersic,~X., 19xx, whatever
+\bibitem[S{\'e}rsic(1963)]{ser}
+S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
 \end{thebibliography}
@@
 \begin{figure}
-\caption{Three ser profiles and approximations.  The top-left,
-  top-right, and bottom-left panels are equivalent to those in the
-  top-right of \figurename~\ref{fig:exp}.  \textsl{bottom-right:} The
-  dependence on the amplitudes $a^{\ser}_m$ and root-variances
-  $\sqrt{v^{\ser}_m}$ on ser index $n$.}
+\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.png}%
+\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.png}
+\caption{Three ser profiles---with $n=2$, $3$, and $5$---and
+  approximations.  The top-left, top-right, and bottom-left panels are
+  equivalent to those in the top-right of \figurename~\ref{fig:exp}.
+  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser}_m$
+  and root-variances $\sqrt{v^{\ser}_m}$ on ser index $n$.}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[width=\figurewidth]{mixtures_vs_K_exp.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_dev.png}\\
+\includegraphics[width=\figurewidth]{mixtures_vs_K_lux.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_luv.png}
 \caption{Comparisons of approximations.  \textsl{top-left:} The

### 2012-10-18 documents/mixtures/mixture_models.tex
notes to self

@@
 
-...Show plots demonstrating quality of fit, one-d and two-d plots...
+The results of the optimizations are shown in the \tablename\ and
+\figurename s... Discuss briefly the import of it all... Why in the
+tables are we giving and plotting root-variances not variances?...
 
-...give tables of results, including Bundy versions and recommended
-versions...Why are we giving and plotting root-variances not
-variances?...
-
-...show dependence of amplitudes and variances on sersic n for some
-M...be sure to synchronize $\xi_{\max}$...
+...make recommendations: Use the $M^{\lux}=6$ lux and the $M^{\luv}=8$
+luv!...
 

### 2012-10-18 code/optimize_mixture_profiles.py
labeling consistent with paper

@@ -209,8 +209,8 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, badne
         plt.plot(x2, pars[k] * not_normal(x2, pars[k + K]), 'k-', alpha=0.5)
     plt.axvline(max_radius, color='k', alpha=0.25)
     badname = "badness"
-    title = r"%s / $K=%d$ / $\xi_{\max} = %d$ / %s $= %.2f\times 10^{%d}$" % (
-        model, len(pars) / 2, max_radius, badname, badness, log10_squared_deviation)
+    title = r"%s / $M^{\mathrm{%s}}=%d$ / $\xi_{\max} = %d$ / %s $= %.2f\times 10^{%d}$" % (
+        model, model, len(pars) / 2, max_radius, badname, badness, log10_squared_deviation)
     plt.title(title)
     plt.xlim(-0.5, 8.5)
     xlim1 = plt.xlim()

### 2012-10-18 documents/herschel/herschel_dust.tex

--- 2012-08-05 documents/herschel_dust.tex
@@
 % -----
-% - outline
+% - write first draft
+% - run cross-validation (ugh)
 % - make figures
@@
 spatial grid $\theta_k$, the data $d_n$, and the data noise variances
-$\sigma_n$.
+$\sigma_n$.  Given the large number of parameters---three per pixel in
+the dust map model---optimization is challenging.  However, the
+problem is very sparse: Most parameters do not affect most data
+pixels.  We use a brutal hand-built sparse optimizer that is part of
+our image modeling project \TheTractor (CITATION); more about this
+below.
 

### 2012-10-18 documents/mixtures/mixture_models.tex
fractional residual plots for log

@@
 \includegraphics[width=\figurewidth]{exp_K06_MR08_residual.png}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_residual_log.png}
+\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.png}
 \caption{\textsl{top-left:} The true exp profile (thin black line),
@@
   on which the badness is computed.  \textsl{bottom-right:} The same
-  but shown logarithmically.\label{fig:exp}}
+  but shown fractionally and logarithmically.\label{fig:exp}}
 \end{figure}
@@
 \includegraphics[width=\figurewidth]{dev_K10_MR08_residual.png}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_residual_log.png}
+\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.png}
 \caption{The dev profile and the best $M^{\dev}=10$ approximation.

### 2012-10-18 code/optimize_mixture_profiles.py
label tweaking

@@ -254,7 +254,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, badne
     plt.xlim(*xlim1)
     plt.ylim(-1., 1.)
     plt.xlabel(xlabel)
-    plt.ylabel(r"fractional intensity residual")
+    plt.ylabel(r"fractional intensity residual ([true$-$approx] / [true])")
     hogg_savefig(prefix + '_fractional')
     plt.semilogx()
     xmin = 0.001

### 2012-10-18 code/optimize_mixture_profiles.py
ylabels

@@ -240,7 +240,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, badne
     plt.xlim(*xlim1)
     plt.ylim(-1., 1.)
     plt.xlabel(xlabel)
-    plt.ylabel(r"intensity residual (relative to intensity at $\xi = 1$)")
+    plt.ylabel(r"intensity residual (units of intensity at $\xi = 1$)")
     hogg_savefig(prefix + '_residual')
     plt.semilogx()
     xmin = 0.001
@@ -254,7 +254,7 @@ def plot_mixture(pars, prefix, model, max_radius, log10_squared_deviation, badne
     plt.xlim(*xlim1)
     plt.ylim(-1., 1.)
     plt.xlabel(xlabel)
-    plt.ylabel(r"fractional intensity residual ([true$-$approx] / [true])")
+    plt.ylabel(r"fractional intensity residual (residual / intensity)")
     hogg_savefig(prefix + '_fractional')
     plt.semilogx()
     xmin = 0.001

### 2012-10-18 documents/mixtures/mixture_models.tex
residuals no longer look bad

@@
 \caption{The dev profile and the best $M^{\dev}=10$ approximation.
-  The panels are equivalent to those in \figurename~\ref{fig:exp}.
-  The residuals look bad in the logarithmic plot, but they are only
-  bad very near the profile center (as expected), where the effect of
-  any realistic point-spread function will average them away.}
+  The panels are equivalent to those in \figurename~\ref{fig:exp}.}
 \end{figure}

### 2012-10-19 documents/mixtures/mixture_models.tex
sdss

@@
 modeled as a mixture of Gaussians, which is not uncommon (for example,
-the \project{Sloan Digital Sky Survey} (\sdss) imaging pipelines
+the \project{Sloan Digital Sky Survey} imaging pipelines
 make approximate mixture-of-Gaussian PSF models for
@@
 on the order of $10^{13}$ pixels.  We are basing our models on the
-\sdss\ Catalog galaxy models, which only include exponential,
-de~Vaucouleurs, and composite (mixture of the two) radial profiles.
-In detail, in fact, the \sdss\ Catalog models are small modifications
-of these (\citealt{lupton}); details below.  So our goal here is to
-provide replacements for these models in order to improve the
-performance of galaxy image modeling and analysis software.  Some of
-the models we present have been used previously in tools for precise
-photometry (Bundy \etal, forthcoming); they could also be used to
-speed other image-fitting systems, like the very successful
-\project{GALFIT} (\citealt{galfit}).
+\project{Sloan Digital Sky Survey} (\sdss) Catalog galaxy models,
+which only include exponential, de~Vaucouleurs, and composite (mixture
+of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
+models are small modifications of these (\citealt{lupton}); details
+below.  So our goal here is to provide replacements for these models
+in order to improve the performance of galaxy image modeling and
+analysis software.  Some of the models we present have been used
+previously in tools for precise photometry (Bundy \etal, forthcoming);
+they could also be used to speed other image-fitting systems, like the
+very successful \project{GALFIT} (\citealt{galfit}).
 

### 2012-10-19 documents/mixtures/mixture_models.tex
punctuation

@@
 
+\noindent
 Gaussians are remarkable distribution functions.  They have the
@@
 
-We are not the first in this space deconvolution and modeling of
+We are not the first in this space:  Deconvolution and modeling of
 galaxy images with mixtures of Gaussians has been done very

### 2012-10-19 documents/mixtures/mixture_models.tex
noindent doesn't work in aastex
affiliations

@@
 \author{David W. Hogg\altaffilmark{1,2,3} \&
-        Dustin Lang\altaffilmark{4}}
+        Dustin Lang\altaffilmark{4,5}}
 \altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
-\altaffiltext{2}{New York University}
-\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie}
-\altaffiltext{4}{Princeton University Observatory}
+\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Phyiscs, New York University, 4 Washington Place, New York, NY 10003, USA}
+\altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{4}{Princeton University Observatory, Princeton, NJ 08544, USA}
+\altaffiltext{5}{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
 
@@
 
-\noindent
 Gaussians are remarkable distribution functions.  They have the

### 2012-10-19 documents/mixtures/mixture_models.tex
figure caption for dstn

@@
   \textsl{bottom-left:} The same but for the lux profile.
-  \textsl{bottom-right:} The same but for the luv profile.}
+  \textsl{bottom-right:} The same but for the luv profile.\label{fig:M}}
+\end{figure}
+
+\clearpage
+\begin{figure}
+% DSTN:  This should be a 3-across by 2-down figure,
+\caption{Demonstration of use of the profiles, or the implicit
+  generative model in this \documentname.  \textsl{top-left:} The
+  circular dimensionless $M^{\luv}=8$ mixture-of-Gaussian
+  approximation to the luv profile, represented on a very fine pixel
+  grid. \textsl{top-middle:} The ellipse representing the non-trivial
+  affine transformation to be applied to the circular, dimensionless
+  profile.  \textsl{top-right:} The sheared profile.
+  \textsl{bottom-left:} A $K=3$ mixture-of-Gaussian model of the
+  pixel-convolved point-spread function, represented on the very fine
+  pixel grid.  \textsl{bottom-middle:} The sheared profile convolved
+  with the PSF, represented on the very fine pixel grid.
+  \textsl{bottom-right:} The sheared luv convolved with the PSF, but
+  now shown on a realistic pixel grid.  Because by assumption the PSF
+  is a pixel-convolved PSF, the representation on the realistic grid
+  is found simply by interpolating the convolved high-resolution
+  model.\label{fig:example}}
 \end{figure}

### 2012-10-20 code/plot_mixture_profiles.py
pickle_to_latex.py

@@ -14,10 +14,7 @@ rc("font",**{"family":"serif","size":12})
 rc("text", usetex=True)
 import pylab as plt
 import numpy as np
-import scipy.optimize as op
-from multiprocessing import Pool
 import cPickle as pickle
-import os as os
 
 models = ["exp", "ser2", "ser3", "dev", "ser5", "lux", "luv"]
 Ks = [4, 6, 8, 10]

### 2012-10-20 code/pickle_to_latex.py
works

@@ -25,45 +25,48 @@ def get_pars(model, K):
     picklefile.close()
     return pars
 
-def write_table_line(strings):
+def write_table_line(fd, strings):
     endofline = r"&"
     for i,string in enumerate(strings):
         if i + 1 == len(strings): endofline = r"\\"
-        print r" %s %s" % (string, endofline)
+        fd.write(r" %s %s" % (string, endofline) + "\n")
     return None
 
 def write_table(model):
+    tablefn = r"table_%s.tex" % model
     if model == "exp" or model == "lux":
         Ks = [4, 6, 8]
     pars = [get_pars(model, K) for K in Ks]
     clist = "c|"
     for K in Ks: clist += "cc|"
-    print r"\begin{tabular}{%s}" % clist
-    print r"&"
-    print r"\multicolumn{%d}{|c|}{%s} \\" % (2 * len(Ks), model)
-    print r"\hline"
-    print r"$M^{\%s}=$ &" % model
-    write_table_line([r"\multicolumn{2}{|c|}{$%d$}" % K for K in Ks])
-    print r"$m$ &"
-    write_table_line([r"$a^{\%s}_m$ & $\sqrt{v^{\%s}_m}$" % (model, model) for K in Ks])
+    fd = open(tablefn, "w")
+    fd.write(r"\begin{tabular}{%s}" % clist + "\n")
+    fd.write(r"&" + "\n")
+    fd.write(r"\multicolumn{%d}{|c|}{%s} \\" % (2 * len(Ks), model) + "\n")
+    fd.write(r"\hline" + "\n")
+    fd.write(r"$M^{\%s}=$ &" % model + "\n")
+    write_table_line(fd, [r"\multicolumn{2}{|c|}{$%d$}" % K for K in Ks])
+    fd.write(r"$m$ &" + "\n")
+    write_table_line(fd, [r"$a^{\%s}_m$ & $\sqrt{v^{\%s}_m}$" % (model, model) for K in Ks])
     for k in range(np.max(Ks)):
-        print r"$%d$" % (k + 1)
+        fd.write(r"$%d$ &" % (k + 1) + "\n")
         aks = []
         svks = []
         for i,K in enumerate(Ks):
             if k < K:
                 aks += [r"%7.5f" % pars[i][k]]
-                svks += [r"%7.5f" % np.sqrt(pars[i][2 * k])]
[…]

### 2012-10-20 code/pickle_to_latex.py
make all tables x

@@ -36,6 +36,8 @@ def write_table(model):
     tablefn = r"table_%s.tex" % model
     if model == "exp" or model == "lux":
         Ks = [4, 6, 8]
+    else:
+        Ks = [6, 8, 10]
     pars = [get_pars(model, K) for K in Ks]
     clist = "c|"
     for K in Ks: clist += "cc|"
@@ -71,3 +73,9 @@ def write_table(model):
 
 if __name__ == "__main__":
     write_table("exp")
+    write_table("ser2")
+    write_table("ser3")
+    write_table("dev")
+    write_table("ser5")
+    write_table("lux")
+    write_table("luv")

### 2012-10-20 documents/mixtures/mixture_models.tex
figure references
note to self
include tabular results

@@
 % -----
+% - add introduction comments about *why* people fit these profiles -- an opportunity to get on the soapbox
 % - make pickles_to_latex_table.py function
@@
 
-The results of the optimizations are shown in the \tablename\ and
-\figurename s... Discuss briefly the import of it all... Why in the
-tables are we giving and plotting root-variances not variances?...
+The results of the optimizations are shown in \tablename
+s~\ref{tab:exp} through \ref{tab:luv} and \figurename s~\ref{fig:exp}
+through \ref{fig:M}... Discuss briefly the import of it all... Why in
+the tables are we giving and plotting root-variances not variances?...
 
@@
 \begin{table}
+\input{table_exp}
 \caption{The amplitudes and root-variances for the best
-  mixture-of-Gaussian approximations to the various profiles, for
-  different mixture sizes.  The badness and total (dimensionless) flux
-  are given for each approximation.}
+  mixture-of-Gaussian approximations to the exp profile, for
+  different mixture sizes.  The total (dimensionless) flux and the
+  badness are given for each approximation.\label{tab:exp}}
+\end{table}
+
+\clearpage
+\begin{table}
+\input{table_dev}
+\caption{Same as \tablename~\ref{tab:exp} but for the dev profile.}
+\end{table}
+
+\clearpage
+\begin{table}
+\input{table_lux}
+\caption{Same as \tablename~\ref{tab:exp} but for the lux profile.}
+\end{table}
+
+\clearpage
+\begin{table}
+\input{table_luv}
+\caption{Same as \tablename~\ref{tab:exp} but for the luv profile.\label{tab:luv}}
 \end{table}

### 2012-10-20 documents/mixtures/mixture_models.tex
acks

@@
 s~\ref{tab:exp} through \ref{tab:luv} and \figurename s~\ref{fig:exp}
-through \ref{fig:M}... Discuss briefly the import of it all... Why in
-the tables are we giving and plotting root-variances not variances?...
+through \ref{fig:M}.  All the results shown in these figures and
+tables and more are available in machine-readable form from author DWH
+upon request.  In the \tablename s and \figurename s we show
+root-variances rather than variances because these have units of
+half-light radii; they are simple standard deviations for the Gaussian
+components.
 
@@
 for valuable comments.  This work was supported in part by NASA (grant
-NNX12AI50G) and the NSF (grant IIS-1124794).
+NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
+the figures and tables and more are available in machine-readable form
+from author DWH upon request.
 

### 2012-10-20 documents/mixtures/mixture_models.tex
notes to self
table formatting

@@
 % -----
+% - URGENT: badness has another term in it -- the regularization term
 % - add introduction comments about *why* people fit these profiles -- an opportunity to get on the soapbox
-% - make pickles_to_latex_table.py function
 % - make figures showing use of the models in a real situation
@@
 % - send to Jim Bosch and Kevin Bundy for comments
-% - fix symbols on figures to match text or vice versa (K, M, v, sigma)
 
@@
 The results of the optimizations are shown in \tablename
-s~\ref{tab:exp} through \ref{tab:luv} and \figurename s~\ref{fig:exp}
+s~\ref{tab:exp} and \ref{tab:lux} and \figurename s~\ref{fig:exp}
 through \ref{fig:M}.  All the results shown in these figures and
@@
 \begin{table}
-\input{table_exp}
+\input{table_exp}\\
+\vspace{2ex}\\
+\input{table_dev}
 \caption{The amplitudes and root-variances for the best
-  mixture-of-Gaussian approximations to the exp profile, for
-  different mixture sizes.  The total (dimensionless) flux and the
-  badness are given for each approximation.\label{tab:exp}}
+  mixture-of-Gaussian approximations to the exp and dev profiles, for
+  different mixture sizes.  The total (dimensionless) fluxes and the
+  badnesses are given for each approximation.\label{tab:exp}}
 \end{table}
@@
 \begin{table}
-\input{table_dev}
-\caption{Same as \tablename~\ref{tab:exp} but for the dev profile.}
-\end{table}
-
-\clearpage
-\begin{table}
-\input{table_lux}
-\caption{Same as \tablename~\ref{tab:exp} but for the lux profile.}
-\end{table}
-
-\clearpage
-\begin{table}
+\input{table_lux}\\
+\vspace{2ex}\\
 \input{table_luv}
-\caption{Same as \tablename~\ref{tab:exp} but for the luv profile.\label{tab:luv}}
+\caption{Same as \tablename~\ref{tab:exp} but for the lux and luv
+  profiles.\label{tab:lux}}
 \end{table}

### 2012-10-20 documents/mixtures/mixture_models.tex
abstract tweaks

@@
 \begin{abstract}
-Exponential and de~Vaucouleurs profiles are simple and successful
-models for fitting two-dimensional images of galaxies.  One numerical
-issue encountered in this kind of fitting is the pixel rendering and
-convolution (or correlation) of the models with the telescope
-point-spread function (PSF); these operations are slow, and easy to
-get slightly wrong at small radii.  Here we exploit the realization
-that these models can be approximated to arbitrary accuracy with a
-mixture (linear superposition) of two-dimensional Gaussians.  Mixtures
-of Gaussians are fast to render and fast to affine-transform.  Most
-importantly, if you have a mixture-of-Gaussian PSF model (which we
-advocate), the PSF-convolved, affine-transformed galaxy models are
-themselves mixtures of Gaussians and therefore very fast to compute
-precisely.  We present worked examples that can be directly used in
-image fitting; we are using them ourselves.  We discuss extensions to
-S\'ersic models and applications of these ideas to three-dimensional
-deprojection.
+Exponential, de~Vaucouleurs, and S\'ersic profiles are simple and
+successful models for fitting two-dimensional images of galaxies.  One
+numerical issue encountered in this kind of fitting is the pixel
+rendering and convolution (or correlation) of the models with the
+telescope point-spread function (PSF); these operations are slow, and
+easy to get slightly wrong at small radii.  Here we exploit the
+realization that these models can be approximated to arbitrary
+accuracy with a mixture (linear superposition) of two-dimensional
+Gaussians.  Mixtures of Gaussians are fast to render and fast to
+affine-transform.  Most importantly, if you have a mixture-of-Gaussian
+PSF model (which we advocate), the PSF-convolved, affine-transformed
+galaxy models are themselves mixtures of Gaussians and therefore very
+fast to compute, integrate, and render precisely.  We present worked
+examples that can be directly used in image fitting; we are using them
+ourselves.  The mixture profiles we provide can be swapped in to
+replace the standard models in any image-fitting code; they ought to
+make that code faster at very little cost in accuracy.
 \end{abstract}

### 2012-10-20 documents/mixtures/mixture_models.tex
notes to self

@@
 
-The performance advantages we will obtain from using mixtures of
-Gaussians are not simply that convolution itself is trivial.  The
-standard unconvolved galaxy models, especially the de~Vaucouleurs and
-S\'ersic models (\citealt{dev}, \citealt{ser}), are very ill-behaved
-near the galaxy center.  Rendering these profiles precisely near the
-center can be very challenging numerically.  In addition, most of the
-rendering time is spent at very small radii, where PSF convolution is
-going to erase all structure anyway.  That is, a mixture-of-Gaussians
-description of the de~Vaucouleurs profile saves time both in rendering
-\emph{and} in PSF convolution; it produces profiles that are
-approximate but very high performance when all real uses of the
-profiles are PSF-convolved, as they usually are.
+... insert here a paragraph about the \emph{wisdom} of fitting these
+standard profiles: When is it scientifically okay?...
+
+The performance advantages we will obtain from using
+mixture-of-Gaussian approximations are not simply that convolution
+itself is trivial.  The standard unconvolved galaxy models, especially
+the de~Vaucouleurs and S\'ersic models (\citealt{dev}, \citealt{ser}),
+are very ill-behaved near the galaxy center.  Rendering these profiles
+precisely near the center can be very challenging numerically.  In
+addition, most of the rendering time is spent at very small radii,
+where PSF convolution is going to erase all structure anyway.  That
+is, a mixture-of-Gaussians description of the de~Vaucouleurs profile
+saves time both in rendering \emph{and} in PSF convolution; it
+produces profiles that are approximate but very high performance when
+all real uses of the profiles are PSF-convolved, as they usually are.
 
@@
 
-...make recommendations: Use the $M^{\lux}=6$ lux and the $M^{\luv}=8$
+... make recommendations: Use the $M^{\lux}=6$ lux and the $M^{\luv}=8$
 luv!...
@@
 
-...pegagogical figure demonstrating these operations and their results...
+... pegagogical figure demonstrating these operations and their results...
 
-...what about mixtures of exps and devs:  No problem!...
+... what about mixtures of exps and devs:  No problem!...
+
+... finally: The insights that \emph{could} be gained by interrogating
+these profile approximations in three dimensions...!
 

### 2012-10-20 documents/mixtures/mixture_models.tex
notes to self
notes re convolution

@@
 % - URGENT: badness has another term in it -- the regularization term
-% - add introduction comments about *why* people fit these profiles -- an opportunity to get on the soapbox
-% - make figures showing use of the models in a real situation
+% - DSTN: make figures showing use of the models in a real situation; see last figure
 % - does the Lupton paper give the profiles, or should we cite the code directly?
+% - make sure the technical issues about convolution and interpolation don't exist *only* in the figure caption
 % - send to Jim Bosch and Kevin Bundy for comments
@@
   is a pixel-convolved PSF, the representation on the realistic grid
-  is found simply by interpolating the convolved high-resolution
-  model.\label{fig:example}}
+  is found simply by interpolating to the pixel centers the convolved
+  high-resolution model.\label{fig:example}}
 \end{figure}

### 2012-10-22 documents/mixtures/mixture_models.tex
badness regularization, advice to users of various kinds

@@
 % -----
-% - URGENT: badness has another term in it -- the regularization term
+% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
 % - DSTN: make figures showing use of the models in a real situation; see last figure
-% - does the Lupton paper give the profiles, or should we cite the code directly?
+% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
 % - make sure the technical issues about convolution and interpolation don't exist *only* in the figure caption
+% - HOGG: need Bundy reference
+% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
 % - send to Jim Bosch and Kevin Bundy for comments
@@
 high-angular-resolution intensity field is convolved with the
-point-spread function (PSF) before being observed.  This is
-particularly useful in data sets in which the PSF itself has also been
-modeled as a mixture of Gaussians, which is not uncommon (for example,
-the \project{Sloan Digital Sky Survey} imaging pipelines
-make approximate mixture-of-Gaussian PSF models for
-every imaging field; \citealt{lupton}).
+point-spread function (PSF) before being observed.  This is always
+easier with Gaussians than cuspy profiles, but is \emph{particularly}
+useful in data sets in which the PSF itself has also been modeled as a
+mixture of Gaussians, which is not uncommon (for example, the
+\project{Sloan Digital Sky Survey} imaging pipelines make approximate
+mixture-of-Gaussian PSF models for every imaging field;
+\citealt{lupton}).
 
@@
 
-... insert here a paragraph about the \emph{wisdom} of fitting these
-standard profiles: When is it scientifically okay?...
+Whenever an investigator is fitting PSF-convolved exponential or
+de~Vaucouleurs profiles, the models presented here will improve code
+performance.  That doesn't mean that doing such fitting is a Good
+Idea.  These profiles are effective models for galaxies at low
+signal-to-noise (as they are used in the \project{SDSS}) and they are
+useful templates for performing consistent photometry across
+morphologically diverse objects (as we use them in \citealt{bundy}).
+They are sometimes used to distinguish disk and bulge light; these
+uses are very prone to over-interpretation: There is no theoretical
+argument and only weak observational arguments that the
+rotation-supported parts of galaxies are always exponential and that
+the kinematically hot components are de~Vaucouleurs.
 
@@
 analysis software.  Some of the models we present have been used
-previously in tools for precise photometry (Bundy \etal, forthcoming);
-they could also be used to speed other image-fitting systems, like the
-very successful \project{GALFIT} (\citealt{galfit}).
+previously in tools for precise photometry (\citealt{bundy}); they
+could also be used to speed other image-fitting systems, like the very
+successful \project{GALFIT} (\citealt{galfit}).
 
@@
 linearly with radius) to make it equivalent to the two-dimensional
-chi-squared.
+chi-squared.  We also add to the badness a very tiny coefficient (on
+the order of $10^{-3}$ of the best-fit badness) times the sum of the
+variances $v_m$ for regularization.  In practice, this term doesn't
+have much effect and could be dropped.
 
@@
 
-... make recommendations: Use the $M^{\lux}=6$ lux and the $M^{\luv}=8$
-luv!...
+In our work on \project{The Tractor}, we use the $M^{\lux}=6$ lux and
+the $M^{\luv}=8$ luv profiles.  Our advice to users would be to do the
+same.  We use the lux and luv over the exp and dev partly because of
+their better behaviors numerically, and partly because they and we are
+both part of the \project{SDSS} tradition.  These---$M^{\lux}=6$ and
+$M^{\luv}=8$ are good compromises between mixture complexity ($M$) and
+quality of fit.  Also, even the best-fitting late-type and early-type
+galaxies deviate from exponential and de~Vaucouleurs fits by more than
+do these high-quality mixture-of-Gaussian approxmations; no precision
+is lost.
 
@@
 
-... pegagogical figure demonstrating these operations and their results...
+DSTN... pegagogical figure demonstrating these operations and their results...
 
@@
 Bovy,~J. \etal, 2012, \aj, 749, 41
+\bibitem[Bundy \etal(2012)]{bundy}
+Bundy,~K.~??
 \bibitem[Cappellari(2002)]{cappellari}

### 2012-10-22 documents/mixtures/mixture_models.tex
last paragraphs

@@
 % - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
-% - make sure the technical issues about convolution and interpolation don't exist *only* in the figure caption
 % - HOGG: need Bundy reference
 % - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
-% - send to Jim Bosch and Kevin Bundy for comments
+% - HOGG: send to Lupton and Bundy for comments
+% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
+% - DSTN: is there anyone we need to ack on your side, or granting agencies?
+% - HOGG: should we make Machine Readable Tables for PASP?
 
@@
 both part of the \project{SDSS} tradition.  These---$M^{\lux}=6$ and
-$M^{\luv}=8$ are good compromises between mixture complexity ($M$) and
-quality of fit.  Also, even the best-fitting late-type and early-type
-galaxies deviate from exponential and de~Vaucouleurs fits by more than
-do these high-quality mixture-of-Gaussian approxmations; no precision
-is lost.
+$M^{\luv}=8$---are good compromises between mixture complexity ($M$)
+and quality of fit (badness).  Also, even the best-fitting late-type
+and early-type galaxies deviate from exponential and de~Vaucouleurs
+fits by more than do these high-quality mixture-of-Gaussian
+approxmations; no precision is lost.
 
@@
 
-... what about mixtures of exps and devs:  No problem!...
+In the \project{SDSS}, \project{GALFIT}, and much of our own work, the
+models that are fit are (effectively) mixtures of exp and dev or exp
+and ser or lux and luv profiles.  Mixtures of profiles that are each
+themselves mixtures of Gaussians are no harder to render than either
+profile separately.  There is some book-keeping, of course, because
+each component gets affine-transformed separately before they are both
+PSF-convolved.
 
-... finally: The insights that \emph{could} be gained by interrogating
-these profile approximations in three dimensions...!
+One amusing aspect of mixture-of-Gaussian profiles has to do with
+projection from three to two dimensions.  The projection of a
+three-dimensional Gaussian is a two-dimensional Gaussian; the
+two-dimensional, rigid, circular approximations we have made for the
+ser profiles can be deprojected to rigid, spherical approximations to
+the three-dimensional profiles trivially.  The two-dimensional models
+we have started with are not accurate models of galaxies in
+detail---no galaxy follows exactly any ser profile---so deprojection
+of our approximations are are not that interesting in themselves.
+However, the general program of fitting two-dimensional sources with
+mixtures of Gaussians may have strong implications in the future for
+three-dimensional modeling and deprojection.
 

### 2012-10-22 documents/mixtures/mixture_models.tex
figure reference
notes to self, text, and MoG acronym

@@
 % -----
+% - DSTN: we need *some* estimate of the speed-up this delivers to The Tractor in the abstract and in the text -- of course this is bullshit because we get to decide how precisely we were doing things prior to using the mixtures -- but we need it since it is our main point.
 % - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
@@
 accuracy with a mixture (linear superposition) of two-dimensional
-Gaussians.  Mixtures of Gaussians are fast to render and fast to
-affine-transform.  Most importantly, if you have a mixture-of-Gaussian
-PSF model (which we advocate), the PSF-convolved, affine-transformed
-galaxy models are themselves mixtures of Gaussians and therefore very
-fast to compute, integrate, and render precisely.  We present worked
-examples that can be directly used in image fitting; we are using them
-ourselves.  The mixture profiles we provide can be swapped in to
-replace the standard models in any image-fitting code; they ought to
-make that code faster at very little cost in accuracy.
+Gaussians (MoGs).  MoGs are fast to render and fast to
+affine-transform.  Most importantly, if you have a MoG PSF model
+(which we advocate), the PSF-convolved, affine-transformed galaxy
+models are themselves MoGs and therefore very fast to compute,
+integrate, and render precisely.  We present worked examples that can
+be directly used in image fitting; we are using them ourselves.  The
+MoG profiles we provide can be swapped in to replace the standard
+models in any image-fitting code; they sped up model fitting in our
+projects by XXX; they ought to make any code faster at very little
+cost in accuracy.
 \end{abstract}
@@
 variance tensors.  The ubiquity of convolution operations in astronomy
-suggests the widespread adoption of mixture-of-Gaussian modeling.  We
-have pioneered this in the area of distribution modeling in large
-numbers of dimensions (\citealt{xd, xdqso, xdqsoz}), where convolution
-occurs because the true or noise-free distribution is convolved with
-the noise before being observed.  Here we are going to capitalize on
-the convolution properties of mixtures of Gaussians in modeling galaxy
+suggests the widespread adoption of mixture-of-Gaussian (MoG)
+modeling.  We have pioneered this in the area of distribution modeling
+in large numbers of dimensions (\citealt{xd, xdqso, xdqsoz}), where
+convolution occurs because the true or noise-free distribution is
+convolved with the noise before being observed.  Here we are going to
+capitalize on the convolution properties of MoGs in modeling galaxy
 morphologies in imaging data, where the true or
@@
 useful in data sets in which the PSF itself has also been modeled as a
-mixture of Gaussians, which is not uncommon (for example, the
-\project{Sloan Digital Sky Survey} imaging pipelines make approximate
-mixture-of-Gaussian PSF models for every imaging field;
-\citealt{lupton}).
+MoG, which is not uncommon (for example, the \project{Sloan Digital
+  Sky Survey} imaging pipelines make approximate MoG PSF models for
+every imaging field; \citealt{lupton}).
 
 We are not the first in this space: Deconvolution and modeling of
-galaxy images with mixtures of Gaussians has been done very
-successfully before (for example, \citealt{bendinelli, emsellem,
-  bendinelli2, cappellari}).  However, the idea behind those projects
-was to use the mixture of Gaussians to provide a very free form for
-the isophotes or morphologies of resolved galaxies or other complex
-scenes.  Here our goals are very limited: We want to improve the
-performance of standard galaxy image model fitting by expressing the
-standard galaxy models---the exponential and de~Vaucouleurs
-profiles---as rigid mixtures of Gaussians.
+galaxy images with MoGs has been done very successfully before (for
+example, \citealt{bendinelli, emsellem, bendinelli2, cappellari}; they
+use the name ``multi-Gaussian expansion'' or MGE).  However, the idea
+behind those projects was to use the MoGs to provide a very free form
+for the isophotes or morphologies of resolved galaxies or other
+complex scenes.  Here our goals are very limited: We want to improve
+the performance of standard galaxy image model fitting by expressing
+the standard galaxy models---the exponential and de~Vaucouleurs
+profiles---as rigid MoGs.
 
@@
 
-The performance advantages we will obtain from using
-mixture-of-Gaussian approximations are not simply that convolution
-itself is trivial.  The standard unconvolved galaxy models, especially
-the de~Vaucouleurs and S\'ersic models (\citealt{dev}, \citealt{ser}),
-are very ill-behaved near the galaxy center.  Rendering these profiles
-precisely near the center can be very challenging numerically.  In
-addition, most of the rendering time is spent at very small radii,
-where PSF convolution is going to erase all structure anyway.  That
-is, a mixture-of-Gaussians description of the de~Vaucouleurs profile
-saves time both in rendering \emph{and} in PSF convolution; it
-produces profiles that are approximate but very high performance when
-all real uses of the profiles are PSF-convolved, as they usually are.
+The performance advantages we will obtain from using MoG
+approximations are not simply that convolution itself is trivial.  The
+standard unconvolved galaxy models, especially the de~Vaucouleurs and
+S\'ersic models (\citealt{dev}, \citealt{ser}), are very ill-behaved
+near the galaxy center.  Rendering these profiles precisely near the
+center can be very challenging numerically.  In addition, most of the
+rendering time is spent at very small radii, where PSF convolution is
+going to erase all structure anyway.  That is, a MoG description of
+the de~Vaucouleurs profile saves time both in rendering \emph{and} in
+PSF convolution; it produces profiles that are approximate but very
+high performance when all real uses of the profiles are PSF-convolved,
+as they usually are.
 
@@
 image synthesis (returned to at the end of this \documentname)
-consists of PSF convolution (an arithmetic operation on the two
-mixtures of Gaussians) followed by sampling (evaluation of the
-resulting mixture at the pixel centers).
+consists of PSF convolution (an arithmetic operation on the two MoGs)
+followed by sampling (evaluation of the resulting MoG at the pixel
+centers).
 
@@
 symmetric $2\times 2$ variance tensor or matrix, and implicitly the
-vectors are column vectors.  A \emph{mixture of Gaussians} is a linear
+vectors are column vectors.  A mixture of Gaussians is a linear
 superposition of Gaussians.  Any positive two-dimensional function
@@
 case any two-dimensional probability distribution function---can be
-represented as a mixture of Gaussians to arbitrary accuracy; that is
+represented as a MoG to arbitrary accuracy; that is
 \begin{eqnarray}
@@
 quantity $\pos$, the $\approx$ symbol implies approximation, $K$ is
-the number of Gaussians used in the mixture, and the $K$ Gaussians
+the number of Gaussians used in the MoG, and the $K$ Gaussians
 have amplitudes $a_k$, means $\mean_k$, and variance tensors $\var_k$.
@@
 function.  We seek the best (where ``best'' will be defined below)
-$M^{\exp}$-Gaussian mixture (where $M^{\exp}$ is an integer)
-approximation
+$M^{\exp}$-Gaussian MoG (where $M^{\exp}$ is an integer) approximation
 \begin{eqnarray}\displaystyle
@@
 where all of the means are exactly zero and all of the variances
-$\var^{\exp}_m$ in the mixture can be represented as a scalar
+$\var^{\exp}_m$ in the MoG can be represented as a scalar
 $v^{\exp}_m$ multiplied by the identity matrix $\identity$
@@
 concentrated profiles dev and luv to determine total fluxes can be
-challenging.  This is not true for the mixture-of-Gaussian
-approximations: Each Gaussian is normalized, so the sum of the
-amplitudes $\sum_m a^{\luv}_m$ (for the luv profile, say) gives the
-total flux for the mixture-of-Gaussians approximation to that profile.
+challenging.  This is not true for the MoG approximations: Each
+Gaussian is normalized, so the sum of the amplitudes $\sum_m
+a^{\luv}_m$ (for the luv profile, say) gives the total flux for the
+MoG approximation to that profile.
 
-We seek the \emph{best} mixture-of-Gaussian approximations.  This
-necessitates definition of the word ``best''.  If we think of the
-profiles as being two-dimensional probability distribution functions
-(for, say, the arrivals of photons), then one natural choice is the
-K-L divergence or similar cross-entropy or information-theoretic
-measure.  However, in typical astronomical imaging, the galaxy is
-superimposed on a substantial, flat sky level, and the noise in the
-data is close to Gaussian.  This suggests more chi-squared-like
-objectives.  We adopt the latter, in part because they are most
-appropriate for our specific proposed application (modeling \sdss-like
-astronomical imaging), but experiments we have performed suggest that
-information-theoretic objectives also lead to good results.
+We seek the \emph{best} MoG approximations.  This necessitates
+definition of the word ``best''.  If we think of the profiles as being
+two-dimensional probability distribution functions (for, say, the
+arrivals of photons), then one natural choice is the K-L divergence or
+similar cross-entropy or information-theoretic measure.  However, in
+typical astronomical imaging, the galaxy is superimposed on a
+substantial, flat sky level, and the noise in the data is close to
+Gaussian.  This suggests more chi-squared-like objectives.  We adopt
+the latter, in part because they are most appropriate for our specific
+proposed application (modeling \sdss-like astronomical imaging), but
+experiments we have performed suggest that information-theoretic
+objectives also lead to good results.
 
@@
 \emph{badness}---is a squared residual between the exact profile
-function $Q(\spos)$ and its mixture-of-Gaussian approximation.  It is
-designed to be equivalent to a chi-squared statistic in a
-homoskedastic two-dimensional image of the profile taken with
-extremely high angular resolution (pixels of size 0.001 the half-light
-radius) and vanishing point-spread function.  Quantitatively the
-badness is defined to be the mean squared residual in the $Q(\spos)$
-functions, which are normalized to have unit intensity at the
-half-light radius, averaged over a two-dimensional circular region in
-the $\spos$ plane centered on the (circularly symmetric) profile and
-extending out to radius $\xi_{\max}$.  We use $\xi_{\max}=8$ for all
-profiles except the $\lux$ profile, for which we use $\xi_{\max}=4$.
-In practice, the badness is computed in a one-dimensional numerical
-integral but the integral is weighted in radius (weight increasing
-linearly with radius) to make it equivalent to the two-dimensional
-chi-squared.  We also add to the badness a very tiny coefficient (on
-the order of $10^{-3}$ of the best-fit badness) times the sum of the
-variances $v_m$ for regularization.  In practice, this term doesn't
-have much effect and could be dropped.
+function $Q(\spos)$ and its MoG approximation.  It is designed to be
+equivalent to a chi-squared statistic in a homoskedastic
+two-dimensional image of the profile taken with extremely high angular
+resolution (pixels of size 0.001 the half-light radius) and vanishing
+point-spread function.  Quantitatively the badness is defined to be
+the mean squared residual in the $Q(\spos)$ functions, which are
+normalized to have unit intensity at the half-light radius, averaged
+over a two-dimensional circular region in the $\spos$ plane centered
+on the (circularly symmetric) profile and extending out to radius
+$\xi_{\max}$.  We use $\xi_{\max}=8$ for all profiles except the
+$\lux$ profile, for which we use $\xi_{\max}=4$.  In practice, the
+badness is computed in a one-dimensional numerical integral but the
+integral is weighted in radius (weight increasing linearly with
+radius) to make it equivalent to the two-dimensional chi-squared.  We
+also add to the badness a very tiny coefficient (on the order of
+$10^{-3}$ of the best-fit badness) times the sum of the variances
+$v_m$ for regularization.  In practice, this term doesn't have much
+effect and could be dropped.
 
@@
 and early-type galaxies deviate from exponential and de~Vaucouleurs
-fits by more than do these high-quality mixture-of-Gaussian
-approxmations; no precision is lost.
+fits by more than do these high-quality MoG approxmations; no
+precision is lost.
 
-The value of these mixture-of-Gaussian galaxy components comes when
-they are to be convolved with a PSF (usually in fact a pixel-convolved
-PSF) that is itself represented as a mixture of Gaussians.  In this
-scenario, the PSF $\psi(\Delta\pos)$---which is thought of as a
-function of focal-plane displacement $\Delta\pos$ away from, say, a
-true stellar position---is represented as a $K$-Gaussian mixture
+The value of these MoG galaxy components comes when they are to be
+convolved with a PSF (usually in fact a pixel-convolved PSF) that is
+itself also represented as a MoG.  In this scenario, the PSF
+$\psi(\Delta\pos)$---which is thought of as a function of focal-plane
+displacement $\Delta\pos$ away from, say, a true stellar position---is
+represented as a $K$-Gaussian MoG
 \begin{eqnarray}\displaystyle
@@
 \end{eqnarray}
-That is, when the PSF is represented as a mixture of Gaussians, any
-image of a star---or indeed any image of any set of stars---is also
-represented as a mixture of Gaussians.
+That is, when the PSF is represented as a MoG, any image of a
+star---or indeed any image of any set of stars---is also represented
+as a MoG.
 
@@
 $\lux$ cases are all similar.  Note the important and key result of
-this \documentname, to wit, that a mixture-of-Gaussian galaxy model
-(with $M$ components) convolved with a mixture-of-Gaussian PSF model
-(with $K$ components) yields a mixture-of-Gaussian image (with
-$[M\,K]$ components).
+this \documentname, to wit, that a MoG galaxy model (with $M$
+components) convolved with a MoG PSF model (with $K$ components)
+yields a MoG model image (with $[M\,K]$ components).
 
-DSTN... pegagogical figure demonstrating these operations and their results...
+In the above we said ``pixel-convolved PSF''.  In every context, when
+modeling images, it is valuable to use the pixel-convolved PSF.  With
+this definition of the PSF, the pixelized image is the PSF-convolved
+true model evaluated at the pixel centers.  This operation is fast.
+Other definitions for the PSF (the non-pixel-convolved, for example)
+require that the user to \emph{two} convolutions, the first withe the
+PSF and the second with the square (or worse) pixel.  Our advice: Only
+fit and use pixel-convolved PSFs.
+
+DSTN... In \figurename~\ref{fig:example} we show...
+
+If your PSF is not in MoG form, it is still the case that convolution
+of a MoG approximation to a dev (say) profile will in general be
+easier than convolution of the original dev profile.  The reason is
+that convolution of a Gaussian with any PSF is fast (indeed most
+image-processing languages have such functions built in); the
+PSF-convolved profile becomes in this case just a mixture of
+Gaussian-convolved PSFs.
+
+The speed-ups that can be obtained by using MoG approximations can be
+very large... DSTN: How much did we benefit when we switched to MoGs
+in \project{The Tractor}?...
 
@@
 
-One amusing aspect of mixture-of-Gaussian profiles has to do with
-projection from three to two dimensions.  The projection of a
-three-dimensional Gaussian is a two-dimensional Gaussian; the
-two-dimensional, rigid, circular approximations we have made for the
-ser profiles can be deprojected to rigid, spherical approximations to
-the three-dimensional profiles trivially.  The two-dimensional models
-we have started with are not accurate models of galaxies in
-detail---no galaxy follows exactly any ser profile---so deprojection
-of our approximations are are not that interesting in themselves.
-However, the general program of fitting two-dimensional sources with
-mixtures of Gaussians may have strong implications in the future for
-three-dimensional modeling and deprojection.
+One amusing aspect of MoG profiles has to do with projection from
+three to two dimensions.  The projection of a three-dimensional
+Gaussian is a two-dimensional Gaussian; the two-dimensional, rigid,
+circular approximations we have made for the ser profiles can be
+deprojected to rigid, spherical approximations to the
+three-dimensional profiles trivially.  The two-dimensional models we
+have started with are not accurate models of galaxies in detail---no
+galaxy follows exactly any ser profile---so deprojection of our
+approximations are are not that interesting in themselves.  However,
+the general program of fitting two-dimensional sources with MoGs may
+have strong implications in the future for three-dimensional modeling
+and deprojection.
 

### 2012-10-22 documents/mixtures/mixture_models.tex
words about speed

@@
 % -----
-% - DSTN: we need *some* estimate of the speed-up this delivers to The Tractor in the abstract and in the text -- of course this is bullshit because we get to decide how precisely we were doing things prior to using the mixtures -- but we need it since it is our main point.
+% - DSTN: check sentences I say about ``ten times faster''.
 % - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
@@
 models in any image-fitting code; they sped up model fitting in our
-projects by XXX; they ought to make any code faster at very little
-cost in accuracy.
+projects by an order of magnitude; they ought to make any code faster
+at very little cost in accuracy.
 \end{abstract}
@@
 The speed-ups that can be obtained by using MoG approximations can be
-very large... DSTN: How much did we benefit when we switched to MoGs
-in \project{The Tractor}?...
+very large.  In our image-modeling project \project{The Tractor}, we
+were PSF-convolving by rendering the profiles (especially the profile
+centers) at very high resolution (hundreds to thousands of resolution
+elements in the central pixel are necessary for good precision on the
+dev profile).  We were then convolving that high-resolution model with
+a low-resolution PSF and rendering to a low-resolution image pixel
+grid.  These expensive operations were obviated by the MoG profiles,
+which involve only rendering a small number of Gaussians at the pixel
+centers on the low-resolution pixel grid.  The MoG approximations
+saved us more than an order of magnitude in compute time, especially
+in optimization, where derivatives have to be taken with respect to
+the unconvolved model properties.
 

### 2012-10-22 documents/mixtures/mixture_models.tex
abstract tweak

@@
 Gaussians (MoGs).  MoGs are fast to render and fast to
-affine-transform.  Most importantly, if you have a MoG PSF model
-(which we advocate), the PSF-convolved, affine-transformed galaxy
+affine-transform.  Most importantly, if you have a MoG model for the
+pixel-convolved PSF, the PSF-convolved, affine-transformed galaxy
 models are themselves MoGs and therefore very fast to compute,
@@
 projects by an order of magnitude; they ought to make any code faster
-at very little cost in accuracy.
+at essentially no cost in precision.
 \end{abstract}

### 2012-10-22 documents/mixtures/mixture_models.tex
small spelling fixes

@@
 to best approximate the traditional galaxy profile functions, under
-some sensible definiton of the word ``best'', as a function of the
+some sensible definition of the word ``best'', as a function of the
 model complexity parameters (numbers of components) $(M^{\exp},
@@
 
-In addition to these, there are general S\'ersic (ser) profiles, of
+In addition to these, there are general S\'ersic (``ser'') profiles, of
 which the exp and dev profiles are special cases.  The general ser
@@
 and early-type galaxies deviate from exponential and de~Vaucouleurs
-fits by more than do these high-quality MoG approxmations; no
+fits by more than do these high-quality MoG approximations; no
 precision is lost.
@@
 Other definitions for the PSF (the non-pixel-convolved, for example)
-require that the user to \emph{two} convolutions, the first withe the
+require that the user do \emph{two} convolutions, the first with the
 PSF and the second with the square (or worse) pixel.  Our advice: Only
@@
 galaxy follows exactly any ser profile---so deprojection of our
-approximations are are not that interesting in themselves.  However,
+approximations are not that interesting in themselves.  However,
 the general program of fitting two-dimensional sources with MoGs may

### 2012-10-22 code/tractor-example-plots.py
mog example plots

@@ -1,0 +1,230 @@
+'''
+% DSTN:  This should be a 3-across by 2-down figure,
+\caption{Demonstration of use of the profiles, or the implicit
+  generative model in this \documentname.  \textsl{top-left:} The
+  circular dimensionless $M^{\luv}=8$ mixture-of-Gaussian
+  approximation to the luv profile, represented on a very fine pixel
+  grid. \textsl{top-middle:} The ellipse representing the non-trivial
+  affine transformation to be applied to the circular, dimensionless
+  profile.  \textsl{top-right:} The sheared profile.
+  \textsl{bottom-left:} A $K=3$ mixture-of-Gaussian model of the
+  pixel-convolved point-spread function, represented on the very fine
+  pixel grid.  \textsl{bottom-middle:} The sheared profile convolved
+  with the PSF, represented on the very fine pixel grid.
+  \textsl{bottom-right:} The sheared luv convolved with the PSF, but
+  now shown on a realistic pixel grid.  Because by assumption the PSF
+  is a pixel-convolved PSF, the representation on the realistic grid
+  is found simply by interpolating the convolved high-resolution
+  model.\label{fig:example}}
+\end{figure}
+'''
+
+import matplotlib
+matplotlib.use('Agg')
+import pylab as plt
+import numpy as np
+
+from tractor import *
+from tractor.sdss_galaxy import *
+from astrometry.util.util import Tan
+
+ps = PlotSequence('trdemo')
+
+ra,dec = 0.,0.
+rd = RaDecPos(ra, dec)
+m1 = Flux(100.)
+re = 1.
+s1 = GalaxyShape(re, 1., 0.)
+s2 = GalaxyShape(re, 0.5, 40.)
+g1 = DevGalaxy(rd, m1, s1)
+
+H,W = 500,500
+data = np.zeros((H,W))
+invvar = np.ones((H,W))
+psf1 = NCircularGaussianPSF([1e-3], [1.])
[…]

### 2012-10-23 code/tractor-example-plots.py
figure tweaks

@@ -27,6 +27,9 @@ import numpy as np
 from tractor import *
 from tractor.sdss_galaxy import *
 from astrometry.util.util import Tan
+from astrometry.util.plotutils import antigray
+
+from matplotlib.ticker import FixedFormatter
 
 ps = PlotSequence('trdemo')
 
@@ -38,11 +41,15 @@ s1 = GalaxyShape(re, 1., 0.)
 s2 = GalaxyShape(re, 0.5, 40.)
 g1 = DevGalaxy(rd, m1, s1)
 
-H,W = 500,500
+H,W = 501,501
+print 'Size', W
 data = np.zeros((H,W))
 invvar = np.ones((H,W))
-psf1 = NCircularGaussianPSF([1e-3], [1.])
-pixscale = re * 8. * 2. / float(W) / 3600.
+psf1 = NCircularGaussianPSF([1e-12], [1.])
+psf1.minradius = 100.
+nre = 3.
+pixscale = re * 8. * nre / float(W) / 3600.
+print 'Pixscale', pixscale
 wcs1 = FitsWcs(Tan(ra, dec, W/2. + 0.5, H/2. + 0.5,
 				   pixscale, 0., 0., pixscale, W, H))
 
@@ -52,27 +59,39 @@ im1 = Image(data=data, invvar=invvar, psf=psf1, wcs=wcs1,
 
 tractor = Tractor([im1], [g1])
 
+print 'getting initial model'
 mod = tractor.getModelImage(0)
+print 'gotit'
 
-ima = dict(interpolation='nearest', origin='lower')
+galext = [-nre*4,nre*4]*2
+ima = dict(interpolation='nearest', origin='lower', cmap=antigray)
 
-logmrange = np.max(np.log(mod)) - np.min(np.log(mod))
+mod = np.maximum(mod, 1e-42)
+logmrange = np.max(np.log10(mod)) - np.min(np.log10(mod))
+logmrange = 6.
+
+def plog(mod):
+	plt.clf()
+	lm = np.log10(mod + 1e-42)
+	mx = lm.max()
+	plt.imshow(lm-mx, vmax=0., vmin=-logmrange,
+			   extent=galext, **ima)
+	#plt.gray()
+	ticks = np.arange(-6, 1)
[…]

### 2012-10-23 code/tractor-example-plots.py
add tractor demo figs

@@ -31,7 +31,7 @@ from astrometry.util.plotutils import antigray
 
 from matplotlib.ticker import FixedFormatter
 
-ps = PlotSequence('trdemo')
+ps = PlotSequence('trdemo', suffixes=['png','pdf'])
 
 ra,dec = 0.,0.
 rd = RaDecPos(ra, dec)
@@ -143,7 +143,7 @@ o = np.array([[2.,0.],[0.,0.3]])
 V2 = np.dot(R.T, np.dot(o, R))
 print 'V', V2
 
-V1 = np.eye(2) * 3.
+V1 = np.eye(2) * 2.
 print 'V', V1
 
 V3 = np.array([[4., -0.5],[-0.5,3.]])
@@ -151,8 +151,9 @@ V3 = np.array([[4., -0.5],[-0.5,3.]])
 VV = np.array([V1, V2, V3])
 print 'VV', VV
 
-psf2 = GaussianMixturePSF(np.array([0.6, 0.6, 0.6]),
-						  np.array([[0.,0.], [-.5,0.], [1.,-0.5]]), VV)
+psf2 = GaussianMixturePSF(np.array([0.6, 0.4, 0.6]),
+						  np.array([[0.,0.5], [-.5,0.], [1.,-0.5]]), VV)
+psf2.computeRadius()
 im1.psf = psf2
 
 p = psf2.getPointSourcePatch(0., 0.,)

### 2012-10-23 documents/mixtures/mixture_models.tex
add tractor demo figs
notes2self
Figure tweaks: 3x2 -> 2x3 remove axis ticks remove colorbars zoom out last two psf-convolved galaxies

@@
 \begin{figure}
+\setlength{\figurewidth}{0.33\textwidth}
 % DSTN:  This should be a 3-across by 2-down figure,
+\begin{tabular}{@{}c@{}c@{}}
+\includegraphics[width=\figurewidth]{trdemo-01} &
+\includegraphics[width=\figurewidth]{trdemo-02} \\
+\includegraphics[width=\figurewidth]{trdemo-03} &
+\includegraphics[width=\figurewidth]{trdemo-06} \\
+\includegraphics[width=\figurewidth]{trdemo-08} &
+\includegraphics[width=\figurewidth]{trdemo-09}
+\end{tabular}
 \caption{Demonstration of use of the profiles, or the implicit
@@
   approximation to the luv profile, represented on a very fine pixel
-  grid. \textsl{top-middle:} The ellipse representing the non-trivial
+  grid. \textsl{top-right:} The ellipse representing the non-trivial
   affine transformation to be applied to the circular, dimensionless
-  profile.  \textsl{top-right:} The sheared profile.
-  \textsl{bottom-left:} A $K=3$ mixture-of-Gaussian model of the
+  profile.  \textsl{middle-left:} The sheared profile.
+  \textsl{middle-right:} A $K=3$ mixture-of-Gaussian model of the
   pixel-convolved point-spread function, represented on the very fine
-  pixel grid.  \textsl{bottom-middle:} The sheared profile convolved
+  pixel grid.  \textsl{bottom-left:} The sheared profile convolved
   with the PSF, represented on the very fine pixel grid.

### 2012-10-23 code/tractor-example-plots.py

@@ -31,6 +31,8 @@ from astrometry.util.plotutils import antigray
 
 from matplotlib.ticker import FixedFormatter
 
+plt.figure(figsize=(4,4))
+
 ps = PlotSequence('trdemo', suffixes=['png','pdf'])
 
 ra,dec = 0.,0.
@@ -70,17 +72,23 @@ mod = np.maximum(mod, 1e-42)
 logmrange = np.max(np.log10(mod)) - np.min(np.log10(mod))
 logmrange = 6.
 
-def plog(mod):
+def clfpos():
 	plt.clf()
+	plt.gca().set_position([0.01, 0.01, 0.98, 0.98])
+
+def plog(mod):
+	clfpos()
 	lm = np.log10(mod + 1e-42)
 	mx = lm.max()
 	plt.imshow(lm-mx, vmax=0., vmin=-logmrange,
 			   extent=galext, **ima)
 	#plt.gray()
-	ticks = np.arange(-6, 1)
-	plt.colorbar(ticks=ticks,
-				 format=FixedFormatter(
-					 ['$10^{%i}$'%i for i in ticks[:-1]] + ['$1$']))
+	#ticks = np.arange(-6, 1)
+	#plt.colorbar(ticks=ticks,
+	#			 format=FixedFormatter(
+	#				 ['$10^{%i}$'%i for i in ticks[:-1]] + ['$1$']))
+	plt.xticks([])
+	plt.yticks([])
 	ps.savefig()
 
 plt.clf()
@@ -103,21 +111,11 @@ print G.shape
 dradec = np.dot(G, uv.T).T
 print 'G', G
 print dradec.shape
-#S = 4.
-#dradec *= S
 dra  = dradec[:,0]
 ddec = dradec[:,1]
-# xy = np.array([wcs1.positionToPixel(RaDecPos(ra + idra, dec + iddec))
-# 			   for idra,iddec in zip(dra,ddec)])
-# xx,yy = xy[:,0], xy[:,1]
-# 
-# plt.clf()
-# plt.plot(xx, yy, 'k-')
-# plt.axis(ax)
-# ps.savefig()
 
 S = 3600.
-plt.clf()
+clfpos()
 plt.plot(dra * S, ddec * S, 'k-')
 for i in [i0, i90]:
 	L = 0.3/4
@@ -127,6 +125,8 @@ for i in [i0, i90]:
[…]

### 2012-10-23 documents/mixtures/mixture_models.tex
caption tweak

@@
 % - DSTN: make figures showing use of the models in a real situation; see last figure
-% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
 % - HOGG: need Bundy reference
@@
 % - DSTN: is there anyone we need to ack on your side, or granting agencies?
+% - HOGG: Cite paper mentioned by Bosch?
 % - HOGG: should we make Machine Readable Tables for PASP?
@@
   is a pixel-convolved PSF, the representation on the realistic grid
-  is found simply by interpolating to the pixel centers the convolved
-  high-resolution model.\label{fig:example}}
+  is found simply by interpolating to the pixel centers the
+  mixture-component Gaussians.\label{fig:example}}
 \end{figure}

### 2012-10-23 documents/mixtures/mixture_models.tex
Address "DSTN" comments, fix affiliation, mention Bosch's work, cite and thank him, mention tractor example figs in the text

--- 2012-10-23 documents/mixtures/mixture_models.tex
@@
 % This file is part of the Tractor project.
-% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton).
+% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton->CMU).
 % All rights reserved.
@@
 % -----
-% - DSTN: check sentences I say about ``ten times faster''.
 % - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
-% - DSTN: make figures showing use of the models in a real situation; see last figure
-% - HOGG: does the Lupton paper give the profiles, or should we cite the code directly?
 % - HOGG: need Bundy reference
@@
 % - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
-% - DSTN: is there anyone we need to ack on your side, or granting agencies?
+% - HOGG: Cite paper mentioned by Bosch?
 % - HOGG: should we make Machine Readable Tables for PASP?
@@
 \altaffiltext{1}{To whom correspondence should be addressed; \texttt{david.hogg@nyu.edu}}
-\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Phyiscs, New York University, 4 Washington Place, New York, NY 10003, USA}
+\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
 \altaffiltext{3}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 \altaffiltext{4}{Princeton University Observatory, Princeton, NJ 08544, USA}
-\altaffiltext{5}{Department of Physics, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
+\altaffiltext{5}{McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
 
@@
 analysis software.  Some of the models we present have been used
-previously in tools for precise photometry (\citealt{bundy}); they
-could also be used to speed other image-fitting systems, like the very
-successful \project{GALFIT} (\citealt{galfit}).
+previously in tools for precise photometry (\citealt{bundy}), and are
+being used in the Large Synoptic Survey Telescope (LSST) analysis
+codes along with shapelet representations for the PSF
+(\citealt{bosch}).  These models could also be used to speed other
+image-fitting systems, like the very successful \project{GALFIT}
+(\citealt{galfit}).
 
@@
 
-DSTN... In \figurename~\ref{fig:example} we show...
+In \figurename~\ref{fig:example} we show how we are using these MoG
+approximations in \project{The Tractor}---a generative modeling
+framework for measuring astronomical objects---to render PSF-convolved
+galaxy images.
 
@@
 \acknowledgements It is a pleasure to thank
-      Brendon Brewer (Auckland)
+      Brendon Brewer (Auckland),
+      Jim Bosch (Princeton),
   and Kevin Bundy (IPMU)
@@
 Bendinelli,~O. \& Parmeggiani,~G., 1995, \aj, 109, 572
+\bibitem[Bosch(2010)]{bosch}
+Bosch,~J., 2010, \aj, 140, 870
 \bibitem[Bovy \etal(2011a)]{xd}
@@
 \setlength{\figurewidth}{0.33\textwidth}
-% DSTN:  This should be a 3-across by 2-down figure,
 \begin{tabular}{@{}c@{}c@{}}

### 2012-10-23 code/tractor-example-plots.py

@@ -87,6 +87,7 @@ def plog(mod):
 	#plt.colorbar(ticks=ticks,
 	#			 format=FixedFormatter(
 	#				 ['$10^{%i}$'%i for i in ticks[:-1]] + ['$1$']))
+	#plt.colorbar()
 	plt.xticks([])
 	plt.yticks([])
 	ps.savefig()
@@ -219,9 +220,10 @@ wcs4 = FitsWcs(Tan(ra, dec, W/2. + 0.5, H/2. + 0.5,
 				   pixscale, 0., 0., pixscale, W, H))
 im1.wcs = wcs4
 
+logmrange = 3
+
 mod = tractor.getModelImage(0)
 plog(mod)
-
 
 H,W = H/S, W/S
 print 'Pix size', W

### 2012-10-23 documents/mixtures/mixture_models.tex
use pdfs for figures

@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.png}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.png}\\
-\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.png}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.png}
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.pdf}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.pdf}\\
+\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.pdf}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.pdf}
 \caption{\textsl{top-left:} The true exp profile (thin black line),
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.png}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.png}\\
-\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.png}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.png}
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.pdf}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.pdf}\\
+\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.pdf}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.pdf}
 \caption{The dev profile and the best $M^{\dev}=10$ approximation.
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.png}%
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.png}\\
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.png}%
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.png}
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.pdf}%
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.pdf}\\
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.pdf}%
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.pdf}
 \caption{The lux and luv profiles and approximations.  The panels are
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.png}%
-\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.png}\\
-\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.png}%
-\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.png}
+\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.pdf}%
+\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.pdf}\\
+\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.pdf}%
+\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.pdf}
 \caption{Three ser profiles---with $n=2$, $3$, and $5$---and
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{mixtures_vs_K_exp.png}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_dev.png}\\
-\includegraphics[width=\figurewidth]{mixtures_vs_K_lux.png}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_luv.png}
+\includegraphics[width=\figurewidth]{mixtures_vs_K_exp}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_dev}\\
+\includegraphics[width=\figurewidth]{mixtures_vs_K_lux}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_luv}
 \caption{Comparisons of approximations.  \textsl{top-left:} The

### 2012-10-23 documents/mixtures/mixture_models.tex
notes to self

@@
-% This file is part of the Tractor project.
-% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton->CMU).
+% This file is part of The Tractor.
+% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton, CMU).
 % All rights reserved.
@@
 % -----
+% - HOGG: need Bundy reference
+% - HOGG: send to Rix and Peng and Lupton and Bundy for comments
+% - HOGG: submit!
 % - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
-% - HOGG: need Bundy reference
 % - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
-% - HOGG: send to Lupton and Bundy for comments
 % - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
-% - HOGG: Cite paper mentioned by Bosch?
-% - HOGG: should we make Machine Readable Tables for PASP?
+% - HOGG: should we make Machine Readable Tables for PASP?  Yes.
 
@@
 components) convolved with a MoG PSF model (with $K$ components)
-yields a MoG model image (with $[M\,K]$ components).
+yields a MoG model image (with $[M\,K]$ components).  In
+\figurename~\ref{fig:example} we show how we are using these MoG
+approximations in \project{The Tractor}---a generative modeling
+framework for measuring astronomical objects---to render PSF-convolved
+galaxy images.
 
@@
 fit and use pixel-convolved PSFs.
-
-In \figurename~\ref{fig:example} we show how we are using these MoG
-approximations in \project{The Tractor}---a generative modeling
-framework for measuring astronomical objects---to render PSF-convolved
-galaxy images.
 
@@
 \bibitem[Bundy \etal(2012)]{bundy}
-Bundy,~K.~??
+Bundy,~K. \etal, \aj, in preparation (arXiv:HOGG)
 \bibitem[Cappellari(2002)]{cappellari}

### 2012-10-23 documents/mixtures/mixture_models.tex
wording tweaks, reference

@@
 % -----
-% - HOGG: need Bundy reference
+% - HOGG: go through remining github issues and close or fix
 % - HOGG: send to Rix and Peng and Lupton and Bundy for comments
@@
 Under this choice, synthesis of a pixelized image involves only
-convolution with the PSF and sampling at the pixel centers.  That is,
-image synthesis (returned to at the end of this \documentname)
-consists of PSF convolution (an arithmetic operation on the two MoGs)
-followed by sampling (evaluation of the resulting MoG at the pixel
+convolution with the PSF and evaluation at the pixel centers.  That
+is, image synthesis (returned to at the end of this \documentname)
+consists of PSF convolution---an arithmetic operation on the two
+MoGs---followed by evaluation of the resulting MoG at the pixel
 centers).
@@
 \bibitem[Bundy \etal(2012)]{bundy}
-Bundy,~K. \etal, \aj, in preparation (arXiv:HOGG)
+Bundy,~K. \etal, \aj, in press
 \bibitem[Cappellari(2002)]{cappellari}

### 2012-10-23 documents/mixtures/mixture_models.tex
LSST text from Jim Bosch

--- 2012-10-23 documents/mixtures/mixture_models.tex
@@
 previously in tools for precise photometry (\citealt{bundy}), and are
-being used in the Large Synoptic Survey Telescope (LSST) analysis
-codes along with shapelet representations for the PSF
+being used in the Large Synoptic Survey Telescope (LSST) prototype
+galaxy photometry pipeline (\citealt{shaw}), in which these models are
+convolved with a shapelet representation of the PSF; because shapelets
+are simply polynomial perturbations on a Gaussian, they can also be
+convolved analytically with mixture-of-Gaussian galaxy models
 (\citealt{bosch}).  These models could also be used to speed other
@@
 S{\'e}rsic,~J.~L., 1963, Boletin de la Asociacion Argentina de Astronomia La Plata Argentina, 6, 41
+\bibitem[Shaw(2012)]{shaw}
+Shaw,~R.~A., ed., 2012, LSST Data Challenge Handbook (Version 2.0; Tucson, AZ: LSST Corp.)
 \end{thebibliography}
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.pdf}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.pdf}
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.png}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.png}
 \caption{\textsl{top-left:} The true exp profile (thin black line),
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.pdf}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.pdf}
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.png}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.png}
 \caption{The dev profile and the best $M^{\dev}=10$ approximation.
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.pdf}%
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.pdf}
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.png}%
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.png}\\
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.png}%
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.png}
 \caption{The lux and luv profiles and approximations.  The panels are
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.pdf}%
-\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.pdf}%
-\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.pdf}
+\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.png}%
+\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.png}\\
+\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.png}
 \caption{Three ser profiles---with $n=2$, $3$, and $5$---and
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{mixtures_vs_K_exp}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_dev}\\
-\includegraphics[width=\figurewidth]{mixtures_vs_K_lux}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_luv}
+\includegraphics[width=\figurewidth]{mixtures_vs_K_exp.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_dev.png}\\
+\includegraphics[width=\figurewidth]{mixtures_vs_K_lux.png}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_luv.png}
 \caption{Comparisons of approximations.  \textsl{top-left:} The
@@
   is a pixel-convolved PSF, the representation on the realistic grid
-  is found simply by interpolating to the pixel centers the
-  mixture-component Gaussians.\label{fig:example}}
+  is found simply by interpolating to the pixel centers the convolved
+  high-resolution model.\label{fig:example}}
 \end{figure}

### 2012-10-23 documents/mixtures/mixture_models.tex
todo done
don't use the word sampling

--- 2012-10-23 documents/mixtures/mixture_models.tex
@@
 morphologically diverse objects (as we use them in \citealt{bundy}).
-They are sometimes used to distinguish disk and bulge light; these
-uses are very prone to over-interpretation: There is no theoretical
-argument and only weak observational arguments that the
-rotation-supported parts of galaxies are always exponential and that
-the kinematically hot components are de~Vaucouleurs.
+More radically, these profiles are sometimes used to distinguish disk
+and bulge light; these uses are very prone to over-interpretation:
+There is no theoretical argument and only weak observational arguments
+that the rotation-supported parts of galaxies are always exponential
+and that the kinematically hot components are always more
+de-Vaucouleurs-like.
 

### 2012-10-23 documents/mixtures/mixture_models.tex
sersic index discussion; typos

@@
 % - HOGG: go through remining github issues and close or fix
-% - HOGG: send to Rix and Peng and Lupton and Bundy for comments
+% - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
 % - HOGG: submit!
@@
 MoG, which is not uncommon (for example, the \project{Sloan Digital
-  Sky Survey} imaging pipelines make approximate MoG PSF models for
-every imaging field; \citealt{lupton}).
+  Sky Survey} imaging pipelines described by \citealt{lupton} make
+approximate MoG PSF models for every imaging field).
 
@@
 of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
-models are small modifications of these (\citealt{lupton}); details
+models are small modifications of these; details
 below.  So our goal here is to provide replacements for these models
@@
 where we have given the constant $\alpha^{\ser}(n)$ for just a few
-values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles give
-the $n=1$ and $n=4$ values).
+values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles given
+above provide values for $n=1$ and $n=4$.)
 
-The \sdss\ pipelines make use of modified profiles, which have been
-truncated smoothly at large radius and (in the case of the
-de~Vaucouleurs profile) ``softened'' at the center (\citealt{lupton}).
-The \sdss\ form of the exponential (lux) profile is
+The \sdss\ pipelines (\citealt{lupton}) make use of modified profiles,
+which have been truncated smoothly at large radius and (in the case of
+the de~Vaucouleurs profile) ``softened'' at the center.  The
+\sdss\ form of the exponential (lux) profile is
 \begin{eqnarray}\displaystyle
@@
 
-The value of these MoG galaxy components comes when they are to be
+In \figurename~\ref{fig:ser}, we show the dependence of amplitudes
+$a^{\ser}_m(n)$ and variances $v^{\ser}_m(n)$ on the ser index $n$.
+There is clearly continuity; a valuable follow-up project would be to
+give expressions for the amplitudes and variances as a function of ser
+index $n$.  In the absence of cleverness our advice would be to make
+use of smooth interpolation.
+
+The value of these MoG approximations comes when they are to be
 convolved with a PSF (usually in fact a pixel-convolved PSF) that is
@@
 an affine transformation $\affine_g$ that takes a circle to the
-relevant ellipse (and is a therefore a general representation of an
+relevant ellipse (and is therefore a general representation of an
 ellipse; it is also the matrix square root of the symmetric variance
@@
   \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser}_m$
-  and root-variances $\sqrt{v^{\ser}_m}$ on ser index $n$.}
+  and root-variances $\sqrt{v^{\ser}_m}$ on ser index $n$.\label{fig:ser}}
 \end{figure}

### 2012-10-23 documents/mixtures/mixture_models.tex
for
of

@@
 PSF and the second with the square (or worse) pixel.  Our advice: Only
-fit and use pixel-convolved PSFs.
+fit for and use pixel-convolved PSFs.
 
 If your PSF is not in MoG form, it is still the case that convolution
-of a MoG approximation to a dev (say) profile will in general be
+of a MoG approximation of a dev (say) profile will in general be
 easier than convolution of the original dev profile.  The reason is

### 2012-10-23 documents/mixtures/mixture_models.tex

@@
 % - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
-% - DSTN: can we be more specific about how Bosch is using them? I would like to cite users.
 % - HOGG: should we make Machine Readable Tables for PASP?  Yes.

### 2012-10-24 documents/mixtures/mixture_models.tex
address
dstn comments

@@
 \begin{eqnarray}\displaystyle
-Q^{\ser}(\spos\given n) &\equiv& \exp(-\alpha^{\ser}(n)\,[|\spos|^{1/n} - 1])
+Q^{\ser(n)}(\spos) &\equiv& \exp(-\alpha^{\ser(n)}\,[|\spos|^{1/n} - 1])
 \\
-\{\alpha^{\ser}(2), \alpha^{\ser}(3), \alpha^{\ser}(5)\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
+\{\alpha^{\ser(2)}, \alpha^{\ser(3)}, \alpha^{\ser(5)}\} &\equiv& \{3.67206075, 5.67016119, 9.66871461\}
 \quad ,
 \end{eqnarray}
-where we have given the constant $\alpha^{\ser}(n)$ for just a few
+where we have given the constant $\alpha^{\ser(n)}$ for just a few
 values of $n$ (\citealt{ciotti}; the $\exp$ and $\dev$ profiles given
@@
 In \figurename~\ref{fig:ser}, we show the dependence of amplitudes
-$a^{\ser}_m(n)$ and variances $v^{\ser}_m(n)$ on the ser index $n$.
+$a^{\ser(n)}_m$ and variances $v^{\ser(n)}_m$ on the ser index $n$.
 There is clearly continuity; a valuable follow-up project would be to
@@
 ellipse (in appropriate units) and $\uv_1$ and $\uv_2$ are the
-eigenvectors on the sky pointing in the major-axis and minor-axis
-directions respectively.  Implicitly all vectors are two-dimensional
-column vectors, and $\affine_g$ is a $2\times 2$ affine transformation
-matrix that contains the ``shape'' (position angle, major-axis, and
-ellipticity) information about the galaxy.  The $\dev$, $\luv$, and
-$\lux$ cases are all similar.  Note the important and key result of
-this \documentname, to wit, that a MoG galaxy model (with $M$
-components) convolved with a MoG PSF model (with $K$ components)
-yields a MoG model image (with $[M\,K]$ components).  In
+eigenvectors in image coordinates pointing in the major-axis and
+minor-axis directions respectively.  Implicitly all vectors are
+two-dimensional column vectors, and $\affine_g$ is a $2\times 2$
+affine transformation matrix that contains the ``shape'' (position
+angle, major-axis, and ellipticity) information about the galaxy.  The
+$\dev$, $\luv$, and $\lux$ cases are all similar.  Note the important
+and key result of this \documentname, to wit, that a MoG galaxy model
+(with $M$ components) convolved with a MoG PSF model (with $K$
+components) yields a MoG model image (with $[M\,K]$ components).  In
 \figurename~\ref{fig:example} we show how we are using these MoG
@@
   equivalent to those in the top-right of \figurename~\ref{fig:exp}.
-  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser}_m$
-  and root-variances $\sqrt{v^{\ser}_m}$ on ser index $n$.\label{fig:ser}}
+  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser(n)}_m$
+  and root-variances $\sqrt{v^{\ser(n)}_m}$ on ser index $n$.\label{fig:ser}}
 \end{figure}

### 2012-10-24 documents/mixtures/mixture_models.tex
more references

@@
 \altaffiltext{4}{Princeton University Observatory, Princeton, NJ 08544, USA}
-\altaffiltext{5}{McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15213, USA}
+\altaffiltext{5}{McWilliams Center for Cosmology, Carnegie Mellon University, 5000 Forbes Avenue, Pittsburgh, PA 15213, USA}
+%\altaffiltext{6}{McWilliams Postdoctoral Fellow}
 
@@
 More radically, these profiles are sometimes used to distinguish disk
-and bulge light; these uses are very prone to over-interpretation:
-There is no theoretical argument and only weak observational arguments
-that the rotation-supported parts of galaxies are always exponential
-and that the kinematically hot components are always more
-de-Vaucouleurs-like.
+and bulge light (as in \citealt{bulgedisk} and references cited
+therein); these uses are prone to over-interpretation: There is no
+theoretical argument and only weak observational arguments that the
+rotation-supported parts of galaxies are always exponential and that
+the kinematically hot components are always more de-Vaucouleurs-like.
 
@@
 
-Along those lines, it is important in any image-modeling situation to
-think of the PSF as the \emph{pixel-convolved} point-spread function.
-Under this choice, synthesis of a pixelized image involves only
-convolution with the PSF and evaluation at the pixel centers.  That
-is, image synthesis (returned to at the end of this \documentname)
-consists of PSF convolution---an arithmetic operation on the two
-MoGs---followed by evaluation of the resulting MoG at the pixel
-centers).
+On the subject of PSF-convolution, it is important in any
+image-modeling situation to think of the PSF as the
+\emph{pixel-convolved} point-spread function.  Under this choice,
+synthesis of a pixelized image involves only convolution with the PSF
+and evaluation at the pixel centers.  That is, image synthesis
+(returned to at the end of this \documentname) consists of PSF
+convolution---an arithmetic operation on the two MoGs---followed by
+evaluation of the resulting MoG at the pixel centers).
 
@@
 of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
-models are small modifications of these; details
-below.  So our goal here is to provide replacements for these models
-in order to improve the performance of galaxy image modeling and
-analysis software.  Some of the models we present have been used
-previously in tools for precise photometry (\citealt{bundy}), and are
-being used in the Large Synoptic Survey Telescope (LSST) prototype
-galaxy photometry pipeline (\citealt{shaw}), in which these models are
-convolved with a shapelet representation of the PSF; because shapelets
-are simply polynomial perturbations on a Gaussian, they can also be
-convolved analytically with mixture-of-Gaussian galaxy models
-(\citealt{bosch}).  These models could also be used to speed other
-image-fitting systems, like the very successful \project{GALFIT}
-(\citealt{galfit}).
+models are small modifications of these; details below.  So our goal
+here is to provide replacements for these models in order to improve
+the performance of galaxy image modeling and analysis software.  Some
+of the models we present have been used previously in tools for
+precise photometry (\citealt{bundy}), and are being used in the Large
+Synoptic Survey Telescope (LSST) prototype galaxy photometry pipeline
+(\citealt{shaw}), in which these models are convolved with a shapelet
+representation of the PSF; because shapelets are simply polynomial
+perturbations on a Gaussian, they can also be convolved analytically
+with mixture-of-Gaussian galaxy models (\citealt{bosch}).  These
+models could also be used to speed other image-fitting systems, like
+the very successful \project{GALFIT} (\citealt{galfit}).  An
+alternative to making MoG profiles is to find exact analytic
+expressions for certain kinds of convolutions.  There \emph{are} some
+analytic results for PSF-convolved S\'ersic models but they involve
+some non-trivial series and special-function expressions
+(\citealt{trujillo}); we won't come back to these again but they might
+be very useful in many situations.
 
@@
 Shaw,~R.~A., ed., 2012, LSST Data Challenge Handbook (Version 2.0; Tucson, AZ: LSST Corp.)
+\bibitem[Simard et al.(2011)]{bulgedisk}
+Simard,~L., Mendel,~J.~T., Patton,~D.~R., Ellison,~S.~L., \& McConnachie,~A.~W., 2011, \apjs, 196, 11
+\bibitem[Trujillo \etal(2001)]{trujillo}
+Trujillo,~I., Aguerri,~J.~A.~L., Cepa,~J., \& Guti{\'e}rrez,~C.~M., 2001, \mnras, 321, 269 
 \end{thebibliography}

### 2012-10-24 documents/mixtures/mixture_models.tex
figurewidth
tweak

@@
 % -----
-% - HOGG: go through remining github issues and close or fix
 % - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
 % - HOGG: submit!
+% - HOGG: post on arXiv
 % - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
@@
 angle, major-axis, and ellipticity) information about the galaxy.  The
-$\dev$, $\luv$, and $\lux$ cases are all similar.  Note the important
-and key result of this \documentname, to wit, that a MoG galaxy model
-(with $M$ components) convolved with a MoG PSF model (with $K$
-components) yields a MoG model image (with $[M\,K]$ components).  In
-\figurename~\ref{fig:example} we show how we are using these MoG
-approximations in \project{The Tractor}---a generative modeling
-framework for measuring astronomical objects---to render PSF-convolved
-galaxy images.
+$\dev$, $\ser(n)$, $\luv$, and $\lux$ cases are all essentially the
+same.  Note the important and key result of this \documentname, to
+wit, that a MoG galaxy model (with $M$ components) convolved with a
+MoG PSF model (with $K$ components) yields a MoG model image (with
+$[M\,K]$ components).  In \figurename~\ref{fig:example} we show how we
+are using these MoG approximations in \project{The Tractor}---a
+generative modeling framework for measuring astronomical objects---to
+render PSF-convolved galaxy images.
 
@@
 \begin{figure}
-\setlength{\figurewidth}{0.33\textwidth}
+\setlength{\figurewidth}{0.25\textwidth}
 \begin{tabular}{@{}c@{}c@{}}

### 2012-10-26 documents/mixtures/mixture_models.tex
move todo list to github
now with eps files; done for submission; about to revert

@@
 
-% to-do
-% -----
-% - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
-% - HOGG: submit!
-% - HOGG: post on arXiv
-% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
-% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
-% - HOGG: should we make Machine Readable Tables for PASP?  Yes.
-
-\documentclass[12pt,pdftex,preprint]{aastex}
+\documentclass[12pt,preprint]{aastex}
 \usepackage{amssymb,amsmath,mathrsfs}
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.pdf}%
-\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.pdf}
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile.eps}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_profile_log.eps}\\
+\includegraphics[width=\figurewidth]{exp_K06_MR08_residual.eps}%
+\includegraphics[width=\figurewidth]{exp_K06_MR08_fractional_log.eps}
 \caption{\textsl{top-left:} The true exp profile (thin black line),
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.pdf}%
-\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.pdf}
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile.eps}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_profile_log.eps}\\
+\includegraphics[width=\figurewidth]{dev_K10_MR08_residual.eps}%
+\includegraphics[width=\figurewidth]{dev_K10_MR08_fractional_log.eps}
 \caption{The dev profile and the best $M^{\dev}=10$ approximation.
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.pdf}%
-\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.pdf}%
-\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.pdf}
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile.eps}%
+\includegraphics[width=\figurewidth]{lux_K06_MR04_profile_log.eps}\\
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile.eps}%
+\includegraphics[width=\figurewidth]{luv_K08_MR08_profile_log.eps}
 \caption{The lux and luv profiles and approximations.  The panels are
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.pdf}%
-\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.pdf}\\
-\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.pdf}%
-\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.pdf}
+\includegraphics[width=\figurewidth]{ser2_K10_MR08_profile_log.eps}%
+\includegraphics[width=\figurewidth]{ser3_K10_MR08_profile_log.eps}\\
+\includegraphics[width=\figurewidth]{ser5_K10_MR08_profile_log.eps}%
+\includegraphics[width=\figurewidth]{mixtures_vs_model_K10.eps}
 \caption{Three ser profiles---with $n=2$, $3$, and $5$---and
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{mixtures_vs_K_exp}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_dev}\\
-\includegraphics[width=\figurewidth]{mixtures_vs_K_lux}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_luv}
+\includegraphics[width=\figurewidth]{mixtures_vs_K_exp.eps}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_dev.eps}\\
+\includegraphics[width=\figurewidth]{mixtures_vs_K_lux.eps}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_luv.eps}
 \caption{Comparisons of approximations.  \textsl{top-left:} The
@@
 \begin{tabular}{@{}c@{}c@{}}
-\includegraphics[width=\figurewidth]{trdemo-01} &
-\includegraphics[width=\figurewidth]{trdemo-02} \\
-\includegraphics[width=\figurewidth]{trdemo-03} &
-\includegraphics[width=\figurewidth]{trdemo-06} \\
-\includegraphics[width=\figurewidth]{trdemo-08} &
-\includegraphics[width=\figurewidth]{trdemo-09}
+\includegraphics[width=\figurewidth]{trdemo-01.eps} &
+\includegraphics[width=\figurewidth]{trdemo-02.eps} \\
+\includegraphics[width=\figurewidth]{trdemo-03.eps} &
+\includegraphics[width=\figurewidth]{trdemo-06.eps} \\
+\includegraphics[width=\figurewidth]{trdemo-08.eps} &
+\includegraphics[width=\figurewidth]{trdemo-09.eps}
 \end{tabular}

### 2012-10-26 documents/mixtures/pdf2eps.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+for file in $(ls *.pdf);
+do
+prefix=$(basename $file .pdf)
+convert -density 300 $file $prefix.png
+convert $prefix.png $prefix.eps
+done

### 2013-01-07 code/optimize_mixture_profiles.py
comment tweaks

@@ -1,6 +1,6 @@
 '''
 This file is part of the Tractor project.
-Copyright 2011 David W. Hogg.
+Copyright 2011, 2012 David W. Hogg.
 
 ### bugs:
 - When using chi-squared-like badness, we should use levmar not bfgs!

### 2013-01-07 documents/mixtures/response_to_referee.txt
started responding to referee

Dear Editor,

This was a very useful referee report and we have revised the paper
accordingly.  I think the referee comments have improved the paper
substantially.

David W. Hogg (on behalf of the authors)

-----

In what follows, the referee comments are indented and our responses
are not indented.

   1. Does the article contain significant new results and/or analysis
   and reflect sufficiently high scientific standards to warrant its
   publication in PASP?

   This paper gives a particularly lucid, concise and clear
   description of a relatively old idea. As such, it represents a
   significant contribution to the literature as a pedagogical
   reference. It should be published but see 4. below.

We agree with the referee that the idea is old but we don't agree with
any implication that the results shown are old.  All references we can
find in the literature show *how* to do mixture-of-Gaussian fitting in
real imaging, but don't give mixture-of-Gaussian approximations to
known profiles.  This paper, in that sense, is more of an "applied
math meets astronomy" paper than the previous work.  We will return to
this below, where we *do* make changes to the text in response to this
general concern of the referee.

   2. Is the paper written with the maximum conciseness compatible
   with accuracy and clarity?

   I thoroughly enjoyed reading the paper for its refreshing
   conciseness and clarity.

Thanks!

   3. Could the order or presentation or English be improved?  Editing
   of English and typographic errors will be corrected by the Journal
   copyediting staff and you do not need to specify these in your
   report.  But if there are more general problems with presentation
   or English please cite this in your report.

   Contractions (don't,  doesn't) should not be used.

   On page 8, paragraph 3: "If your PSF is not in MoG form ..." should read "If the
    PSF is not in MoG form ..."

   4. Do you have any comments or criticisms that may be helpful to the author(s)
   to improve or correct the paper? In cases that may be ambiguous, please specify
   whether you consider the suggested changes to be mandatory for publication or
   advisory.

   GENERAL COMMENT: Although I do think that this "Note" would stand on its own as
   a separate publication, I would nonetheless strongly suggest that the authors
   consider merging it into the forthcoming Lang et al. paper on The Tractor. This
   Note appears to be an "ad" for The Tractor. In the words of the authors, the main
   point of the Note is to emphasize the performance gain that the MoG approach
   would bring to the analysis of truly large galaxy datasets.  However, the current
   draft does not show any analysis of real data to support this performance claim.
   Other methods have been proposed in the past with similar promises of significant
   performance improvements (e.g., fast nonlinear regression via eigenimages,
   Brigham et al.). They looked promising on paper but were found to exhibit unexpected
   biases when applied to real data. I would hate to see this happen to MoGs. This
   is why it would be really nice to have the beautiful description of MoGs
   presented in this Note side-by-side with some tests on real data in the same paper.

   SPECIFIC COMMENTS:

   Page 2, paragraph 3: "There is no theoretical argument and only weak observational
   arguments that the rotation-supported parts of galaxies are always exponential
   and that the kinematically hot components are always more de Vaucouleurs-like".
   Although your warning about over-interpretating exp and dev components using
   bulges and disks is well taken (and others have given the same warning I might
   add), I think your statement above is too strong. On the theoretical side, some
   works (e.g., Brooks et al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410,
   1391) have performed photometric decompositions of their model galaxies to see how
   photometric components mapped onto kinematical ones. The agreement is good and
   does suggest that it is often a reasonable mapping. It is certainly not a "radical"
   assumption to make. Also, while it is true that some dispersion-supported
   galaxies such as dwarf ellipticals have exponential profiles, they certainly do not
    make up the bulk of the galaxy population in large surveys. Finally, I note the
    recent ATLAS3D paper by Krajnovic et al. 2012, arXiv:1210.816 in which it was
   found that "early-type" galaxies with significant rotational support also
   exhibited significant exponential ("disk") components.

   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k say, to
   avoid confusion with the \alpha's in the following equations.

   Page 6, first line: "The half-light inverse radius parameters \alpha^{exp} and \
   alpha^{dev} -- and the softening and cutoff radius parameters -- are taken from
   the SDSS codebase." I think you meant to write \alpha^{lux} and \alpha^{luv} here.

   Page 6, paragraph 2: "Although there is an analytic result for the dev profile,
   numerical integration of the concentrated profiles dev and luv to determine
   total fluxes can be challenging."  Why are you flagging the numerical integration of
    the dev profile as a challenge given that there is an analytic expression for it?

   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler divergence"

   Page 6 - Definition of "best": The authors should really have been "bolder" by
   insisting on the use of an information-theoretic measure here rather than the
   "traditional" chi-squared. The MoGs are a natural fit to an image entropy measure
   because maximizing entropy, say, would naturally lead to minimizing the number of
    components in a MoG. Chi-squared is also biased at low signal-to-noise ratios.
   Please consider.

   Page 6: Why is "badness" computed in a one-dimensional numerical integral? For
   the sake of computational speed? What happens when dealing with heavily blended
   objects? Where is the center for the 1D integral then? Please explain. A general
   explanation of how object deblending is performed with MoGs would be welcome.

   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv = 8 -- are good
    compromises between mixture complexity and quality of fit" warrants more
   explanation. Everytime I read about using MoGs for galaxy fitting, I have the same
   reaction: it needs a large number of highly redundant to model something a single
   exponential profile. Fans of the eigenimage decompositions are probably appalled!
    Granted the mathematical simplicity of MoGs can make up for this by enabling
   faster fitting times, but it is definitely an un-appealing aspect of the method.
   It would be helpful to comment on this to counter a reaction that I think may be
   shared by others. More importantly, I do not quite understand how you decided on
    M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for lux has clearly
   converged and there is no gain in going higher than M=6. I cannot say the same
   thing for the luv MoG. Badness goes down by a factor of 16 going from M=6 to M=8
   and then goes down by a factor of 5 going from M=8 to M=10. How can you claim
   that your badness has "converged" for M=8 then? A quantitative threshold and its
   rationale are clearly needed here to justify the "good compromises" here. Please
   be more quantitative or clarify if I misunderstood something.

   Equation 19: I did not quite understand your distinction between the \Delta X
   vector here and the x vector in equation 2. Please clarify.

   ==============================================

### 2013-01-07 documents/mixtures/mixture_models.tex
revert to pdf
typo
contractions

--- 2012-10-24 documents/mixtures/mixture_models.tex
@@
 
-% to-do
-% -----
-% - HOGG: send to Bertin and Rix and Peng and Lupton and Bundy for comments
-% - HOGG: submit!
-% - HOGG: post on arXiv
-% - HOGG: can we drop the regularization in the badness and still succeeed?  If so, let's do that (and drop it in the text).
-% - HOGG: make what I say about M=6 lux and M=8 lux true -- ie, edit the Tractor code!
-% - HOGG: should we make Machine Readable Tables for PASP?  Yes.
-
 \documentclass[12pt,pdftex,preprint]{aastex}
+% \documentclass[12pt,preprint]{aastex}
 \usepackage{amssymb,amsmath,mathrsfs}
@@
 probability distribution functions and it becomes remarkable that we
-don't do everything we do in terms of Gaussians.
+do not do everything we do in terms of Gaussians.
 
@@
 de~Vaucouleurs profiles, the models presented here will improve code
-performance.  That doesn't mean that doing such fitting is a Good
+performance.  That does not mean that doing such fitting is a Good
 Idea.  These profiles are effective models for galaxies at low
@@
 some non-trivial series and special-function expressions
-(\citealt{trujillo}); we won't come back to these again but they might
-be very useful in many situations.
+(\citealt{trujillo}); we will not come back to these again but they
+might be very useful in many situations.
 
@@
 $10^{-3}$ of the best-fit badness) times the sum of the variances
-$v_m$ for regularization.  In practice, this term doesn't have much
+$v_m$ for regularization.  In practice, this term does not have much
 effect and could be dropped.
@@
 
-If your PSF is not in MoG form, it is still the case that convolution
-of a MoG approximation of a dev (say) profile will in general be
-easier than convolution of the original dev profile.  The reason is
-that convolution of a Gaussian with any PSF is fast (indeed most
-image-processing languages have such functions built in); the
+If the PSF is not represented in MoG form, it is still the case that
+convolution of a MoG approximation of a dev (say) profile will in
+general be easier than convolution of the original dev profile.  The
+reason is that convolution of a Gaussian with any PSF is fast (indeed
+most image-processing languages have such functions built in); the
 PSF-convolved profile becomes in this case just a mixture of
@@
 \begin{figure}
-\includegraphics[width=\figurewidth]{mixtures_vs_K_exp}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_dev}\\
-\includegraphics[width=\figurewidth]{mixtures_vs_K_lux}%
-\includegraphics[width=\figurewidth]{mixtures_vs_K_luv}
+\includegraphics[width=\figurewidth]{mixtures_vs_K_exp.pdf}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_dev.pdf}\\
+\includegraphics[width=\figurewidth]{mixtures_vs_K_lux.pdf}%
+\includegraphics[width=\figurewidth]{mixtures_vs_K_luv.pdf}
 \caption{Comparisons of approximations.  \textsl{top-left:} The
@@
 \begin{tabular}{@{}c@{}c@{}}
-\includegraphics[width=\figurewidth]{trdemo-01} &
-\includegraphics[width=\figurewidth]{trdemo-02} \\
-\includegraphics[width=\figurewidth]{trdemo-03} &
-\includegraphics[width=\figurewidth]{trdemo-06} \\
-\includegraphics[width=\figurewidth]{trdemo-08} &
-\includegraphics[width=\figurewidth]{trdemo-09}
+\includegraphics[width=\figurewidth]{trdemo-01.pdf} &
+\includegraphics[width=\figurewidth]{trdemo-02.pdf} \\
+\includegraphics[width=\figurewidth]{trdemo-03.pdf} &
+\includegraphics[width=\figurewidth]{trdemo-06.pdf} \\
+\includegraphics[width=\figurewidth]{trdemo-08.pdf} &
+\includegraphics[width=\figurewidth]{trdemo-09.pdf}
 \end{tabular}

### 2013-01-24 documents/mixtures/response_to_referee.txt
contractions
notes to self

@@
 
-In what follows, the referee comments are indented and our responses
-are not indented.
+We very much appreciate the thorough and sensible referee comments;
+they have led to substantial improvements in the paper.  In what
+follows, the referee comments are indented and our responses are not
+indented.
 
@@
 
-We agree with the referee that the idea is old but we don't agree with
-any implication that the results shown are old.  All references we can
+We agree with the referee that the idea of using Gaussians to fit
+galaxies is old but we don't agree with any implication that the
+method or applied math results shown are old.  All references we can
 find in the literature show *how* to do mixture-of-Gaussian fitting in
@@
 general concern of the referee.
+
+DSTN NOTES:.. math convenience... pre-compiling in Photo... adaptive
+sampling challenges... Bundy and Marshall uses... programming
+complexity... These points need to go in the paper, not just this
+response!
 
@@
 
+Replaced all contractions.
+
    On page 8, paragraph 3: "If your PSF is not in MoG form ..." should read "If the
     PSF is not in MoG form ..."
+
+Fixed to "If the PSF is not represented in MoG form ...".
 
@@
    presented in this Note side-by-side with some tests on real data in the same paper.
+
+[ let's do this and include it ]
 
@@
 
-   ==============================================
+In other news, we also made the following changes to the text, in
+response to other comments and issues:
+
+- we did foo
+
+- we did bar

### 2013-02-10 documents/mixtures/response_to_referee.txt
notes to self
one more paragraph of response to referee

@@
 math meets astronomy" paper than the previous work.  We will return to
-this below, where we *do* make changes to the text in response to this
-general concern of the referee.
+this below, but we *did* make changes to the text in response to this
+general concern of the referee: We added sentences in the introductory
+paragraphs and at the end of the paper to emphasize the difficulty of
+rendering the profiles without these approximations.
 
-DSTN NOTES:.. math convenience... pre-compiling in Photo... adaptive
-sampling challenges... Bundy and Marshall uses... programming
-complexity... These points need to go in the paper, not just this
-response!
+To be more specific: we added some sentences to emphasize that the
+currently used methods for rendering profiles are effectively mixtures
+of delta functions (the profile is sampled at a finite set of
+locations) and rendering requires as many PSF renderings as there are
+sample points.  This is a huge computation.
+
+We also added an introductory paragraph that points out that the code
+complexity required to do rendering correctly is large and this leads
+to engineering overheads in testing and correctness that go way beyond
+the simple numerical speed losses.
+
+[should we also add a late paragraph re-iterating and showing a comparison?]
 
@@
 
-   On page 8, paragraph 3: "If your PSF is not in MoG form ..." should read "If the
-    PSF is not in MoG form ..."
+   On page 8n, paragraph 3: "If your PSF is not in MoG form ..." should
+   read "If the PSF is not in MoG form ..."
 
@@
 
-   4. Do you have any comments or criticisms that may be helpful to the author(s)
-   to improve or correct the paper? In cases that may be ambiguous, please specify
-   whether you consider the suggested changes to be mandatory for publication or
-   advisory.
+   4. Do you have any comments or criticisms that may be helpful to
+   the author(s) to improve or correct the paper? In cases that may be
+   ambiguous, please specify whether you consider the suggested
+   changes to be mandatory for publication or advisory.
 
-   GENERAL COMMENT: Although I do think that this "Note" would stand on its own as
-   a separate publication, I would nonetheless strongly suggest that the authors
-   consider merging it into the forthcoming Lang et al. paper on The Tractor. This
-   Note appears to be an "ad" for The Tractor. In the words of the authors, the main
-   point of the Note is to emphasize the performance gain that the MoG approach
-   would bring to the analysis of truly large galaxy datasets.  However, the current
-   draft does not show any analysis of real data to support this performance claim.
-   Other methods have been proposed in the past with similar promises of significant
-   performance improvements (e.g., fast nonlinear regression via eigenimages,
-   Brigham et al.). They looked promising on paper but were found to exhibit unexpected
-   biases when applied to real data. I would hate to see this happen to MoGs. This
-   is why it would be really nice to have the beautiful description of MoGs
-   presented in this Note side-by-side with some tests on real data in the same paper.
+   GENERAL COMMENT: Although I do think that this "Note" would stand
+   on its own as a separate publication, I would nonetheless strongly
+   suggest that the authors consider merging it into the forthcoming
+   Lang et al. paper on The Tractor. This Note appears to be an "ad"
+   for The Tractor. In the words of the authors, the main point of the
+   Note is to emphasize the performance gain that the MoG approach
+   would bring to the analysis of truly large galaxy datasets.
+   However, the current draft does not show any analysis of real data
+   to support this performance claim.  Other methods have been
+   proposed in the past with similar promises of significant
+   performance improvements (e.g., fast nonlinear regression via
+   eigenimages, Brigham et al.). They looked promising on paper but
+   were found to exhibit unexpected biases when applied to real
+   data. I would hate to see this happen to MoGs. This is why it would
+   be really nice to have the beautiful description of MoGs presented
+   in this Note side-by-side with some tests on real data in the same
+   paper.
 
-[ let's do this and include it ]
+As noted above, we have added sentences in one paragraph and one new
+paragraph that makes it clearer what the performance gains are.
+
+[ should we also get quantitative at the end? ]
 
@@
 
-   Page 2, paragraph 3: "There is no theoretical argument and only weak observational
-   arguments that the rotation-supported parts of galaxies are always exponential
-   and that the kinematically hot components are always more de Vaucouleurs-like".
-   Although your warning about over-interpretating exp and dev components using
-   bulges and disks is well taken (and others have given the same warning I might
-   add), I think your statement above is too strong. On the theoretical side, some
-   works (e.g., Brooks et al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410,
-   1391) have performed photometric decompositions of their model galaxies to see how
-   photometric components mapped onto kinematical ones. The agreement is good and
-   does suggest that it is often a reasonable mapping. It is certainly not a "radical"
-   assumption to make. Also, while it is true that some dispersion-supported
-   galaxies such as dwarf ellipticals have exponential profiles, they certainly do not
-    make up the bulk of the galaxy population in large surveys. Finally, I note the
-    recent ATLAS3D paper by Krajnovic et al. 2012, arXiv:1210.816 in which it was
-   found that "early-type" galaxies with significant rotational support also
-   exhibited significant exponential ("disk") components.
+   Page 2, paragraph 3: "There is no theoretical argument and only
+   weak observational arguments that the rotation-supported parts of
+   galaxies are always exponential and that the kinematically hot
+   components are always more de Vaucouleurs-like".  Although your
+   warning about over-interpretating exp and dev components using
+   bulges and disks is well taken (and others have given the same
+   warning I might add), I think your statement above is too
+   strong. On the theoretical side, some works (e.g., Brooks et
+   al. 2011, ApJ, 728, 51; Agertz et al. 2011, MNRAS, 410, 1391) have
+   performed photometric decompositions of their model galaxies to see
+   how photometric components mapped onto kinematical ones. The
+   agreement is good and does suggest that it is often a reasonable
+   mapping. It is certainly not a "radical" assumption to make. Also,
+   while it is true that some dispersion-supported galaxies such as
+   dwarf ellipticals have exponential profiles, they certainly do not
+   make up the bulk of the galaxy population in large
+   surveys. Finally, I note the recent ATLAS3D paper by Krajnovic et
+   al. 2012, arXiv:1210.816 in which it was found that "early-type"
+   galaxies with significant rotational support also exhibited
+   significant exponential ("disk") components.
+
+Point well taken!  We softened this paragraph so that it no longer
+makes any claim on this point; we now only attempt to induce some
+feelings about conservativeness.
 

### 2013-02-10 documents/mixtures/mixture_models.tex
preample trivialities
notes to self
one more paragraph of response to referee

@@
 % This file is part of The Tractor.
-% Copyright 2012 David W. Hogg (NYU) and Dustin Lang (Princeton, CMU).
+% Copyright 2012, 2013 David W. Hogg (NYU) and Dustin Lang (Princeton, CMU).
 % All rights reserved.
 
+% to-do items
+% -----------
+% - respond to referee
+% - resubmit
+
 \documentclass[12pt,pdftex,preprint]{aastex}
-% \documentclass[12pt,preprint]{aastex}
 \usepackage{amssymb,amsmath,mathrsfs}
@@
 and bulge light (as in \citealt{bulgedisk} and references cited
-therein); these uses are prone to over-interpretation: There is no
-theoretical argument and only weak observational arguments that the
-rotation-supported parts of galaxies are always exponential and that
-the kinematically hot components are always more de-Vaucouleurs-like.
+therein); these uses are prone to over-interpretation: There certainly
+could be rotation-supported parts of galaxies that are not exponential
+and kinematically hot components that are not de-Vaucouleurs-like.
 
 The performance advantages we will obtain from using MoG
-approximations are not simply that convolution itself is trivial.  The
-standard unconvolved galaxy models, especially the de~Vaucouleurs and
-S\'ersic models (\citealt{dev}, \citealt{ser}), are very ill-behaved
-near the galaxy center.  Rendering these profiles precisely near the
-center can be very challenging numerically.  In addition, most of the
-rendering time is spent at very small radii, where PSF convolution is
-going to erase all structure anyway.  That is, a MoG description of
-the de~Vaucouleurs profile saves time both in rendering \emph{and} in
-PSF convolution; it produces profiles that are approximate but very
-high performance when all real uses of the profiles are PSF-convolved,
-as they usually are.
+approximations are not simply that it makes the convolution trivial.
+The standard unconvolved galaxy models, especially the de~Vaucouleurs
+and S\'ersic models (\citealt{dev}, \citealt{ser}), are very
+ill-behaved near the galaxy center.  Rendering these profiles
+precisely near the center can be very challenging numerically.
+Computer codes that do this rendering and PSF convolution well use
+enormous numbers of samples; mathematically this sampling
+representation of the models is a mixture of delta functions.  In
+order for these representations of the profiles to be precise, large
+numbers of sample points are required.  The convolution and rendering
+step involves as many PSF evaluations as there are samples---which can
+number in the thousands.  In addition, most of the rendering time is
+spent at very small radii---most of the samples in any integration are
+in or near the central pixel---where PSF convolution is going to erase
+all structure anyway.  That is, a MoG description of the
+de~Vaucouleurs profile saves time both in rendering \emph{and} in PSF
+convolution; it produces code that is very high performance when all
+real uses of the profiles are PSF-convolved, as they usually are.
+
+Not only does a sampling representation of the profiles require many
+more compute operations per render than a MoG representation, it also
+requires more code complexity.  For the sampling to be both accurate
+and efficient, it has to be strongly adaptive (with much higher
+density in the center).  This adds complexity and choices into the
+code, leading to engineering and software maintenance and testing
+overheads that cost not just compute time but human time.  One
+response to this is to build large number of pre-computed profile PSF
+convolutions and interpolate among them.  This is the strategy taken
+in the \project{SDSS} pipelines (\citealt{lupton}), but that requires
+even \emph{more} human overhead, because not only does the sampling
+and convolution have to be done correctly, so does the numerical
+interpolation.
 

### 2013-02-11 documents/mixtures/response_to_referee.txt
two more referee comments

@@
 
+fixed.
+
    Page 6 - Definition of "best": The authors should really have been "bolder" by
@@
    explanation of how object deblending is performed with MoGs would be welcome.
+
+All we are trying to do in this paper is come up with *replacements*
+for the exponential and deVaucouleurs profiles and variants.  We are
+not actually fitting real galaxy images.  Indeed that requires
+deblending and so on.  Literally, all we are doing is replacing the
+exp{r^{1/n}) with a sum of Gaussians.  We are not fitting real scenes.
+
+[ we should make a text change here ]
 

### 2013-02-11 documents/mixtures/mixture_models.tex
two more referee comments

@@
 two-dimensional probability distribution functions (for, say, the
-arrivals of photons), then one natural choice is the K-L divergence or
-similar cross-entropy or information-theoretic measure.  However, in
-typical astronomical imaging, the galaxy is superimposed on a
-substantial, flat sky level, and the noise in the data is close to
-Gaussian.  This suggests more chi-squared-like objectives.  We adopt
-the latter, in part because they are most appropriate for our specific
-proposed application (modeling \sdss-like astronomical imaging), but
-experiments we have performed suggest that information-theoretic
-objectives also lead to good results.
+arrivals of photons), then one natural choice is the Kullback-Leibler
+divergence or similar cross-entropy or information-theoretic measure.
+However, in typical astronomical imaging, the galaxy is superimposed
+on a substantial, flat sky level, and the noise in the data is close
+to Gaussian.  This suggests more chi-squared-like objectives.  We
+adopt the latter, in part because they are most appropriate for our
+specific proposed application (modeling \sdss-like astronomical
+imaging), but experiments we have performed suggest that
+information-theoretic objectives also lead to good results.
 

### 2013-02-15 documents/mixtures/mixture_models.tex
a few more referee points

@@
 overheads that cost not just compute time but human time.  One
-response to this is to build large number of pre-computed profile PSF
+response to this is to build large numbers of pre-computed profile PSF
 convolutions and interpolate among them.  This is the strategy taken
@@
 and convolution have to be done correctly, so does the numerical
-interpolation.
+interpolation.  Finally, a sampling approximation---or any finite
+numerical approximation---is effectively something like a ``mixture of
+delta functions'' or ``mixture of integration kernels'' approximation.
+Any such approximation (at fixed accuracy) will have far more degrees
+of freedom (complexity) than the MoG approximations given below.
 
@@
 \begin{eqnarray}
-p(\pos) &\approx& \sum_{k=1}^K a_k\,\normal(\pos\given\mean_k,\var_k)
+p(\pos) &\approx& \sum_{k=1}^K c_k\,\normal(\pos\given\mean_k,\var_k)
 \\
-1 &=& \sum_{k=1}^K a_k
+1 &=& \sum_{k=1}^K c_k
 \quad ,
@@
 the number of Gaussians used in the MoG, and the $K$ Gaussians
-have amplitudes $a_k$, means $\mean_k$, and variance tensors $\var_k$.
+have amplitudes $c_k$, means $\mean_k$, and variance tensors $\var_k$.
 The sum-to-one condition ensures that the probability distribution
@@
 \end{eqnarray}
-The half-light inverse-radius parameters $\alpha^{\exp}$ and
-$\alpha^{\dev}$---and the softening and cutoff radius parameters---are
+The half-light inverse-radius parameters $\alpha^{\lux}$ and
+$\alpha^{\luv}$---and the softening and cutoff radius parameters---are
 taken from the \sdss\ codebase.
@@
 investigator wants profiles that are normalized to have unit total
-flux (intensity integrated over solid angle).  Although there is an
-analytic result for the dev profile, numerical integration of the
-concentrated profiles dev and luv to determine total fluxes can be
-challenging.  This is not true for the MoG approximations: Each
-Gaussian is normalized, so the sum of the amplitudes $\sum_m
-a^{\luv}_m$ (for the luv profile, say) gives the total flux for the
-MoG approximation to that profile.
+flux (intensity integrated over solid angle).  Although there happens
+to be an analytic result for the dev profile, direct numerical
+integration of the concentrated profiles dev and luv to determine
+total fluxes can be challenging.  This is not true for the MoG
+approximations: Each Gaussian is normalized, so the sum of the
+amplitudes $\sum_m a^{\luv}_m$ (for the luv profile, say) gives the
+total flux for the MoG approximation to that profile.
+
+Since the dev profile has an analytic integral, numerical integration
+of the dev profile is never strictly needed, but the fact that the
+numerical integration is challenging is an argument that the function
+is ill-behaved.  Also, there aren't analytic integrals for most
+partial footprints (that is, definite integrals with finite domain or
+out to finite radius); these are also much easier to evaluate in the
+MoG approximation, since the small Gaussians (in the center) will lie
+(almost) entirely within a finite domain, and the larger Gaussians
+(where numerical integration is required) are easier to sample and
+integrate than the exact profile.
 

### 2013-02-16 documents/mixtures/response_to_referee.txt
a few more referee points
more response; reflow paragraphs

@@
 to engineering overheads in testing and correctness that go way beyond
-the simple numerical speed losses.
+the simple numerical speed losses.  We alsoo bring up the point that
+conventional numerical integration methods effectively treat the
+profiles as very complex mixtures of kernels.
 
-[should we also add a late paragraph re-iterating and showing a comparison?]
+[ should we also add a late paragraph re-iterating and showing a
+comparison? ]
 
@@
 
-   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k say, to
-   avoid confusion with the \alpha's in the following equations.
+   Equations 2 and 3: I suggest you replace a_k by another symbol, c_k
+   say, to avoid confusion with the \alpha's in the following
+   equations.
 
-   Page 6, first line: "The half-light inverse radius parameters \alpha^{exp} and \
-   alpha^{dev} -- and the softening and cutoff radius parameters -- are taken from
-   the SDSS codebase." I think you meant to write \alpha^{lux} and \alpha^{luv} here.
+Good point.  Fixed.
 
-   Page 6, paragraph 2: "Although there is an analytic result for the dev profile,
-   numerical integration of the concentrated profiles dev and luv to determine
-   total fluxes can be challenging."  Why are you flagging the numerical integration of
-    the dev profile as a challenge given that there is an analytic expression for it?
+   Page 6, first line: "The half-light inverse radius parameters
+   \alpha^{exp} and \ alpha^{dev} -- and the softening and cutoff
+   radius parameters -- are taken from the SDSS codebase." I think you
+   meant to write \alpha^{lux} and \alpha^{luv} here.
 
-   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler divergence"
+Absolutely; fixed.
+
+   Page 6, paragraph 2: "Although there is an analytic result for the
+   dev profile, numerical integration of the concentrated profiles dev
+   and luv to determine total fluxes can be challenging."  Why are you
+   flagging the numerical integration of the dev profile as a
+   challenge given that there is an analytic expression for it?
+
+Indeed.  Adjusted that paragraph slightly and added a subsequent
+paragraph explaining the implications for integrals over finite
+domains.
+
+   Page 6, paragraph 2: "K-L divergence" --> "Kullback-Leibler
+   divergence"
 
@@
 
-   Page 6 - Definition of "best": The authors should really have been "bolder" by
-   insisting on the use of an information-theoretic measure here rather than the
-   "traditional" chi-squared. The MoGs are a natural fit to an image entropy measure
-   because maximizing entropy, say, would naturally lead to minimizing the number of
-    components in a MoG. Chi-squared is also biased at low signal-to-noise ratios.
-   Please consider.
+   Page 6 - Definition of "best": The authors should really have been
+   "bolder" by insisting on the use of an information-theoretic
+   measure here rather than the "traditional" chi-squared. The MoGs
+   are a natural fit to an image entropy measure because maximizing
+   entropy, say, would naturally lead to minimizing the number of
+   components in a MoG. Chi-squared is also biased at low
+   signal-to-noise ratios.  Please consider.
 
-   Page 6: Why is "badness" computed in a one-dimensional numerical integral? For
-   the sake of computational speed? What happens when dealing with heavily blended
-   objects? Where is the center for the 1D integral then? Please explain. A general
-   explanation of how object deblending is performed with MoGs would be welcome.
+Chi-squared would definitely be biased if what we were fitting in
+astronomical images were distributions of photon arrivals with no
+background or read noise.  The presence of background (sky photons)
+and read noise makes chi-squared a better statistic for optimization
+than, say, the Poisson likelihood.  However, the *precise* measure you
+would use in *principle* depends (weakly) on the properties of the
+data you have.
+
+[ make text change here ]
+
+As for the information-theoretic things mentioned by the referee: Yes,
+in principle things like AIC and BIC and Bayes integral etc measures
+for "how many components to use" would be very useful.  Unfortunately,
+unlike with the decision about chi-squared vs Poisson likelihood, the
+value of the AIC or BIC does depend very very sensitively on the noise
+model (object, sky, and read noise levels) in every pixel of every
+image.  That is, someone doing this correctly would use a higher K
+expansion for the brighter galaxies and a lower K expansion for the
+fainter galaxies.  That is very relevant and important...
+
+[ we must add text for this ]
+
+   Page 6: Why is "badness" computed in a one-dimensional numerical
+   integral? For the sake of computational speed? What happens when
+   dealing with heavily blended objects? Where is the center for the
+   1D integral then? Please explain. A general explanation of how
+   object deblending is performed with MoGs would be welcome.
 
@@
 
-[ we should make a text change here ]
+In the Tractor papers which are forthcoming, we *are* solving this
+deblending problem, but that is a separate matter, involving image
+likelihoods, noise models, and priors (and utilities, perhaps) on
+parameters, so it is way beyond the scope of this paper.
 
-   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv = 8 -- are good
-    compromises between mixture complexity and quality of fit" warrants more
-   explanation. Everytime I read about using MoGs for galaxy fitting, I have the same
-   reaction: it needs a large number of highly redundant to model something a single
-   exponential profile. Fans of the eigenimage decompositions are probably appalled!
-    Granted the mathematical simplicity of MoGs can make up for this by enabling
-   faster fitting times, but it is definitely an un-appealing aspect of the method.
-   It would be helpful to comment on this to counter a reaction that I think may be
-   shared by others. More importantly, I do not quite understand how you decided on
-    M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for lux has clearly
-   converged and there is no gain in going higher than M=6. I cannot say the same
-   thing for the luv MoG. Badness goes down by a factor of 16 going from M=6 to M=8
-   and then goes down by a factor of 5 going from M=8 to M=10. How can you claim
-   that your badness has "converged" for M=8 then? A quantitative threshold and its
-   rationale are clearly needed here to justify the "good compromises" here. Please
-   be more quantitative or clarify if I misunderstood something.
+In this paper *all* we are doing is the "applied math" problem of
+replacing the analytic profiles with MoG approximations.
 
-   Equation 19: I did not quite understand your distinction between the \Delta X
-   vector here and the x vector in equation 2. Please clarify.
+[ we must make a text change here in response to this issue ]
 
-In other news, we also made the following changes to the text, in
-response to other comments and issues:
+   Page 7, paragraph 2: The statement "These -- M^lux = 6 and M^luv =
+   8 -- are good compromises between mixture complexity and quality of
+   fit" warrants more explanation. Everytime I read about using MoGs
+   for galaxy fitting, I have the same reaction: it needs a large
+   number of highly redundant to model something a single exponential
+   profile. Fans of the eigenimage decompositions are probably
+   appalled!  Granted the mathematical simplicity of MoGs can make up
+   for this by enabling faster fitting times, but it is definitely an
+   un-appealing aspect of the method.  It would be helpful to comment
+   on this to counter a reaction that I think may be shared by
+   others. More importantly, I do not quite understand how you decided
+   on M^lux = 6 and M^luv = 8. If I look at Table 2, the badness for
+   lux has clearly converged and there is no gain in going higher than
+   M=6. I cannot say the same thing for the luv MoG. Badness goes down
+   by a factor of 16 going from M=6 to M=8 and then goes down by a
+   factor of 5 going from M=8 to M=10. How can you claim that your
+   badness has "converged" for M=8 then? A quantitative threshold and
+   its rationale are clearly needed here to justify the "good
+   compromises" here. Please be more quantitative or clarify if I
+   misunderstood something.
 
-- we did foo
+[ response ; what changes ]
 
-- we did bar
+   Equation 19: I did not quite understand your distinction between
+   the \Delta X vector here and the x vector in equation 2. Please
+   clarify.
+
+[ response; what changes ]
+
+[ and search email for other issues arising / suggestions / etc... ]

### 2013-02-17 documents/mixtures/mixture_models.tex
why Delta x and not x

@@
 proportional to the identity or even diagonal because the PSF will not
-in general be round.  An example that illustrates the use of this PSF
-is the following: A star of flux $S_s$ at focal-plane position
-$\pos_s$ will lead to an image (PSF-convolved intensity map) of the
-form
+in general be round.  The reason we use $\Delta\pos$ in the argument
+of $\psi(\cdot)$---and not $\pos$---is that the PSF describes the
+probability distribution for two-dimensional \emph{displacements} of
+detected photons from true source positions.  It is equivalent to say
+that the PSF is a function that is only used to \emph{convolve} (or
+correlate) two-dimensional models; it is never used as an image model
+\emph{itself}.  The PSF is only used in the context of convolving some
+high-resolution model, which might be as complex as a multi-component
+galaxy, or might be as simple as a delta function, in the case of a
+star.
+
+An example that illustrates the use of this PSF is the following: A
+star of flux $S_s$ at focal-plane position $\pos_s$ will lead to an
+image (PSF-convolved intensity map) of the form
 \begin{eqnarray}\displaystyle

### 2013-02-24 documents/mixtures/response_to_referee.txt
why Delta x and not x
close some doors

@@
 We also added an introductory paragraph that points out that the code
-complexity required to do rendering correctly is large and this leads
-to engineering overheads in testing and correctness that go way beyond
-the simple numerical speed losses.  We alsoo bring up the point that
-conventional numerical integration methods effectively treat the
-profiles as very complex mixtures of kernels.
-
-[ should we also add a late paragraph re-iterating and showing a
-comparison? ]
+complexity required to do straight profile rendering (ie without the
+MoGs) correctly is large and this leads to engineering overheads in
+testing and correctness that go way beyond the simple numerical speed
+losses.  We alsoo bring up the point that conventional numerical
+integration methods effectively treat the profiles as very complex
+mixtures of kernels.
 
@@
 paragraph that makes it clearer what the performance gains are.
-
-[ should we also get quantitative at the end? ]
 
@@
 
-[ response; what changes ]
+Agreed.  We added three sentences shortly after equation (19)
+explaining explicitly why we use \Delta x and not x here.  It relates
+to the fact that the PSF is only used as a convolution kernel, never
+as a model in itself.
 

### 2013-03-10 documents/mixtures/response_to_referee.txt
finished responding to referee issues

@@
 and read noise makes chi-squared a better statistic for optimization
-than, say, the Poisson likelihood.  However, the *precise* measure you
-would use in *principle* depends (weakly) on the properties of the
-data you have.
-
-[ make text change here ]
+than, say, the Poisson likelihood.  (That wouldn't be true if we were
+fitting to *actual* images in which we know the sky level and read
+noise, but we aren't.  Here we are just making *replacements* for the
+standard profiles in a generic setting.)  All that said, the *precise*
+measure you would use in *principle* depends (weakly) on the
+properties of the data you have.  We added two sentences to this
+effect in the relevant place (right after the K-L divergence mention).
 
@@
 model (object, sky, and read noise levels) in every pixel of every
-image.  That is, someone doing this correctly would use a higher K
-expansion for the brighter galaxies and a lower K expansion for the
-fainter galaxies.  That is very relevant and important...
-
-[ we must add text for this ]
+image.  That is, someone doing this correctly would use a higher M
+expansion for the brighter galaxies and a lower M expansion for the
+fainter galaxies.  That is very relevant and important.  We added text
+about this where we give the M_exp=6 advice.  We note that although it
+might make sense to go to large M from a data precision point of view,
+it doesn't from a *model* precision point of view, since the models
+don't fit the data at this level (there are spiral arms, deviations
+from ellipsoidal symmetry and so on), as the referee evidently knows!
+We added text about this too.
 
@@
 likelihoods, noise models, and priors (and utilities, perhaps) on
-parameters, so it is way beyond the scope of this paper.
+parameters, so it is way beyond the scope of this paper.  This paper
+is just about the profiles *themselves*, not any specific data
+analysis problem that we can solve with them.
 
 In this paper *all* we are doing is the "applied math" problem of
-replacing the analytic profiles with MoG approximations.
-
-[ we must make a text change here in response to this issue ]
+replacing the analytic profiles with MoG approximations.  As we note
+in the introductory paragraphs, *any* use of deV and exp profiles
+involves some approximation when convolving with a realistic PSF.  We
+are just arguing for using MoG approximations instead of the
+alternatives, most of which are finite, adaptive samplings, which are
+more computationally expensive and less accurate.
 
@@
 
-[ response ; what changes ]
+We don't mean to claim that the badness has converged, and we modified
+the text to say so.  We just mean that as a trade-off between code
+speed and precision, in *practice* this is a good place to stop.
+Again, this is just for *replacing* the analytic profiles, not for
+general galaxy fitting.  For general galaxy fitting, we agree that
+many more Gaussians can be required, especially for nearby
+well-resolved galaxies.
 
@@
 
+In addition to the changes given above, we also made the following small changes:
+
 [ and search email for other issues arising / suggestions / etc... ]
+
+* We added an acknowledgement to the referee.

### 2013-03-10 documents/mixtures/mixture_models.tex
finished responding to referee issues

@@
 imaging), but experiments we have performed suggest that
-information-theoretic objectives also lead to good results.
+information-theoretic objectives also lead to good results.  In
+principle the objective that should be used to make these
+approximations could be slightly different in different contexts.  In
+practice, since the dependence of ``best objective'' on data
+properties is weak and because the point here is to make generic
+replacements for the analytic profiles, chi-squared is a good choice.
 
@@
 their better behaviors numerically, and partly because they and we are
-both part of the \project{SDSS} tradition.  These---$M^{\lux}=6$ and
-$M^{\luv}=8$---are good compromises between mixture complexity ($M$)
-and quality of fit (badness).  Also, even the best-fitting late-type
+both part of the \project{SDSS} tradition.  We find that
+these---$M^{\lux}=6$ and $M^{\luv}=8$---are good compromises between
+mixture complexity ($M$) and quality of fit (badness) for our purposes
+fitting \project{SDSS} data.  We expect this to be true for many
+typical data sets.  In principle, $M$ should rise as the data quality
+improves, because better data requires better models, and as $M$
+increases, the approximations to the analytic profiles improve.  It is
+important to remember, however, if you are tempted to go to very large
+$M$, that this can be effort wasted: Even the best-fitting late-type
 and early-type galaxies deviate from exponential and de~Vaucouleurs
@@
       Jim Bosch (Princeton),
-  and Kevin Bundy (IPMU)
+      Kevin Bundy (IPMU),
+      and an anonymous referee
 for valuable comments.  This work was supported in part by NASA (grant

### 2013-03-10 documents/mixtures/mixture_models.tex
working through email comments

@@
 galaxy images with MoGs has been done very successfully before (for
-example, \citealt{bendinelli, emsellem, bendinelli2, cappellari}; they
+example, \citealt{bendinelli, emsellem, bendinelli2, kochanek, cappellari}; they
 use the name ``multi-Gaussian expansion'' or MGE).  However, the idea
@@
 and convolution have to be done correctly, so does the numerical
-interpolation.  Finally, a sampling approximation---or any finite
-numerical approximation---is effectively something like a ``mixture of
-delta functions'' or ``mixture of integration kernels'' approximation.
-Any such approximation (at fixed accuracy) will have far more degrees
-of freedom (complexity) than the MoG approximations given below.
+interpolation.  Another less common strategy is to find other kinds of
+function expansions (\citealt{bendi}).  Finally, a sampling
+approximation---or any finite numerical approximation---is effectively
+something like a ``mixture of delta functions'' or ``mixture of
+integration kernels'' approximation.  Any such approximation (at fixed
+accuracy) will have far more degrees of freedom (complexity) than the
+MoG approximations given below.
 
@@
 of the two) radial profiles.  In detail, in fact, the \sdss\ Catalog
-models are small modifications of these; details below.  So our goal
+models are small modifications of these; details below.  Our goal
 here is to provide replacements for these models in order to improve
@@
 s~\ref{tab:exp} and \ref{tab:lux} and \figurename s~\ref{fig:exp}
-through \ref{fig:M}.  All the results shown in these figures and
-tables and more are available in machine-readable form from author DWH
+through \ref{fig:M}.  All the results shown in these \figurename s and
+\tablename s and more are available in machine-readable form from author DWH
 upon request.  In the \tablename s and \figurename s we show
@@
 galaxy follows exactly any ser profile---so deprojection of our
-approximations are not that interesting in themselves.  However,
-the general program of fitting two-dimensional sources with MoGs may
-have strong implications in the future for three-dimensional modeling
-and deprojection.
+approximations are not that interesting in themselves.  However, the
+general program of fitting two-dimensional sources with MoGs may have
+strong implications in the future for three-dimensional modeling and
+deprojection; some work has been done in this direction previously
+(\citealt{vdb}); it is promising.
 
@@
 Bendinelli,~O., 1991, \apj, 366, 599
+\bibitem[Bendinelli et al.(1993)]{bendi}
+Bendinelli,~O., Ciotti,~L., \& Parmeggiani,~G.\ 1993, \aap, 279, 668 
 \bibitem[Bendinelli \& Parmeggiani(1995)]{bendinelli2}
@@
 Emsellem,~E., Monnet,~G., Bacon,~R., \& Nieto,~J.-L., 1994, \aap, 285, 739
+\bibitem[Kochanek et al.(2000)]{kochanek}
+Kochanek,~C.~S., Falco,~E.~E., Impey,~C.~D., et al., 2000, \apj, 535, 692
 \bibitem[Lupton \etal(2001)]{lupton}
@@
 Trujillo,~I., Aguerri,~J.~A.~L., Cepa,~J., \& Guti{\'e}rrez,~C.~M., 2001, \mnras, 321, 269 
+\bibitem[van~den~Bosch et al.(2008)]{vdb}
+van~den~Bosch,~R.~C.~E., van de Ven,~G., Verolme,~E.~K., Cappellari,~M., \& de~Zeeuw,~P.~T., 2008, \mnras, 385, 647
 \end{thebibliography}

### 2013-03-10 documents/mixtures/mixture_models.tex
cites and acks

@@
 interpolation.  Another less common strategy is to find other kinds of
-function expansions (\citealt{bendi}).  Finally, a sampling
-approximation---or any finite numerical approximation---is effectively
-something like a ``mixture of delta functions'' or ``mixture of
-integration kernels'' approximation.  Any such approximation (at fixed
-accuracy) will have far more degrees of freedom (complexity) than the
-MoG approximations given below.
+function expansions or approximations (\citealt{bendi},
+\citealt{spergel}).  Finally, a sampling approximation---or any finite
+numerical approximation---is effectively something like a ``mixture of
+delta functions'' or ``mixture of integration kernels'' approximation.
+Any such approximation (at fixed accuracy) will have far more degrees
+of freedom (complexity) than the MoG approximations given below.
 
@@
       Kevin Bundy (IPMU),
+      Luca Ciotti (Bologna),
+      Rachel Mandelbaum (CMU),
+      Chien Peng (OCIW),
       and an anonymous referee
@@
 Simard,~L., Mendel,~J.~T., Patton,~D.~R., Ellison,~S.~L., \& McConnachie,~A.~W., 2011, \apjs, 196, 11
+\bibitem[Spergel(2010)]{spergel}
+Spergel,~D.~N.\ 2010, \apjs, 191, 58
 \bibitem[Trujillo \etal(2001)]{trujillo}

### 2013-03-22 code/sernorm.py
add a little test code for Jim Bosch's sersic half-light radius scaling formula

@@ -1,0 +1,34 @@
+# Just checking out Jim Bosch's formula for the half-light radius scaling
+# "sernorm".
+import matplotlib
+matplotlib.use('Agg')
+import pylab as plt
+import numpy as np
+
+from astrometry.util.plotutils import *
+
+from scipy.special import gammaincinv
+
+def sernorm(n):
+	return gammaincinv(2.*n, 0.5)
+
+ps = PlotSequence('sernorm')
+X = np.linspace(0., 8., 8001)
+dx = (X[1]-X[0])
+X = X[:-1] + dx/2.
+for n in np.arange(1, 6, 0.5):
+	print 'sernorm(%f)' % n, sernorm(n)
+	S = np.exp(-sernorm(n) * (X**(1./n) - 1.))
+	plt.clf()
+	plt.plot(X, S, 'r-')
+	cs = np.cumsum(S*X)*dx
+	plt.plot(X, cs / (cs[-1]), 'r:')
+	plt.axhline(1., color='k', alpha=0.5)
+	plt.axhline(0.5, color='k', alpha=0.5)
+	plt.axvline(1., color='k', alpha=0.5)
+	plt.title('n = %f' % n)
+	#ps.savefig()
+	plt.xscale('log')
+	plt.yscale('log')
+	ps.savefig()
+

### 2013-03-22 code/sernorm.py
Ciotti & Bertin ref fix

@@ -16,19 +16,18 @@ ps = PlotSequence('sernorm')
 X = np.linspace(0., 8., 8001)
 dx = (X[1]-X[0])
 X = X[:-1] + dx/2.
-for n in np.arange(1, 6, 0.5):
+cc = ['r','b','g','m','y','c','k',(1.,0.5,0),'0.3',(0,0.5,1.)]
+plt.clf()
+for i,n in enumerate(np.arange(1, 6, 0.5)):
 	print 'sernorm(%f)' % n, sernorm(n)
 	S = np.exp(-sernorm(n) * (X**(1./n) - 1.))
-	plt.clf()
-	plt.plot(X, S, 'r-')
+	plt.plot(X, S, '-', color=cc[i])
 	cs = np.cumsum(S*X)*dx
-	plt.plot(X, cs / (cs[-1]), 'r:')
-	plt.axhline(1., color='k', alpha=0.5)
-	plt.axhline(0.5, color='k', alpha=0.5)
-	plt.axvline(1., color='k', alpha=0.5)
-	plt.title('n = %f' % n)
-	#ps.savefig()
-	plt.xscale('log')
-	plt.yscale('log')
-	ps.savefig()
-
+	plt.plot(X, cs / (cs[-1]), ':', color=cc[i])
+plt.axhline(1., color='k', alpha=0.5)
+plt.axhline(0.5, color='k', alpha=0.5)
+plt.axvline(1., color='k', alpha=0.5)
+plt.xscale('log')
+plt.yscale('log')
+plt.xlim(X.min(), X.max())
+ps.savefig()

### 2013-03-22 documents/mixtures/mixture_models.tex
Ciotti & Bertin ref fix
Add paragraph about gammaincinv, and call out Bosch in acks

@@
 
+\cite{ciotti} give an equation for $\alpha^{\ser(n)}$ in their
+equation 5:
+\begin{eqnarray}
+\gamma(2 n, \alpha^{\ser(n)}) &=& \frac{\Gamma(2 n)}{2}
+\end{eqnarray}
+where $\gamma$ is the \emph{left} or \emph{lower} incomplete gamma function,
+\begin{eqnarray}
+\gamma(a, x) &=& \int_0^x \exp(-t) \, t^{a-1} \, \mathrm{d} t \quad .
+\end{eqnarray}
+Fortunately for us, this can be evaluated using
+\begin{eqnarray}
+\alpha^{\ser(n)} &=& \mathrm{gammaincinv}(2 n, \, \frac{1}{2})
+\end{eqnarray}
+using the gammaincinv (inverse incomplete gamma) function in the
+\project{scipy.special} library.
+
 The \sdss\ pipelines (\citealt{lupton}) make use of modified profiles,
@@
       Brendon Brewer (Auckland),
-      Jim Bosch (Princeton),
       Kevin Bundy (IPMU),
@@
       and an anonymous referee
-for valuable comments.  This work was supported in part by NASA (grant
+for valuable comments.
+%
+We thank Jim Bosch (Princeton) for bringing to our attention the
+\emph{gammaincinv} function.
+%
+  This work was supported in part by NASA (grant
 NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
@@
 \bibitem[Ciotti \& Bertin(1999)]{ciotti}
-Ciotti,~L. \& Bertin,~G., 1999, \aa, 352, 447
+Ciotti,~L. \& Bertin,~G., 1999, \aap, 352, 447
 \bibitem[de~Vaucouleurs(1948)]{dev}

### 2013-03-22 documents/mixtures/response_to_referee.txt
working through email comments
cites and acks
em-dash between K&L in K--L div; alsoo -> also

@@
 testing and correctness that go way beyond the simple numerical speed
-losses.  We alsoo bring up the point that conventional numerical
+losses.  We also bring up the point that conventional numerical
 integration methods effectively treat the profiles as very complex
@@
 
-In addition to the changes given above, we also made the following small changes:
+In addition to the changes given above, we also made the following
+small changes:
 
-[ and search email for other issues arising / suggestions / etc... ]
+* Added references to Kochanek 2000 and van den Bosch 2008 and
+  Bendinelli 1993 and Spergel 2010.
 
-* We added an acknowledgement to the referee.
+* Fixed a few typos.
+
+* Added some acknowledgements, including one to the referee.

### 2013-03-22 documents/mixtures/mixture_models.tex
em-dash between K&L in K--L div; alsoo -> also

@@
 two-dimensional probability distribution functions (for, say, the
-arrivals of photons), then one natural choice is the Kullback-Leibler
+arrivals of photons), then one natural choice is the Kullback--Leibler
 divergence or similar cross-entropy or information-theoretic measure.

### 2013-03-23 documents/mixtures/mixture_models.tex
tweak text, fix amplitude notation throughout

@@
 \begin{eqnarray}\displaystyle
-Q^{\exp}(\spos) &\approx& \sum_{m=1}^{M^{\exp}} a^{\exp}_m\,\normal(\spos\given\zero,\var^{\exp}_m)
+Q^{\exp}(\spos) &\approx& \sum_{m=1}^{M^{\exp}} c^{\exp}_m\,\normal(\spos\given\zero,\var^{\exp}_m)
 \\
@@
 \\
-Q^{\dev}(\spos) &\approx& \sum_{m=1}^{M^{\dev}} a^{\dev}_m\,\normal(\spos\given\zero,\var^{\dev}_m)
+Q^{\dev}(\spos) &\approx& \sum_{m=1}^{M^{\dev}} c^{\dev}_m\,\normal(\spos\given\zero,\var^{\dev}_m)
 \\
@@
 \begin{eqnarray}
-\{a^{\exp}_m,v^{\exp}_m\}_{m=1}^{M^{\exp}}~,~\{a^{\dev}_m,v^{\dev}_m\}_{m=1}^{M^{\dev}}
+\{c^{\exp}_m,v^{\exp}_m\}_{m=1}^{M^{\exp}}~,~\{c^{\dev}_m,v^{\dev}_m\}_{m=1}^{M^{\dev}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\alpha^{\ser(n)} &=& \mathrm{gammaincinv}(2 n, \, \frac{1}{2})
+\alpha^{\ser(n)} &=& \texttt{gammaincinv}(2 n, \, \frac{1}{2})
 \end{eqnarray}
-using the gammaincinv (inverse incomplete gamma) function in the
+using the \texttt{gammaincinv} (inverse incomplete gamma) function in the
 \project{scipy.special} library.
@@
 approximations: Each Gaussian is normalized, so the sum of the
-amplitudes $\sum_m a^{\luv}_m$ (for the luv profile, say) gives the
+amplitudes $\sum_m c^{\luv}_m$ (for the luv profile, say) gives the
 total flux for the MoG approximation to that profile.
@@
 In \figurename~\ref{fig:ser}, we show the dependence of amplitudes
-$a^{\ser(n)}_m$ and variances $v^{\ser(n)}_m$ on the ser index $n$.
+$c^{\ser(n)}_m$ and variances $v^{\ser(n)}_m$ on the ser index $n$.
 There is clearly continuity; a valuable follow-up project would be to
@@
 \begin{eqnarray}\displaystyle
-I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,a^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
+I(\pos\given\exp,S_g,\pos_g,\affine_g) &=& \sum_{k=1}^K \sum_{m=1}^{M^{\exp}} S_g\,c^{\exp}_m\,p_k\,\normal(\pos\given\pos_g+\mean_k,\var_{gm}+\var_k)
 \\
@@
 We thank Jim Bosch (Princeton) for bringing to our attention the
-\emph{gammaincinv} function.
+formula involving \texttt{gammaincinv}.
 %
-  This work was supported in part by NASA (grant
-NNX12AI50G) and the NSF (grant IIS-1124794).  All the results shown in
-the figures and tables and more are available in machine-readable form
-from author DWH upon request.
+This work was supported in part by NASA (grant NNX12AI50G) and the NSF
+(grant IIS-1124794).  All the results shown in the figures and tables
+and more are available in machine-readable form from author DWH upon
+request.  This research made use of the Python projects
+\project{numpy} and \project{scipy} and the NASA \project{Astrophysics
+  Data System}.
 
@@
   equivalent to those in the top-right of \figurename~\ref{fig:exp}.
-  \textsl{bottom-right:} The dependence on the amplitudes $a^{\ser(n)}_m$
+  \textsl{bottom-right:} The dependence on the amplitudes $c^{\ser(n)}_m$
   and root-variances $\sqrt{v^{\ser(n)}_m}$ on ser index $n$.\label{fig:ser}}
@@
 \caption{Comparisons of approximations.  \textsl{top-left:} The
-  dependence of amplitude $a^{\exp}_m$ and root-variance
+  dependence of amplitude $c^{\exp}_m$ and root-variance
   $\sqrt{v^{\exp}_m}$ on $M^{\exp}$ for the exp profile.

### 2013-03-23 code/pickle_to_latex.py

@@ -50,7 +50,7 @@ def write_table(model):
     fd.write(r"$M^{\%s}=$ &" % model + "\n")
     write_table_line(fd, [r"\multicolumn{2}{|c|}{$%d$}" % K for K in Ks])
     fd.write(r"$m$ &" + "\n")
-    write_table_line(fd, [r"$a^{\%s}_m$ & $\sqrt{v^{\%s}_m}$" % (model, model) for K in Ks])
+    write_table_line(fd, [r"$c^{\%s}_m$ & $\sqrt{v^{\%s}_m}$" % (model, model) for K in Ks])
     for k in range(np.max(Ks)):
         fd.write(r"$%d$ &" % (k + 1) + "\n")
         aks = []
@@ -64,7 +64,7 @@ def write_table(model):
                 svks += [r"~"] # [r"\cdots"]
         write_table_line(fd, [r"$%s$ & $%s$" % (a, sv) for a, sv in zip(aks, svks)])
     fd.write(r"\hline" + "\n")
-    fd.write(r"$\sum_m a^{\%s}_m=$ &" % model + "\n")
+    fd.write(r"$\sum_m c^{\%s}_m=$ &" % model + "\n")
     write_table_line(fd, [r"\multicolumn{2}{|c|}{$%6.3f$}" % q for q in [np.sum(pars[i][0:K]) for i, K in enumerate(Ks)]])
     fd.write(r"badness~$=$ &" + "\n")
     MR = 8.

### 2013-03-23 code/plot_mixture_profiles.py

@@ -47,7 +47,7 @@ def plot_mixtures_vs_model(K):
         plt.annotate(label, [sigmas[-1], amps[-1]], xytext=[4,-4+tweak], textcoords="offset points", va="baseline", ha="left")
     model = "ser"
     plt.xlabel(r"root-variance $\sqrt{v^{\mathrm{%s}}_m}$ (units of half-light radii)" % model)
-    plt.ylabel(r"amplitudes $a^{\mathrm{%s}}_m$" % model)
+    plt.ylabel(r"amplitudes $c^{\mathrm{%s}}_m$" % model)
     plt.title(r"approximations to ser profiles at $M^{\mathrm{ser}}=%d$" % K)
     plt.loglog()
     plt.xlim(0.5e-4, 2.e1)
@@ -77,7 +77,7 @@ def plot_mixtures_vs_K(model):
         plt.annotate(label, [sigmas[0], amps[0]], xytext=[-4,-4], textcoords="offset points", va="baseline", ha="right")
         plt.annotate(label, [sigmas[-1], amps[-1]], xytext=[4,-4+tweak], textcoords="offset points", va="baseline", ha="left")
     plt.xlabel(r"root-variance $\sqrt{v^{\mathrm{%s}}_m}$ (units of half-light radii)" % model)
-    plt.ylabel(r"amplitudes $a^{\mathrm{%s}}_m$" % model)
+    plt.ylabel(r"amplitudes $c^{\mathrm{%s}}_m$" % model)
     plt.title(r"approximations to %s" % model)
     plt.loglog()
     plt.xlim(0.5e-4, 2.e1)

### 2013-03-23 documents/mixtures/mixture_models.tex
say less about a^ser(n)

--- 2013-03-22 documents/mixtures/mixture_models.tex
@@
 \end{eqnarray}
-where $\gamma$ is the \emph{left} or \emph{lower} incomplete gamma function,
-\begin{eqnarray}
-\gamma(a, x) &=& \int_0^x \exp(-t) \, t^{a-1} \, \mathrm{d} t \quad .
-\end{eqnarray}
+where $\gamma$ is the lower incomplete gamma function.
 Fortunately for us, this can be evaluated using
@@
 \end{eqnarray}
-using the gammaincinv (inverse incomplete gamma) function in the
-\project{scipy.special} library.
+where gammaincinv is the inverse incomplete gamma function implemented
+in the \project{scipy.special} library.
 

### 2013-03-23 documents/mixtures/mixture_models.tex
add footnote to scipy
dstn style -> hogg style

--- 2013-03-23 documents/mixtures/mixture_models.tex
@@
 
-\cite{ciotti} give an equation for $\alpha^{\ser(n)}$ in their
-equation 5:
+The general equation for $\alpha^{\ser(n)}$ is (\citealt{ciotti} equation~5):
 \begin{eqnarray}
 \gamma(2 n, \alpha^{\ser(n)}) &=& \frac{\Gamma(2 n)}{2}
+\\
+\gamma(q, x) &\equiv& \int_0^x \exp(-t) \, t^{q-1} \, \mathrm{d}t
+\quad ,
 \end{eqnarray}
-where $\gamma$ is the \emph{left} or \emph{lower} incomplete gamma function,
-\begin{eqnarray}
-\gamma(a, x) &=& \int_0^x \exp(-t) \, t^{a-1} \, \mathrm{d} t \quad .
-\end{eqnarray}
-Fortunately for us, this can be evaluated using
+where $\gamma(\cdot)$ is the \emph{left} or \emph{lower} incomplete
+gamma function.  Fortunately for us, this can be evaluated using
 \begin{eqnarray}

### 2013-03-24 documents/mixtures/mixture_models.tex

@@
 \end{eqnarray}
-where $\gamma(\cdot)$ is the \emph{left} or \emph{lower} incomplete
-gamma function.  Fortunately for us, this can be evaluated using
+where $\gamma(\cdot)$ is the lower incomplete gamma function.
+Fortunately for us, this can be evaluated using
 \begin{eqnarray}
@@
 \end{eqnarray}
-using the \texttt{gammaincinv} (inverse incomplete gamma) function in the
-\project{scipy.special} library.
+where \texttt{gammaincinv} is the inverse incomplete gamma function implemented
+in the \project{scipy.special} library.\footnote{\url{http://scipy.org}}
 

### 2013-03-24 code/optimize_mixture_profiles.py
badness fiddling

@@ -146,8 +146,8 @@ def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation):
     Compute chi-squared deviation between profile and
     mixture-of-Gaussian approximation, where the chi-squared is
     appropriate for two-dimensional intensity fitting.
-
-    Note magic number in penalty for large variance.
+    
+    Magic number in extrabadness.
     """
     pars = np.exp(lnpars)
     x = np.arange(0.0005, max_radius, 0.001)
@@ -155,11 +155,13 @@ def badness_of_fit(lnpars, model, max_radius, log10_squared_deviation):
     # note radius (x) weighting, this makes the differential proportional to 2 * pi * r * dr
     numerator = np.sum(x * residual * residual)
     denominator = np.sum(x)
-    badness = numerator / denominator / 10.**log10_squared_deviation
+    badness = numerator / denominator
     K = len(pars) / 2
     var = pars[K:]
-    extrabadness = 0.0001 * np.sum(var) / max_radius**2
-    return badness + extrabadness
+    extrabadness = 1e-10 * np.mean(var / max_radius**2)
+    if extrabadness > badness:
+        print "EXTRABAD:", model, pars, max_radius, log10_squared_deviation
+    return (badness + extrabadness) / 10.**log10_squared_deviation
 
 def negative_score(lnpars, model, max_radius, log10_squared_deviation):
     """

### 2013-03-24 documents/mixtures/mixture_models.tex
typesetting

@@
 \end{eqnarray}
-where \texttt{gammaincinv} is the inverse incomplete gamma function implemented
+where $\texttt{gammaincinv}(\cdot)$ is the inverse incomplete gamma function implemented
 in the \project{scipy.special} library.\footnote{\url{http://scipy.org}}
